### 2016-10-14 paper/paper.tex
more fixes

\documentclass{sig-alternate-05-2015}

\usepackage{url}                  % format URLs
\usepackage{listings}          % format code
\lstset{
  mathescape, 
  language={C},
  basicstyle=\small
}
\usepackage{enumitem}      % adjust spacing in enums
\usepackage[colorlinks=true,allcolors=blue,breaklinks,draft=false]{hyperref}   
\usepackage{graphicx}
\usepackage{float}
\usepackage{subcaption}
\usepackage{xspace,framed}
\usepackage{colortbl}
\usepackage{calc}
\usepackage[dvipsnames]{xcolor}
\usepackage{amssymb,amsmath,amsfonts} 
\usepackage{mathtools}

\usepackage{algorithm}
\usepackage{amsfonts}
\usepackage{multicol}
\usepackage{multirow}
\usepackage{tikz}
\usetikzlibrary{positioning, automata, shapes.arrows, calc, shapes, arrows}
\usepackage[justification=centering]{caption}
\usepackage{stmaryrd}
\usepackage{hhline}
\usepackage{pifont}
\usepackage{longtable}
\usepackage{afterpage}
\usepackage{wasysym}
%\usetikzlibrary{decorations}
%\usetikzlibrary{decorations.pathmorphing}

\newcommand{\blue}[1]{{\color{blue}#1}}
\newcommand{\red}[1]{{\color{red}#1}}
\newcommand{\green}[1]{{\color{green}#1}}

\newtheorem{myassumption}{Assumption}
\newtheorem{mylemma}{Lemma}
\newtheorem{myprop}{Proposition}

\include{macros}

\begin{document}

% Copyright
\setcopyright{acmcopyright}

% DOI
\doi{}

% ISBN
\isbn{}

%
% --- Author Metadata here ---
%\conferenceinfo{WOODSTOCK}{'97 El Paso, Texas USA}
%\CopyrightYear{2007} % Allows default copyright year (20XX) to be over-ridden - IF NEED BE.
%\crdata{0-12345-67-8/90/01}  % Allows default copyright data (0-89791-88-6/97/05) to be over-ridden - IF NEED BE.
% --- End of Author Metadata ---

\title{
Sound and Automated Synthesis of Digital Stabilizing Controllers for Continuous Plants
%Counterexample-guided Synthesis of Closed-loop Digital Control Systems
}
%\subtitle{[Extended Abstract]
%\titlenote{A full version of this paper is available as
%\textit{Author's Guide to Preparing ACM SIG Proceedings Using
%\LaTeX$2_\epsilon$\ and BibTeX} at
%\texttt{www.acm.org/eaddress.htm}}}
%
% You need the command \numberofauthors to handle the 'placement
% and alignment' of the authors beneath the title.
%
% For aesthetic reasons, we recommend 'three authors at a time'
% i.e. three 'name/affiliation blocks' be placed beneath the title.
%
% NOTE: You are NOT restricted in how many 'rows' of
% "name/affiliations" may appear. We just ask that you restrict
% the number of 'columns' to three.
%
% Because of the available 'opening page real-estate'
% we ask you to refrain from putting more than six authors
% (two rows with three columns) beneath the article title.
% More than six makes the first-page appear very cluttered indeed.
%
% Use the \alignauthor commands to handle the names
% and affiliations for an 'aesthetic maximum' of six authors.
% Add names, affiliations, addresses for
% the seventh etc. author(s) as the argument for the
% \additionalauthors command.
% These 'additional authors' will be output/set for you
% without further effort on your part as the last section in
% the body of your article BEFORE References or any Appendices.

%\numberofauthors{8} %  in this sample file, there are a *total*
% of EIGHT authors. SIX appear on the 'first-page' (for formatting
% reasons) and the remaining two appear in the \additionalauthors section.
%
%% \author{
%% % You can go ahead and credit any number of authors here,
%% % e.g. one 'row of three' or two rows (consisting of one row of three
%% % and a second row of one, two or three).
%% %
%% % The command \alignauthor (no curly braces needed) should
%% % precede each author name, affiliation/snail-mail address and
%% % e-mail address. Additionally, tag each line of
%% % affiliation/address with \affaddr, and tag the
%% % e-mail address with \email.
%% %
%% % 1st. author
%% \alignauthor
%% Ben Trovato\titlenote{Dr.~Trovato insisted his name be first.}\\
%%        \affaddr{Institute for Clarity in Documentation}\\
%%        \affaddr{1932 Wallamaloo Lane}\\
%%        \affaddr{Wallamaloo, New Zealand}\\
%%        \email{trovato@corporation.com}
%% % 2nd. author
%% \alignauthor
%% G.K.M. Tobin\titlenote{The secretary disavows
%% any knowledge of this author's actions.}\\
%%        \affaddr{Institute for Clarity in Documentation}\\
%%        \affaddr{P.O. Box 1212}\\
%%        \affaddr{Dublin, Ohio 43017-6221}\\
%%        \email{webmaster@marysville-ohio.com}
%% }

\author{Author names omitted for review}

\newcommand\tool{{\sf DSSynth}\xspace}

\maketitle

\begin{abstract}
%
Modern control is implemented using digital microcontrollers, which are
designed to work in an embedded environment, within a dynamical plant that
represents physical components.
%
We present a new algorithm based on counterexample guided inductive
synthesis that automates the design of digital controllers that are
correct by construction.  The synthesis result is sound with respect to the
complete range of approximations, including time discretisation,
quantisation effects, and finite-precision arithmetic and their related
rounding errors.
%
We have implemented our new algorithm in a tool called \tool, and are able
to automatically generate stable controllers for a set of intricate plant
models taken from the literature within minutes.
%
\end{abstract}

%
%  Use this command to print the description
%
\printccsdesc

% We no longer use \terms command
%\terms{Theory}

\keywords{Control synthesis, CEGIS, finite word length, time sampling, quantisation, interval analysis}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Introduction}

Modern implementations of embedded control systems have proliferated with the
availability of low-cost devices that can perform highly non-trivial control
tasks, with significant impact in numerous application areas such as
environmental control and robotics~\cite{astrom1997computer, Franklin15}.

Correct and sound control is non-trivial, however. The problem is exacerbated by
artefacts specific to digital control, such as the effects of
finite-precision arithmetic, time discretisation, and quantisation noise
introduced by A/D and D/A conversion.  Thus, the programming is a key
barrier to broad adoption of digital control, and requires considerable
expertise.

Beyond classical a-posteriori validation procedures in digital control, 
there has been plenty of previous work aiming at \emph{verifying} a given
controller designed by the engineer.  Recent work can show the stability of 
digital controllers with considerations to implementation aspects, i.e.,
fixed-point arithmetic and the word length used~\cite{Bessa16}.  This work
exploits advances in bit-accurate verification of C programs to obtain a
verifier for software-implemented digital control.

By contrast, we leverage very recent step-change in the scalability of
\emph{program synthesis}.  Program synthesis engines use a program
specification as the starting point, and subsequently generate a sequence of
candidate programs.  The candidate programs are iteratively refined to match
the specification more closely.  Program synthesisers that implement
Counter-Example Guided Inductive Synthesis (CEGIS)~\cite{sketch} are now
able to generate programs for highly non-trivial specifications with a very
high degree of automation.  Modern CEGIS engines implement combinations of
automated testing, genetic algorithms and SMT-based automated
reasoning~\cite{DBLP:conf/lpar/DavidKL15, DBLP:conf/cav/0001A14}.

By applying a state-of-the-art CEGIS engine we are now able to present a
tool that automatically \emph{generates} digital controllers for a given
continuous plant model that are correct by construction.  This approach promises to
reduce the cost and time of development of digital control dramatically, and
requires considerably less expertise than methodologies driven by
verification.  Specifically, we synthesise controllers for hybrid
closed-loop systems, i.e., software-implemented embedded controllers along
with a model of their physical environment (the plant), that are
\emph{stable}.  Due to the complexity of such systems, in this work we focus on linear
models with known configurations and perform parametric synthesis of stabilising digital controllers.

Our work addresses particularly challenging aspects of the control synthesis
problem.  We perform digital control synthesis over a hybrid model, where
the plant exhibits continuous behaviour whereas the controller operates in 
discrete time and over a quantised domain.  
Inspired by a classical approach \cite{astrom1997computer}, 
we translate the problem into a single digital domain.  
Given our primary interest in stability despite digitisation effects, we have modelled
a digital equivalent of the plant by evaluating the effects of the
quantisers (A/D and D/A converters) as time discretisation elements with
quantisation noise.  The resulting closed-loop system is a program loop that
operates on bit-vectors encoded using fixed-point arithmetic with finite-word
length (FWL). The deterministic effects of the FWL and
quantisation errors are incorporated into the model as an additive
uncertainty, which is taken into account during the CEGIS-based synthesis of
the control software for the plant.

In summary, this paper makes the following original contributions.
%
\begin{itemize}

\item We design and implement an automatic approach for generating
  {\em corect-by-construction} digital controllers that is based on
  state-of-the-art inductive synthesis techniques.  Our application of
  program synthesis is non-trivial and requires addressing challenges
  specific to embedded control systems, such as the effects of quantisers and
  FWL. In particular, we have found that a two-stage verification engine
  that continuously refines the precision of the fixed-point
  representation is key to the performance of our approach yielding,
  in most cases, a speed-up of at least two orders of magnitude over a
  conventional one-stage verification engine.

\item We provide experimental results showing that our tool is able to
synthesise stable controllers for a set of intricate benchmarks taken from
the literature efficiently: the median runtime for our benchmark set is
$197$s, i.e., half of the controllers can be synthesised in less than five
minutes.

%% \item We exploit a CEGIS-based approach that uses inductive synthesis
%% together with a bit-precise formal verification algorithm for synthesising
%% correct-by-design robust digital controllers for non-fragile closed-loop
%% stability.

%% \item Experimental results with our synthesiser show that the
%% combination of interval arithmetic to represent the controller coefficients
%% and a two-stage verification approach to continuously refine precision
%% enables us to compute a parametric, sound solution for $14$ out of $15$
%% digital stabilizing controllers for a set of continuous and discrete plants
%% with parametric uncertainties.

\end{itemize}

%\begin{figure}
%\centering
%\resizebox{.35\textwidth}{!}{
% \begin{tikzpicture}[scale=0.3,-,>=stealth',shorten >=.2pt,auto,
%     semithick, initial text=, ampersand replacement=\&,]

%  \matrix[nodes={draw, fill=none, shape=rectangle, minimum height=.2cm, minimum width=.2cm, align=center}, row sep=.5cm, column sep=.5cm] {
%   \coordinate (aux0);
%   \& \node[circle] (circle) {}; 
%   \& \node[fill=yellow!20] (da) {{\sc D/A}}; 
%   \& \node[fill=yellow!20] (p) {{\sc P}};
%   \& \node[fill=yellow!20] (ad) {{\sc A/D}}; 
%   \& \coordinate (aux1);\\
%   \& \coordinate (aux3); 
%   \&
%   \&
%   \node[fill=yellow!20] (c) {C};
%   \& 
%   \& \coordinate (aux2);\\
%  };

%  \path[->] (aux0) edge (circle.west);
%  \path  
%   (circle.east) edge (da.west)
%   (da.east) edge (p.west)
%   (p.east) edge (ad.west)
%   (ad.east) edge (aux1.west)
%   (aux1.south) edge (aux2.north)
%   (aux2.west) edge (c.east)
%   (c.west) edge (aux3.east); 
%  \path[->]  (aux3.north) edge (circle.south);
% \end{tikzpicture}
%}
% \caption{Closed-Loop System. \label{fig:closed-system}}
%\end{figure}

We cover preliminaries on plant discretisation, the verification of
closed-loop control systems, and stability in
Section~\ref{sec:preliminaries}.  Our core contribution is in
Section~\ref{sec:synthesis}: we present our CEGIS-based algorithm for
generating stable controllers for a given plant model.  Our experimental
setup and experimental results are given in Section~\ref{sec:experiments},
and we discuss how we extend the state-of-the-art beyond the existing
literature in Section~\ref{sec:related}. We conclude and describe future
work in Section~\ref{sec:conclusions}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Preliminaries}\label{sec:preliminaries}

%\blue{Please insert a figure clearly distinguishing between continuous and discretised/quantised plant - a template is in a picture sent out by Lucas late last week. Please introduced signals on the plot.

%Let us straighten the notations next. 
%We should have 
%\begin{itemize}
%\item 
%$G(s) = \frac{s^{M_G}b_{0}+s^{M_G-1}b_{1}+...+b_{M_G}}{s^{N_G}a_{0}+s^{N_G-1}a_{1}+...+a_{N_G}}$ -- this is the physical plant, which is focus of interest and to which we refer our specifications/properties 
%\item 
%$G(z, T) = \frac{\bar b_{0}+\bar b_{1}z^{-1}+...+\bar b_{M_G}z^{-M_G}}{\bar a_{0}+\bar a_{1}z^{-1}+...+\bar a_{N_G}z^{-N_G}}$ -- this is the time-discretised plant (notice the notation encompasses explicitly the sample time $T$; please write out explicitly the relationship between the $a_i$ and the $\bar a_i$ and likewise for the numerator's coefficients 
%(in particular, are the orders $M_G, N_G$ the same? 
%\item 
%we have $Y(z) = G(z,T) U(z)$; then, we introduce $\hat Y(z) \doteq Y(z) + \nu (z)$, so that 
%$$
%\frac{\hat Y(z)}{U(z)} = G(z,T) + \frac{\nu (z)}{U(z)} = G(z,T) + \Delta G_Q(z) \doteq \hat G(z) ,  
%$$ 
%where $\Delta G_Q(z)$ is an error term accounting for the quantisation effects 
%\item 
%finally, we encompass the errors resulting from employing finite parameters representation, 
%which we encompass in a term $\Delta G_P(z)$: is this error additive, so that we would get to a 
%$$
%\tilde G(z) \doteq \hat G(z) + \Delta G_P(z) = G(z,T) + \Delta G_Q(z) + \Delta G_P(z)?  
%$$ 
%Dario, can you write it out explicitly? In other words, if 
%$$
%\hat G(z) = G(z,T) + \Delta G_Q(z) = \frac{\hat b_{0}+\hat b_{1}z^{-1}+...+\hat b_{M_G}z^{-M_G}}{\hat a_{0}+\hat a_{1}z^{-1}+...+\hat a_{N_G}z^{-N_G}},  
%$$ 
%then what is the explicit expression of 
%$$
%\tilde G(z) = \frac{\tilde b_{0}+\tilde b_{1}z^{-1}+...+\tilde b_{M_G}z^{-M_G}}{\tilde a_{0}+\tilde a_{1}z^{-1}+...+\tilde a_{N_G}z^{-N_G}}?
%$$
%(Notice that the orders $M_G, N_G$ need to be properly taken care of.)
%\item 
%Now, it's over the transfer function $\tilde G(z)$ that we want to synthesise the parameters of a template $C(z)$. 
%This controller's transfer function will be subject to the same quantisation and precision limits as the plant's: 
%we need to argue that the way we do synthesis affects manipulation of quantities that relate to specific errors and be explicit about these. 
%\item[]
%\item 
%An additional source of error can be related to the fact that a plant transfer function, 
%per se a physical entity, is actually represented in a program imprecisely: 
%this would mean that $G(s)$ is actually 
%$$
%G^\ast (s) \doteq \frac{s^{M_G^\ast}b_{0}+s^{M_G^\ast-1}b_{1}+...+b_{M_G}}{s^{N_G^\ast}a_{0}+s^{N_G^\ast-1}a_{1}+...+a_{N_G}}, 
%$$ 
%which perhaps can indeed be re-written as $G^\ast (s) = G(s) + \Delta G_P(s)$? (This proof should come from that above.) 
%We should argue that this error is encompassed by the one derived above (or isn't it?). 
%\end{itemize}
%}

%------------------------------------
\subsection{Discretising the Plant}
\label{ssec:SandH}
%------------------------------------

The digital controllers synthesised using the algorithm we present in this
paper are typically used in closed loops with continuous (physical)
plants.  Thus, the proposed approach has to consider dynamics of continuous
parts of the system (i.e., the plant) and the discrete parts (i.e., the
digital controller).  To obtain a joint model, we fully discretise the continuous
plant. We briefly recall the standard foundations of this step.

We only consider transfer function models, and require a $z$-domain
transfer function $G(z)$ that captures all aspects of the continuous plant (with a
Laplace domain transfer function $G(s)$) and the sampling and hold process.
%Note that $G(z)$ is a discrete model of a physical, continuous
%plant, i.e., 
The continuous model of the plant must be discretised to obtain the corresponding 
%discrete equivalent 
coefficients of $G(z)$.
%via ZOH discretisation.

%\red{[why do we discuss the ZOH at this point?]}

Amongst the several discretisation methods available in the
literature~\cite{Franklin15}, the zero-order hold (ZOH) discretisation is
considered to be suitable for representing the sample and hold processes that are
typically employed in complex systems~\cite{istepanian2012digital}. 
The ZOH discretisation models the exact effect of sampling and DAC interpolation over the plant.

\begin{myassumption}
%
The sample and hold effects of the ADC and the ZOH of the DAC are synchronised,
i.e., there is no delay between sampling the plant output at ADC and
updating the DAC accordingly.  The DAC interpolator is an ideal ZOH. 
%
\end{myassumption}

\begin{mylemma}\cite{astrom1997computer}
%
Given a synchronised ZOH input and sample and hold output on the plant with
a sample time $T$ satisfying Nyquist criterion, the discrete pulse
transfer function $G(z,T)$ is an exact z-domain representation of $G(s)$ 
that can be computed using the following formula:
%
\begin{equation}
\label{eq:pulsetf}
G(z,T) = %\mathcal{ZOH}\left\lbrace{G(s)}\right\rbrace =
(1-z^{-1})\mathcal{Z}\left\lbrace{\mathcal{L}^{-1}\left\lbrace{\frac{G(s)}{s}}\right\rbrace_{t=kT}}\right\rbrace.
\end{equation}
%
For the sake of brevity, we will use the notation $G(z)$ to represent the
pulse transfer function.
%
\end{mylemma}

Lemma~\ref{eq:pulsetf} ensures the poles and zeros matching under the 
$\mathcal{Z}\left\lbrace{\mathcal{L}^{-1}\left\lbrace{\cdot}\right\rbrace_{t=kT}}\right\rbrace$
operations, 
and it includes the ZOH dynamics in the $(1-z^{-1})$ term, 
which is sufficient for stability studies over $G(s)$.

%------------------------------------
\subsection{Verifying Closed-loop Control Systems}
\label{verifying-closed-loop-control-systems}
%------------------------------------

%\red{[This section and the next should be merged into a crisp section with the background on DSVerifier, if not included in previous work? 
%Please stick to the introduced notations. Please clarify what the function FWL does, compared to the general approximations introduced above.]}

%\red{Alessandro, I've made the changes as requested. Please let me know if this is good enough or if I made it too short}

%\textcolor{blue}{[IB: I'm proposing some changes in this section based on Alessandro's comments and trying to obtain a simpler and uniform notation.]}

CEGIS-based control synthesis requires a formal verifier to check whether a
candidate controller meets the requirements when combined with the plant. 
We use the Digital-System Verifier (DSVerifier)~\cite{IsmailBCFF15} for this
purpose.  It checks the stability of closed-loop control systems and
considers finite-word length (FWL) effects in the digital controller and
uncertainty parameters in the plant model (plant intervals)~\cite{Bessa16}. 
Let $C(z)$ be a digital controller and $G(z)$ be a plant model given as
%
\begin{align}
\small
\label{controller_plant_tf}
C(z)&=\frac{C_n(z)}{C_d{z}}=\frac{\beta_{0}+\beta_{1}z^{-1}+...+\beta_{M_C}z^{-M_C}}{\alpha_{0}+\alpha_{1}z^{-1}+...+\alpha_{N_C}z^{-N_C}}, \\
%\label{plant_tf}
G(z)&=\frac{G_n(z)}{G_d(z)}=\frac{b_{0}+b_{1}z^{-1}+...+b_{M_G}z^{-M_G}}{a_{0}+a_{1}z^{-1}+...+a_{N_G}z^{-N_G}}.
\end{align}
%
\noindent where $\vec{\beta}$ and $\vec{\alpha}$ are vectors containing the
controller's coefficients; similarly $\vec{b}$ and $\vec{a}$ encompass the
plant's coefficients; and $N_{(\cdot)}$ and $M_{(\cdot)}$ denote the number
of poles and zeros for the corresponding block.

Uncertainties in $G(z)$ may appear due to uncertainties in $G(s)$
(\emph{i.e.} $\Delta_p{G}_n(s),\Delta_p{G}_d(s)$ which come from tolerances/imprecisions in the original model), errors in the numeric
calculations, or roundoffs in the quantisation.  These uncertainties are 
parametrically expressed by additive terms, resulting in a uncertain model
$\hat{G}(z)$, such that:
%
\begin{equation}
\hat{G}(z)=\frac{G_n(z)+\Delta G_n(z)}{G_d(z)+\Delta G_d(z)},
\end{equation}
%
which will be represented by the following transfer function:
%
\begin{equation}
\hat{G}(z)=\frac{\hat{b}_{0}+\hat{b}_{1}z^{-1}+...+\hat{b}_{M_G}z^{-M_G}}{\hat{a}_{0}+\hat{a}_{1}z^{-1}+...+\hat{a}_{N_G}z^{-N_G}} : N_G \geq M_G .
\end{equation}
%
Due to the nature of the methods we use for the stability check, we require
that the parametric errors in the plant have the same polynomial order as
the plant itself (all other errors described in this paper are ensured to fulfil this property in our
implementation).

In order to simplify the presentation, we employ a notation based on the
coefficients of the polynomials.  Let $\mathcal{P}^{N}$ be a polynomial
domain of order $N$. Let $P \in \mathcal{P}^{M,N}$ be a rational polynomial
$\frac{P_n}{P_d}$, where $P_n \in \mathcal{P}^{M}$ and $P_d \in \mathcal{P}^{N}$.  
For a vector of coefficients
%
\begin{equation}
\vec{P} \in \mathbb{R}^{N+M+2}=[n_{0}\ n_{1}\ \hdots \ n_{M}\ d_{0}\ d_{1}\ \hdots\ d_{N}\ ]^T
\label{eq:coefficients}
\end{equation}
%
and an uncertainty vector 
%
\begin{equation}
\Delta{\vec{P}}=[\Delta{n}_{0}\ \Delta{n}_{1}\ \hdots \ \Delta{n}_{M}\ \Delta{d}_{0}\ \Delta{d}_{1}\ \hdots\ \Delta{d}_{N}\ ]^T \; 
\label{eq:delta_coefficients}
\end{equation}
%
we write $\vec{\hat{G}}=\vec{G}+\Delta{\vec{G}}$.   
In the following we will either represent the functions $G(z)$, $C(z)$, 
or work over their respective coefficients $\vec{G}, \vec{C}$ in vector form.

%\begin{equation}
%\Delta{\vec{p}}=\begin{bmatrix}
%\hat{b_{0}} - b_{0}\\
%\hat{b_{1}} - b_{1}\\
%\vdots \\
%\hat{b}_{M_G} - b_{M_{G}}\\
%\hat{a_{0}} - a_{0}\\
%\hat{a_{1}} - a_{1}\\
%\vdots\\
%\hat{a}_{N_G} - a_{N_{G}}\\
%\end{bmatrix}=\begin{bmatrix}
%\vec{\hat{b}}\\
%\vec{\hat{a}}
%\end{bmatrix}-\begin{bmatrix}
%\vec{b}\\
%\vec{a}
%\end{bmatrix}
%\label{eq:uncertainty}
%\end{equation}

A typical digital control system system with a continuous plant and a discrete
controller is illustrated in Figure~\ref{fig:sampledsystem}.  The DAC and
ADC converters introduce quantisation errors which are modeled as noises $\nu_{1}(z)$ and $\nu_{2}(z)$,
$G(s)$ is the continuous-time plant model with parametric additive
uncertainty $\Delta_p{G}_n(s)$ and $\Delta_p{G}_d(s)$, $R(z)$ is the reference signal, $U(z)$ is
the control signal, and $\hat{Y}(z)$ is the output signal affected by the
noises and uncertainties in the closed-loop system.

\begin{figure*}[htb]
\centering

\tikzset{add/.style n args={4}{
    minimum width=6mm,
    path picture={
        \draw[circle] 
            (path picture bounding box.south east) -- (path picture bounding box.north west)
            (path picture bounding box.south west) -- (path picture bounding box.north east);
        \node[draw=none] at ($(path picture bounding box.south)+(0,0.13)$)     {\small #1};
        \node[draw=none] at ($(path picture bounding box.west)+(0.13,0)$)      {\small #2};
        \node[draw=none] at ($(path picture bounding box.north)+(0,-0.13)$)    {\small #3};
        \node[draw=none] at ($(path picture bounding box.east)+(-0.13,0)$)     {\small #4};
        }
    }
 }

\resizebox{1.0\textwidth}{!}{
 \begin{tikzpicture}[scale=0.6,-,>=stealth',shorten >=.2pt,auto,
     semithick, initial text=, ampersand replacement=\&,]

  \matrix[nodes={draw, fill=none, shape=rectangle, minimum height=.2cm, minimum width=.2cm, align=center}, row sep=.6cm, column sep=.6cm] {
    \node[draw=none] (r) {$R(z)$};
%   \& \coordinate (aux0);
   \& \node[circle,add={-}{+}{}{}] (circle) {};
   \node[draw=none] (ez) at ([xshift=1cm,yshift=.15cm]circle)  {$e(z)$};
   \& \node[rectangle,draw,
	minimum width=1cm,
	minimum height=1cm,
        label=\textbf{Controller}] (cz) {{\sc C(z)}};
   \node[draw=none] (ud) at ([xshift=1cm,yshift=.15cm]cz)  {$U(z)$};

   \& complexnode/.pic={ 
      \node[rectangle,draw,
	minimum width=3cm,
	minimum height=1.6cm,
	label=\textbf{DAC},] (dac) {};
     \node[circle,add={}{+}{+}{},fill=gray!20] (q2) at ([xshift=-.65cm]dac.center) {};
     \node[draw=none] (q2t)  at ([xshift=-.65cm,yshift=-.65cm]dac.center) {{\sc Q2}};
     \node[draw=none] (v2)  at ([xshift=-.65cm,yshift=1.5cm]dac.center) {$\nu_2(z)$};
     \node[fill=gray!20] (zoh) at ([xshift=.65cm]dac.center) {{\sc ZOH}};}   
   \& \node[rectangle,draw,
	minimum width=1cm,
	minimum height=1cm,
        label=\textbf{Plant}] (gs) {{\sc $\hat{G}(s)$}};
   \node[draw=none] (ud) at ([xshift=-2cm,yshift=.15cm]gs)  {$U(s)$};
   \node[draw=none] (y) at ([xshift=2cm,yshift=.15cm]gs)  {$Y(s)$};
   \& complexnode/.pic={ 
     \node[rectangle,draw,
	minimum width=4cm,
	minimum height=1.6cm,
	label=\textbf{ADC},] (adc) {};
   \draw[] ([xshift=-1cm]adc.center) -- ++(0.5,0.2cm);
   \coordinate (switch1) at ([xshift=-1cm]adc.center);
   \coordinate (switch2) at ([xshift=-0.4cm]adc.center);
   \node[circle,add={}{+}{+}{},fill=gray!20] (q1) at ([xshift=1cm]adc.center) {};} 
     \node[draw=none] (q2t)  at ([xshift=1cm,yshift=-.65cm]adc.center) {{\sc Q1}};
   \node[draw=none] (v1)  at ([xshift=1cm,yshift=1.5cm]adc.center) {$\nu_1(z)$};
   \& \coordinate (aux1);
   \& \node[draw=none] (yz) {$\hat{Y}(z)$};\\
   \& \coordinate (aux3); 
   \&
   \&
   \& 
   \& 
   \& \coordinate (aux2);\\
  };

  \path[->] (v1) edge (q1.north);
  \path[->] (v2) edge (q2.north);
  \path[->] (r) edge (circle.west);
  \path[->] (aux1) edge (yz);
  \path  
   (circle.east) edge (cz)
   (cz.east) edge (q2.west)
   (q2.east) edge (zoh.west)
   (zoh.east) edge (gs.west)
   (switch2) edge (q1.west)
   (q1.east) edge (aux1.west)
   (gs.east) edge (switch1.west)
   (aux1.south) edge (aux2.north)
   (aux2.west) edge (aux3.east); 
  \path[->]  (aux3.north) edge (circle.south);
 \end{tikzpicture}
}
 \caption{Closed-loop digital control system\label{fig:sampledsystem}}
\end{figure*}

The ADC and DAC may be abstracted by transforming the closed-loop system in
Figure~\ref{fig:sampledsystem} into the digital system in 
Figure~\ref{fig:digsystem1}, where the parametric uncertainties are
represented by~$\Delta_p \vec{G}$
%
%=\mathcal{Z}\left\lbrace{G_{P}(s)}\right\rbrace$
%
and the entire effect of $\nu_{1}$ and $\nu_{2}$ in the output $Y(z)$ is
represented as an additive output noise~$\nu(z)$. 
%, thus
%$\hat{Y}(z)=Y(z)+\nu(z)$.  The measurement noise $\nu(k)$ is enough for
%represent both quantisation noises ($\nu_{1}$ and $\nu_{2}$) what can be
%obtained reorganizing the equations~\ref{eq:outputfunctions}:
%%
%\begin{equation}
%\label{eq:outputfunctions2}
%\hat{Y}(z)=H_{1}R(z)+H_{1}\nu_{1}(z)+H_{2}(z)\nu_{2}(z),
%\end{equation}
%%
%where $H_{1}R(z)$ is exactly equal to $Y(z)$ without quantisation noise effects, thus:
%%
%\begin{equation}
%\nu(z)=H_{1}\nu_{1}(z)+H_{2}(z)\nu_{2}(z).
%\end{equation}

%The system in equation~\ref{controller_plant_tf} (disregarding the $\Delta{G}_{P}(z)$ may be represented by the following difference equation model:
%\begin{equation}
%\begin{split}
%y(k)=-a_{1}y(k-1)-a_{2}y(k-2)-\cdots - a_{N}y(k-N)\\
%+b_{0}u(k)+b_{1}u(k-1)+b_{2}u(k-2)+\cdots+b_{M}u(k-M),
%\end{split}
%\end{equation}
%and consequently $\hat{y}(k)$ may be represented by the following difference equation:
%\begin{equation}
%\label{eq:differencewithnoise}
%\begin{split}
%\hat{y}(k)=-a_{1}y(k-1)-a_{2}y(k-2)-\cdots - a_{N_{G}}y(k-N_{G})\\
%+b_{0}u(k)+b_{1}u(k-1)+b_{2}u(k-2)+\cdots+b_{M_{G}}u(k-M_{G})+\nu(k).
%\end{split}
%\end{equation}
%
%The z-transform of~\eqref{eq:differencewithnoise} is:
%\begin{equation}
%\begin{split}
%\hat{Y}(z)=(-a_{1}z^{-1}-a_{2}z^{-2}-\cdots-a_{N_{G}}z^{-N_{G}})Y(z)\\
%+(b_{0}+b_{1}z^{-1}+b_{2}z^{-2}+\cdots+b_{M_{G}}z^{-M_{G}})U(z)+\nu(z),
%\end{split}
%\end{equation}
%dividing both sides by $U(z)$, it is equivalent to:
%\begin{equation}
%\frac{\hat{Y}(z)}{U(z)}=[1-D_{G}(z)]G(z)+U(z)D_{G}(z)+\frac{\nu(z)}{U(z)}=G(z)+\Delta{G}_{Q}(z),
%\end{equation}

%Considering the quantisation noise $\nu(k)$ as an output error in discrete plant model, we may represent it as an additive noise in the plant output $y(k)$ resulting in the noisy output $\hat{y}(k)$:
%\begin{equation}
%
%\end{equation}

\begin{figure*}[ht]
\centering

\tikzset{add/.style n args={4}{
    minimum width=6mm,
    path picture={
        \draw[circle] 
            (path picture bounding box.south east) -- (path picture bounding box.north west)
            (path picture bounding box.south west) -- (path picture bounding box.north east);
        \node[draw=none] at ($(path picture bounding box.south)+(0,0.13)$)     {\small #1};
        \node[draw=none] at ($(path picture bounding box.west)+(0.13,0)$)      {\small #2};
        \node[draw=none] at ($(path picture bounding box.north)+(0,-0.13)$)    {\small #3};
        \node[draw=none] at ($(path picture bounding box.east)+(-0.13,0)$)     {\small #4};
        }
    }
 }

\resizebox{.6\textwidth}{!}{
 \begin{tikzpicture}[scale=0.3,-,>=stealth',shorten >=.2pt,auto,
     semithick, initial text=, ampersand replacement=\&,]

  \matrix[nodes={draw, fill=none, shape=rectangle, minimum height=.2cm, minimum width=.2cm, align=center}, row sep=.6cm, column sep=.3cm] {
    \node[draw=none] (r) {$R(z)$};
%   \& \coordinate (aux0);
   \& \node[circle,add={-}{+}{}{}] (circle) {};
   \node[draw=none] (ez) at ([xshift=1cm,yshift=.15cm]circle)  {$e(z)$};
   \& \node[rectangle,draw,
	minimum width=1cm,
	minimum height=1cm,
        label=\textbf{Controller}] (cz) {{\sc C(z)}};
   \node[draw=none] (ud) at ([xshift=1cm,yshift=.15cm]cz)  {$U(z)$};
   \& \node[rectangle,draw,
	minimum width=1cm,
	minimum height=1cm,
        label=\textbf{Plant}] (gs) {{\sc $\hat{G}(z)$}};
   \node[draw=none] (y) at ([xshift=2cm,yshift=.15cm]gs)  {$Y(z)$};

   \& \node[circle,add={}{+}{+}{}] (circle2) {};
%   \node[draw=none] (tyz) at ([xshift=-1cm,yshift=.15cm]circle2)  {$Y(z)$};
   \node[draw=none] (nu) at ([yshift=1cm]circle2)  {$\nu(z)$};
   \& \coordinate (aux1);
   \& \node[draw=none] (yz) {$\tilde Y(z)$};\\
   \& \coordinate (aux3); 
   \&
   \& 
   \& 
   \& \coordinate (aux2);\\
  };

%  \path[->] (v1) edge (q1.north);
%  \path[->] (v2) edge (q2.north);
  \path[->] (r) edge (circle.west);
  \path[->] (nu) edge (circle2.north);
  \path[->] (aux1) edge (yz);
  \path  
   (circle.east) edge (cz)
   (cz.east) edge (gs.west)
   (gs.east) edge (circle2.west)
   (circle2.east) edge (aux1.west)
   (aux1.south) edge (aux2.north)
   (aux2.west) edge (aux3.east); 
  \path[->]  (aux3.north) edge (circle.south);
 \end{tikzpicture}
}
\caption{System equivalent to the system in Figure~\ref{fig:sampledsystem} 
\label{fig:digsystem1}}
\end{figure*}

%
In Figure~\ref{fig:digsystem1}, two sources of uncertainties are
illustrated: parametric uncertainties related to modelling errors, and
uncertainties introduced by quantisations in sampling and holding process,
i.e., ADC and DAC conversions.  Assuming linearity and independence between
these sources, the uncertain model may be rewritten as a vector of coefficients in the z-domain using equations \ref{eq:coefficients} and \ref{eq:delta_coefficients}
%
\begin{equation}
\label{eq:complete_unc_model}
\vec{\hat{G}}=\vec{G}+\Delta_p \vec{G}+\Delta_q \vec{G},
\end{equation}
%
where $\Delta_p \vec{G}$ is the term that corresponds to the parametric
uncertainties and $\Delta_q \vec{G}$ is the term that reflects the
quantisation effects on the plant model.

%\red{[next figure is a bit unnecessary, does not ad much to the previous one.]} Figure~\ref{fig:digsystem2} shows the digital closed loop systems with $\Delta G_{P}(z)$ and $\Delta G_{Q}(z)$.

%\begin{figure}[ht]
%\centering
%
%\tikzset{add/.style n args={4}{
%    minimum width=6mm,
%    path picture={
%        \draw[circle] 
%            (path picture bounding box.south east) -- (path picture bounding box.north west)
%            (path picture bounding box.south west) -- (path picture bounding box.north east);
%        \node[draw=none] at ($(path picture bounding box.south)+(0,0.13)$)     {\small #1};
%        \node[draw=none] at ($(path picture bounding box.west)+(0.13,0)$)      {\small #2};
%        \node[draw=none] at ($(path picture bounding box.north)+(0,-0.13)$)    {\small #3};
%        \node[draw=none] at ($(path picture bounding box.east)+(-0.13,0)$)     {\small #4};
%        }
%    }
% }
%
%\resizebox{.5\textwidth}{!}{
% \begin{tikzpicture}[scale=0.3,-,>=stealth',shorten >=.2pt,auto,
%     semithick, initial text=, ampersand replacement=\&,]
%
%  \matrix[nodes={draw, fill=none, shape=rectangle, minimum height=.2cm, minimum width=.2cm, align=center}, row sep=.6cm, column sep=.3cm] {
%  %
%	\& 
%	\&
%	\& \coordinate (aux7); 
%	\& \node[rectangle,draw,
%	minimum width=1cm,
%	minimum height=1cm] (deltag2) {{\sc $\Delta G_{Q}(z)$}};
%	\& \coordinate (aux8);\\
%    \node[draw=none] (r) {$R(z)$};
%%   \& \coordinate (aux0);
%   \& \node[circle,add={-}{+}{}{}] (circle) {};
%   \node[draw=none] (ez) at ([xshift=1cm,yshift=.15cm]circle)  {$e(z)$};
%   \& \node[rectangle,draw,
%	minimum width=1cm,
%	minimum height=1cm,
%        label=\textbf{Controller}] (cz) {{\sc C(z)}};
%   \node[draw=none] (ud) at ([xshift=1cm,yshift=.15cm]cz)  {$U(z)$};
%   \& \coordinate (aux4);
%   \& \node[rectangle,draw,
%	minimum width=1cm,
%	minimum height=1cm,
%        label=\textbf{Plant}] (gs) {{\sc G(z)}};
%   \node[draw=none] (y) at ([xshift=1cm,yshift=.15cm]gs)  {$Y(z)$};
%   \& \node[circle,add={+}{+}{+}{}] (circle2) {};
%%   \node[draw=none] (tyz) at ([xshift=-1cm,yshift=.15cm]circle2)  {$Y(z)$};
%%   \node[draw=none] (nu) at ([yshift=1cm]circle2)  {$\nu(z)$};
%   \& \coordinate (aux1);
%   \& \node[draw=none] (yz) {$\hat{Y}(z)$};\\
%   \& 
%   \&
%   \& \coordinate (aux5); 
%   \& \node[rectangle,draw,
%	minimum width=1cm,
%	minimum height=1cm] (deltag) {{\sc $\Delta G_{P}(z)$}};
%   \& \coordinate (aux6);\\
%   \& \coordinate (aux3); 
%   \&
%   \& 
%   \& 
%   \&
%   \& \coordinate (aux2);\\
%  };
%
%
%%  \path[->] (v1) edge (q1.north);
%%  \path[->] (v2) edge (q2.north);
%  \path[->] (r) edge (circle.west);
%%  \path[->] (nu) edge (circle2.north);
%  \path[->] (aux1) edge (yz);
%  \path  
%   (aux4.north) edge (aux5.south)
%   (aux4.north) edge (aux7.south)
%   (aux5.east) edge (deltag.west)
%   (aux7.east) edge (deltag2.west)
%   (deltag.east) edge (aux6.west)
%   (deltag2.east) edge (aux8.west)
%   (circle.east) edge (cz)
%   (cz.east) edge (gs.west)
%   (circle2.east) edge (aux1.west)
%   (aux1.south) edge (aux2.north)
%   (aux2.west) edge (aux3.east); 
%  \path[->] (aux8.south) edge (circle2.north); 
%  \path[->] (aux6.south) edge (circle2.south);
%  \path[->] (gs.east) edge (circle2.west);
%  \path[->] (aux3.north) edge (circle.south);
% \end{tikzpicture}
%}
% \caption{Digital system with additive uncertainties $\Delta G_{P}(z)$ and $\Delta G_{Q}(z)$. \label{fig:digsystem2}}
%\end{figure}

%\textcolor{blue}{[IB: I propose to exchange the following paragraphs by the previous blue]}

%\textcolor{green}{DC: I believe Alessandro wanted to be more specific about the nature of the modelling error which is why I was trying to explain it. In this case I believe there are also two separate sources for $\Delta{G}_p(z)$. The first is $\Delta{G}_p(s)$ (which is basically a known tolerance that may come from G(t)), whilst the second is the error introduced by the calculations in the $\mathcal{Z}$ transform. Alessandro, do you want to address this second term (possibly as explained below) or just leave the blue section as is?}

%Since $G(z)$ is an approximation of $G(s)$ we introduce an error term caused during its calculation as $\Delta{G}_P$
%
%%
%\begin{align}
%\small
%\label{parametric_plant_tf}
%\hat{G}(z)&=\frac{(b_{0}+\Delta_p b_{0}) +...+(b_{M_G}+\Delta_p b_{M_G})z^{-M_G}}{(a_{0}+\Delta_p a_{0})+...+(a_{N_G}+\Delta_p a_{N_G})z^{-N_G}} \nonumber \\
%&=G(z)+\Delta{G}_P(z).
%\end{align}
%%
%Note that $\Delta{G}_P$ is undefined and upperbound by a user defined
%value such that $\hat{G}(z)$ corresponds to the set of
%possible polynomial forms of $G(z, T)$ that overapproximate $G(s)$.
%This calculation is performed outside of the scope of this work since it depends on the nature of the tool used to obtain the $\mathcal{Z}$ form. The error comes from the bounded horizon and mathematical precision used to calculate the $\mathcal{Z}$ transform, which will be tool and user dependent.
%\red{[The previous statement is confusing and I must disagree here: I want hard error bounds on a specific value of T.]}
%\green{[You are correct. It was absolutely wrong. Hopefully this explanation is clear. We use MATLAB, but still don't know the details of their implementation, so I don't have hard data on how this error is calculated.}

Since the controller is implemented using a FWL, $C(z)$ also suffers
disturbances from the FWL effects, with roundoffs in coefficients that may
change the closed-loop poles and zeros position, and consequently affect its
stability, as argued in~\cite{Bessa16}.

Let $\hat{C}(z)$ be the transfer function of the digital controller
represented using this FWL with integer size $I$ and fractional size~$F$. 
We shall denote the FWL domain of coefficients as $\mathbb{R}\langle I,F
\rangle$
%
%Since the controller is implemented in the $\mathbb{R}\langle I,F \rangle$ 
%domain, {\it i.e.} a Fixed-Point Finite Word Length with integer size $I$ and fractional
%size $F$, we shall denote this domain as $\mathbb{R}\langle I,F \rangle$ 
%
and assume that there exists a function
$\mathit{FWL}_n[\cdot]:\mathcal{P}^{n}\rightarrow \mathcal{P}^{n}\langle I,F \rangle$,
which applies the FWL effects to a polynomial, with an equivalent $\mathit{FWL}_{n,m}[\cdot]:\mathcal{P}^{n,m}\rightarrow \mathcal{P}^{n,m}\langle I,F \rangle$ that applies the same effect to a ratio of polynomials, where
$\mathcal{P}^{n}$ is the
space of polynomials of $n$-th order, $\mathcal{P}^{n}\langle I,F
\rangle$ is is the space of polynomials with coefficients in
$\mathbb{R}\langle I,F \rangle$, and $\mathcal{P}^{n,m}$,  $\mathcal{P}^{n,m}\langle I,F
\rangle$ are their rational equivalent.

Thus, the perturbed controller model $\hat{C}(z)$ may be obtained from the
original model $C(z)=\frac{C_{n}(z)}{C_{d}(z)}$ as follows:
%
\begin{equation}
\hat{C}(z)=\mathit{FWL}_{M_C,N_C}[C(z)]=\frac{\mathit{FWL}_{M_C}[C_n(z)]}{\mathit{FWL}_{N_C}[C_d(z)]}.
\end{equation}
%
In the case of a synthesised controller, because the synthesis is performed
using FWL, $\hat{C}(z) \equiv C(z)$.

%\textcolor{blue}{[IB: And I propose to exchange the remaining paragraphs by the previous blue]}
%
%\textcolor{green}{[DC: I changed it to a separate section to differentiate the effects, so we should keep both]}

%------------------------------------
\subsection{Stability under FWL and Noise}
\label{sec:stability}
%------------------------------------

Correct synthesis of the digital controller requires consideration of the
effect of FWL in controller and quantisation noises in the closed loop
system.  Let the quantiser $Q1$ be the source of a white noise $\nu_{1}$ and
$Q2$ be the source of a white noise $\nu_{2}$.  The following equation
models the system in Figure~\ref{fig:sampledsystem}, encompassing the
parametric uncertainties $\Delta \vec{G}$ and the FWL effects on
the controller ($\hat{C}(z)$):
%
\begin{equation}
\hat{Y}(z)=\hat{C}(z)\hat{G}(z)[R(z)+\nu_{1}(z)]+\nu_{2}(z)\hat{G}(z)-\hat{Y}(z)\hat{C}(z)\hat{G}(z),
\end{equation}
%
The above can be rewritten as follows:
%
\begin{equation}
\label{eq:outputfunctions}
\hat{Y}(z)=H_{1}[R(z)+\nu_{1}(z)]+H_{2}(z)\nu_{2}(z),
\end{equation}
%
where
%
\begin{equation}
H_{1}=\frac{\hat{C}(z)\hat{G}(z)}{1+\hat{C}(z)\hat{G}(z)},
\end{equation}
%
and
%
\begin{equation}
H_{2}=\frac{\hat{G}(z)}{1+\hat{C}(z)\hat{G}(z)}.
\end{equation}

\begin{myassumption}
\label{whitenoise}
%
The quantisation noises $\nu_{1}$ (from $Q1$) and $\nu_{2}$ (from $Q2$) are
uncorrelated white noises and their amplitudes are always bounded by the
half of quantisation step~\cite{astrom1997computer}, i.e., $\vert \nu_{1}
\vert \leq \frac{q_{1}}{2}$ and $\vert \nu_{2} \vert \leq \frac{q_{2}}{2}$,
where $q_{1}$ and $q_{2}$, respectively, are the quantisation steps of ADC
and DAC.
% 
\end{myassumption}

%\begin{mylemma}
%\label{lemma:variance}
%%
%In particular, Widrow \cite{widrow1956} models the quantisation noise as a
%gaussian model in which the second moment (variance) of the probability
%density function (PDF) of the quantisation noise is given by
%%
%\begin{equation}
%\label{eq:variancelemma}
%\sigma_{Q}^{2}=\frac{q^{2}}{12},
%\end{equation}
%where $q$ is the step size of quantiser. (Proof is on \cite{widrow1956})
%\end{mylemma}
%
%\begin{mylemma}
%\label{lemma:boundednoise}
%%
%The power spectral density of the quantisation noise (supposed white)
%$S_{\nu}(z)$ is bounded and may be computed through the z-transform of
%autocorrelation function of $\nu(k)$, thus:
%%
%\begin{equation}
%S_{\nu}(z)= \sigma_{Q}^{2}=\frac{q^{2}}{12}.
%\end{equation}
%\end{mylemma}
%
%\begin{proof}
%%
%Considering $\nu(k)$ a random process, where $\nu(k)$ is independent from
%$\nu(k+i)$ for every integer $k\neq 0$.  The autocorrelation function of
%$\nu(k)$, {\it i.e.} $R_{\nu\nu}(\tau)$ may be computed as follows:
%%
%\begin{equation}
%R_{\nu\nu}(\tau)=E[\nu(t+\tau)\nu(t)]=\sigma_{Q}^{2}\cdot \delta(\tau),
%\end{equation}
%%
%where $\delta(\tau)$ is the Dirac delta function at delay time $\tau$.
%
%By definition~\cite{poularikas2000transforms}, the power spectral density of
%$\nu(k)$ is the z-transform of $R_{\nu\nu}(\tau)$, thus
%%
%\begin{equation}
%S_{\nu}(z)=  \sum\limits_{k=-\infty}^{\infty}{R_{\nu\nu}(k)z^{-k}} = \sigma_{Q}^{2},
%\end{equation}
%considering the Lemma~\ref{lemma:variance}:
%\begin{equation}
%S_{\nu}(z)=  \frac{q^{2}}{12}.
%\end{equation}
%\end{proof}
%
%Finally, the quantisation noise spectrum may be considered a energy-bounded
%(Lemma~\ref{lemma:boundednoise}), and consequently the Parseval's theorem
%ensures that the quantisation noise is also energy-bounded in time-domain. 
%Thus $\hat{Y}(z)$ will be bounded and the closed-loop system stable if the
%conditions in Lemma~\ref{lemma:eq_int_stab} are followed.

%\red{[Add remark on bound (q/2) for signal. We need a bounded signal to be able to leverage BIBO stability results. ]}

It is known that a discrete-time linear system is said to be Bounded-Input and Bounded-Output
(BIBO) stable if and only if every pole of its transfer function lies inside
the unit circle \cite{Astrom08}.  Analyzing Eq.~\eqref{eq:outputfunctions}, the following
proposition states necessary conditions for BIBO stability of the closed-loop system
in Figure~\ref{fig:sampledsystem}, since the exogenous signals $R(z)$,
$\nu_{1}$, and $\nu_{2}$ are bounded.

\begin{myprop}{\cite{Bessa16,fadali}}
\label{prop:eq_int_stab} 
%
A feedback closed-loop control system as given in Figure~\ref{fig:sampledsystem},
with a FWL implementation of the digital controller $\hat{C}(z) =\mathit{FWL}_{M_C,N_C}[C(z)]$ and uncertain
discrete model of the plant 
$$\hat{G}(z) =\frac{\hat{G}_n(z)}{\hat{G}_d(z)}: \vec{\hat{G}}=\vec{G}+\Delta_p \vec{G}+\Delta_q
\vec{G} \text{(from  \ref{eq:coefficients},\ref{eq:delta_coefficients})}$$ ,  
$\hat{G}(z)$, is
BIBO-stable if and only if:
%
\begin{itemize}
\item  the roots of characteristic polynomial $S(z)$ are inside the open unit circle, where $S(z)$ is:
\begin{equation}
\label{eq:internal_stab_lemma}
S(z)=\mathit{FWL}_{M_C}[C_n(z)]\hat{G}_n(z)+\mathit{FWL}_{N_C}[C_d(z)]\hat{G}_d(z);
\end{equation}
\item the direct loop product $\mathit{FWL}_{M_C,N_C}[C(z)]\cdot G(z)$ has no pole-zero cancellation on or outside the unit circle.
\end{itemize}
\end{myprop}

Proposition~\ref{prop:eq_int_stab} provides the necessary conditions for the digital controller to stabilize the closed loop system, 
considering plant parametric uncertainties (i.e., $\Delta_p \vec{G}$),
quantisation noises ($\nu_{1}$ and $\nu_{2}$) and FWL effects in the control
software.  
Notice that the model for plant uncertainty due to quantisation noise (i.e., $\Delta_q \vec{G}$) may be disregarded if the conditions on
Proposition~\ref{prop:eq_int_stab} are satisfied and if the quantisation noise is
bounded.  The boundedness of the quantisation noise comes from Assumption~\ref{whitenoise}.

%\red{[first: lemmae precede theorems. ]}

% Let $\tilde C(z)$ and $\tilde G(z)$ be the transfer functions represented
% using Fixed-Point bit vectors with integer size $I_b$ and fractional
% size $F_b$, whose domain is $\mathbb{R}\langle I_b,F_b \rangle \supseteq \mathbb{R}\langle I,F \rangle$.
% %
% \begin{align}
% \small
% \label{digital_controller_plant_tf}
% \tilde C(z)&=\frac{\tilde \beta_{0}+\tilde \beta_{1}z^{-1}+...+\tilde \beta_{M_C}z^{-M_C}}{\tilde \alpha_{0}+\tilde \alpha_{1}z^{-1}+...+\tilde \alpha_{N_C}z^{-N_C}}, \\
% %\label{plant_tf}
% \tilde G(z)&=\frac{\tilde b_{0}+\tilde b_{1}z^{-1}+...+\tilde b_{M_G}z^{-M_G}}{\tilde a_{0}+\tilde a_{1}z^{-1}+...+\tilde a_{N_G}z^{-N_G}}.
% \end{align}

% Since the controller is synthesized in the $\mathbb{R}\langle I,F \rangle$
% domain,
% %
%  $$\forall i \leq N_C, j \leq M_C\ \  \tilde \beta_{i} \equiv \beta_{i} \wedge \tilde \alpha_{j} \equiv \alpha_{j} \Leftrightarrow \tilde C(z) \equiv C(z)$$

% Two approaches have been taken to synthesise a stable controller based on this model.
% The first uses interval arithmetic, representing the coefficients $[{c}_i-\Delta_p{c}_i-\Delta_b{c}_i\ ,\ {c}_i+\Delta_p{c}_i-\Delta_b{c}_i+(2^{-F_b})]$

% \cdsay{Dario, what coefficients are these?}

% and rounding operations outwards during the process. This results in a tighter calculation of the error at the cost of performance.

%, thus supporting this assumption.

%------------------------------------
%\subsection{Using Fixed Point Arithmetic for verification}
%\label{fixedpoint_verification}
%------------------------------------

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Automated Program Synthesis \\ for Digital
Control}\label{sec:synthesis}

%------------------------------------
\subsection{Overview of the Synthesis Process}
\label{verification-flow}
%------------------------------------

In order to synthesise closed-loop digital control systems, we use a program
synthesis engine.  Our program synthesiser implements Counter-Example Guided
Inductive Synthesis (CEGIS)~\cite{sketch}.  We start by presenting its
general architecture followed by describing the parts specific to
closed-loop control systems.  A high-level view of the synthesis process is
given in Figure~\ref{DSSynth_process}.  Steps $1$ to $3$ are performed by
the user and Steps~A to D are automatically performed by our tool for
Digital Systems Synthesis, named \tool.
%
%, which accepts the
%digital controller transfer functions together with the plant model.  For
%any digital controller, implementation details should be provided ({\it
%e.g.}, number of bits, realization, and sample time).

\begin{figure*}[t]
\centering
\includegraphics[width=\textwidth]{figures/verification-flow.png}
\vspace{1ex}
\caption{Overview of the synthesis process\label{DSSynth_process}}
\end{figure*}

%In Step $1$, the user provides inputs $G(z)$ and $\Delta{G}(z)$
%%$\%$ ({\it i.e.}, $\Delta{G}(z)$ as a percentage of $G(z)$)
%, which contain the plant model and the
%interval, respectively.  In Step $2$, a digital controller must be designed
%with any preferred method, where $C(z)$ is obtained.  The controller
%numerical representation and realization form are chosen in Steps $3$ and
%$4$, respectively.  In Step $5$, the user finally configures the
%verification parameters ({\it e.g.}, verification time, properties, and BMC
%tool).  After that, DSVerifier verification engine performs an (automatic)
%verification of the desired property $\phi$ ({\it i.e.}, stability).  The
%user steps described in Steps $1$--$5$ result in an ANSI-C code that should
%be used as an input for DSVerifier.

Given a model for a plant for which to synthesise a controller, provided in
an ANSI-C input file (Steps 1--3), \tool constructs a non-deterministic
model to represent the plant family (Step~A), and formulates a function
(Step~B) using implementation details provided in Steps~$2$ and $3$ to
calculate the parameters of the controller to be synthesised (Step~C). 
Finally, \tool builds an intermediate ANSI-C code for the digital system
implementation, which is used as input for the CEGIS engine (Step~D).

This intermediate ANSI-C code model contains a specification $\phi$ for the
property of interest (i.e., robust stability) and is passed to the
Counterexample-Guided Inductive Synthesis (CEGIS) module of
CBMC~\cite{ClarkeKL04}, where the controller is marked as the input variable
to synthesise.  The properties specified in the ANSI-C code model can only
be satisfied if CEGIS synthesises controller coefficients that
satisfy~$\phi$.  CEGIS employs an iterative, counterexample-guided
refinement process which is explained in detail in
section~\ref{synthesiser-general}.  A~successful synthesis result is
reported if CEGIS generates a controller that is safe with respect to~$\phi$. 
In particular, the ANSI-C code model guarantees that a synthesised solution
is complete and sound with respect to the stability property~$\phi$, since it
does not depend on system inputs and outputs.  In the case of stability, the
specification~$\phi$ consists in a number of assumptions on the polynomial
coefficients, following Jury's Criteria, as well as the restrictions set in
the representation of these coefficients as discussed in more detail in 
Section~\ref{synthesis-elements}.

% by the function $f$ (cf.~Eq.~\eqref{FWL_tf})
%In Step A, DSVerifier constructs a non-deterministic model to represent the
%plant family $\mathfrak{G}$ using $g_0$ and $\Delta{g}\%$, which are
%provided in Step $1$.  DSVerifier then formulates $\mathit{FWL}[\cdot]$ in
%Step B using implementation details provided from Steps $2$ and $3$, and
%then computes $\mathit{FWL}[c_0]$ in Step C.  Thus, DSVerifier builds an
%intermediate ANSI-C code for the digital system implementation, which is
%used as input for the model-checker, as pointed in Step D.

%This intermediate ANSI-C code model has three main modules: digital
%controller code to be embedded into the microprocessor; plant model code,
%which simulates the plant model dynamics considering uncertainties; and
%``assert'' and ``assume'' statements, which control the verification flow. 
%For the digital controller code, transfer functions coefficients are
%quantised and deterministic, and all operations use fixed-point arithmetic. 
%\textcolor{red}{In the plant model code, transfer function coefficients are
%not quantised, but represented with maximum precision based on float
%data-type variables}; they are treated as non-deterministic variables to
%support model uncertainties.  The directive ``assume'' bounds
%non-deterministic variables, {\it i.e.}, inputs and plant uncertain
%coefficients.
%
%The translation of ANSI-C code into SAT/SMT formulae
%is done by a back-end model-checking tool ({\it e.g.},
%CBMC~\cite{ClarkeKL04} or ESBMC~\cite{CordeiroFM12}).  Here, DSVerifier
%symbolically checks a given property $\phi$ with respect to closed-loop
%systems, which are composed by $f(c_i)$ and every $G$ in $\mathfrak{G}$. 
%If any property violation is found, then DSVerifier reports a
%counterexample, which contains system inputs or parametric deviations
%$\Delta{G}$ that lead the system to a failure.  A successful verification
%result is reported if the system is safe with respect to $\phi$.  In
%particular, the stability verification is complete and sound, since it does
%not depend on system outputs and inputs.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%\section{Modelling Errors}

%In this section, we show how to embed discretisation errors on the plant model, as follows: 
%%$\hat{G}(z)$ is split into two terms:
%%
%\begin{equation}
%\label{eq:uncertainplant}
%%\hat{G}(z)=G(z)+\Delta{G(z)},
%\tilde G(z) = \hat G(z) + \Delta G_P(z) = G(z,T) + \Delta G_Q(z) + \Delta G_P(z),   
%\end{equation}
%where $G(z,T)$ is the time discretisation of the nominal physical plant $G(s)$, 
%$\Delta{G_{Q}(z)}$ represents the quantisation uncertainty model, 
%and $\Delta{G_{P}(z)}$ describes the parametric uncertainty model.

%\noindent where $G(z)$ is the discretised nominal plant model and
%$\Delta{G(z)}$ is the additive uncertainty, which is related to parametric
%uncertainties~\cite{Bessa16}.

%\section{An Uncertainty Model for the quantisation Error} 
%\label{sec:uncertainty-model-quantisation-error}
%
%%Here, an uncertainty model to accommodate the
%%quantisation noise effects on $\hat{G}(z)$ is proposed.  
%%As a result, Eq.~\eqref{eq:uncertainplant} is rewritten as
%%
%%\begin{equation}
%%\label{eq:uncertainFWLplant}
%%\hat{G}(z)=G(z)+\Delta{G_{Q}(z)}+\Delta{G_{P}(z)},
%%\end{equation}
%
%A model for $\Delta{G_{Q}(z)}$ that can represent the quantisation
%noise effect in the plant output, and consequently the closed-loop system
%output, is proposed.  
%%\red{[eliminate next:] 
%%For convenience, the parametric uncertainties will be described here.}
%
%Consider the closed-loop sampled system in Figure~\ref{fig:sampledsystem},
%where $C(z)$ is the digital controller model, whose control output signal is
%quantised ($Q2$) and interpolated via a ZOH using digital-to-analog
%converter (DAC).  The quantisation occurs because the DAC resolution may be
%different from the fixed point representation of the controller.  $G(s)$ is
%the continuous-time model of the plant, from which output samples are taken
%and quantised ($Q1$) by means of an analog-to-digital converter (ADC)
%synchronized with the DAC.  Each quantiser ($Q1$ and $Q2$) will present some
%contribution for the total quantisation noise.
%    
%\begin{figure*}[ht]
%\centering
%
%\tikzset{add/.style n args={4}{
%    minimum width=6mm,
%    path picture={
%        \draw[circle] 
%            (path picture bounding box.south east) -- (path picture bounding box.north west)
%            (path picture bounding box.south west) -- (path picture bounding box.north east);
%        \node[draw=none] at ($(path picture bounding box.south)+(0,0.13)$)     {\small #1};
%        \node[draw=none] at ($(path picture bounding box.west)+(0.13,0)$)      {\small #2};
%        \node[draw=none] at ($(path picture bounding box.north)+(0,-0.13)$)    {\small #3};
%        \node[draw=none] at ($(path picture bounding box.east)+(-0.13,0)$)     {\small #4};
%        }
%    }
% }
%
%\resizebox{.8\textwidth}{!}{
% \begin{tikzpicture}[scale=0.6,-,>=stealth',shorten >=.2pt,auto,
%     semithick, initial text=, ampersand replacement=\&,]
%
%  \matrix[nodes={draw, fill=none, shape=rectangle, minimum height=.2cm, minimum width=.2cm, align=center}, row sep=.6cm, column sep=.3cm] {
%    \node[draw=none] (r) {$\tilde R(z)$};
%%   \& \coordinate (aux0);
%   \& \node[circle,add={-}{+}{}{}] (circle) {};
%   \node[draw=none] (ez) at ([xshift=1cm,yshift=.15cm]circle)  {$\tilde e(z)$};
%   \& \node[rectangle,draw,
%	minimum width=1cm,
%	minimum height=1cm,
%        label=\textbf{Controller}] (cz) {{\sc C(z)}};
%   \node[draw=none] (ud) at ([xshift=1cm,yshift=.15cm]cz)  {$\tilde U(z)$};
%     
%   \& complexnode/.pic={ 
%      \node[rectangle,draw,
%	minimum width=3cm,
%	minimum height=1.6cm,
%	label=\textbf{DAC},] (dac) {};
%     \node[circle,add={}{+}{+}{},fill=yellow!20] (q2) at ([xshift=-.65cm]dac.center) {};
%     \node[draw=none] (q2t)  at ([xshift=-.65cm,yshift=-.65cm]dac.center) {{\sc Q2}};
%     \node[draw=none] (v2)  at ([xshift=-.65cm,yshift=1.5cm]dac.center) {$\nu_2(z)$};
%     \node[fill=yellow!20] (zoh) at ([xshift=.65cm]dac.center) {{\sc ZOH}};}
%   \& \node[rectangle,draw,
%	minimum width=1cm,
%	minimum height=1cm,
%        label=\textbf{Plant}] (gs) {{\sc G(s)}};
%   \node[draw=none] (ud) at ([xshift=-1cm,yshift=.15cm]gs)  {$U(s)$};
%   \node[draw=none] (y) at ([xshift=1cm,yshift=.15cm]gs)  {$Y(s)$};
%   \& complexnode/.pic={ 
%     \node[rectangle,draw,
%	minimum width=4cm,
%	minimum height=1.6cm,
%	label=\textbf{ADC},] (adc) {};
%   \draw[] ([xshift=-1cm]adc.center) -- ++(0.5,0.2cm);
%   \coordinate (switch1) at ([xshift=-1cm]adc.center);
%   \coordinate (switch2) at ([xshift=-0.4cm]adc.center);
%   \node[circle,add={}{+}{+}{},fill=yellow!20] (q1) at ([xshift=1cm]adc.center) {};} 
%     \node[draw=none] (q2t)  at ([xshift=1cm,yshift=-.65cm]adc.center) {{\sc Q1}};
%   \node[draw=none] (v1)  at ([xshift=1cm,yshift=1.5cm]adc.center) {$\nu_1(z)$};
%   \& \coordinate (aux1);
%   \& \node[draw=none] (yz) {$\tilde Y(z)$};\\
%   \& \coordinate (aux3); 
%   \&
%   \&
%   \& 
%   \& 
%   \& \coordinate (aux2);\\
%  };
%
%
%  \path[->] (v1) edge (q1.north);
%  \path[->] (v2) edge (q2.north);
%  \path[->] (r) edge (circle.west);
%  \path[->] (aux1) edge (yz);
%  \path  
%   (circle.east) edge (cz)
%   (cz.east) edge (q2.west)
%   (q2.east) edge (zoh.west)
%   (zoh.east) edge (gs.west)
%   (switch2) edge (q1.west)
%   (q1.east) edge (aux1.west)
%   (gs.east) edge (switch1.west)
%   (aux1.south) edge (aux2.north)
%   (aux2.west) edge (aux3.east); 
%  \path[->]  (aux3.north) edge (circle.south);
% \end{tikzpicture}
%}
% \caption{Sampled System. \label{fig:sampledsystem}}
%\end{figure*}
%
%%% \begin{figure}[ht]
%%% \centering
%%% \includegraphics[width=\columnwidth]{figures/hsystembd.pdf}
%%% \caption{Sampled system.}
%%% \label{fig:sampledsystem}
%%% \end{figure}
%
%
%\begin{figure}[ht]
%\centering
%
%\tikzset{add/.style n args={4}{
%    minimum width=6mm,
%    path picture={
%        \draw[circle] 
%            (path picture bounding box.south east) -- (path picture bounding box.north west)
%            (path picture bounding box.south west) -- (path picture bounding box.north east);
%        \node[draw=none] at ($(path picture bounding box.south)+(0,0.13)$)     {\small #1};
%        \node[draw=none] at ($(path picture bounding box.west)+(0.13,0)$)      {\small #2};
%        \node[draw=none] at ($(path picture bounding box.north)+(0,-0.13)$)    {\small #3};
%        \node[draw=none] at ($(path picture bounding box.east)+(-0.13,0)$)     {\small #4};
%        }
%    }
% }
%
%\resizebox{.5\textwidth}{!}{
% \begin{tikzpicture}[scale=0.3,-,>=stealth',shorten >=.2pt,auto,
%     semithick, initial text=, ampersand replacement=\&,]
%
%  \matrix[nodes={draw, fill=none, shape=rectangle, minimum height=.2cm, minimum width=.2cm, align=center}, row sep=.6cm, column sep=.3cm] {
%    \node[draw=none] (r) {$R(z)$};
%%   \& \coordinate (aux0);
%   \& \node[circle,add={-}{+}{}{}] (circle) {};
%   \node[draw=none] (ez) at ([xshift=1cm,yshift=.15cm]circle)  {$e(z)$};
%   \& \node[rectangle,draw,
%	minimum width=1cm,
%	minimum height=1cm,
%        label=\textbf{Controller}] (cz) {{\sc C(z)}};
%   \node[draw=none] (ud) at ([xshift=1cm,yshift=.15cm]cz)  {$U(z)$};
%   \& \node[rectangle,draw,
%	minimum width=1cm,
%	minimum height=1cm,
%        label=\textbf{Plant}] (gs) {{\sc G(z)}};
%   \node[draw=none] (y) at ([xshift=1cm,yshift=.15cm]gs)  {$Y(z)$};
%
%   \& \node[circle,add={}{+}{+}{}] (circle2) {};
%%   \node[draw=none] (tyz) at ([xshift=-1cm,yshift=.15cm]circle2)  {$Y(z)$};
%   \node[draw=none] (nu) at ([yshift=1cm]circle2)  {$\nu(z)$};
%   \& \coordinate (aux1);
%   \& \node[draw=none] (yz) {$\tilde Y(z)$};\\
%   \& \coordinate (aux3); 
%   \&
%   \& 
%   \& 
%   \& \coordinate (aux2);\\
%  };
%
%
%%  \path[->] (v1) edge (q1.north);
%%  \path[->] (v2) edge (q2.north);
%  \path[->] (r) edge (circle.west);
%  \path[->] (nu) edge (circle2.north);
%  \path[->] (aux1) edge (yz);
%  \path  
%   (circle.east) edge (cz)
%   (cz.east) edge (gs.west)
%   (gs.east) edge (circle2.west)
%   (circle2.east) edge (aux1.west)
%   (aux1.south) edge (aux2.north)
%   (aux2.west) edge (aux3.east); 
%  \path[->]  (aux3.north) edge (circle.south);
% \end{tikzpicture}
%}
% \caption{Sampled System. \label{fig:sampledsystem}}
%\end{figure}
%
%%
%%
%
%\begin{figure}[ht]
%\centering
%
%\tikzset{add/.style n args={4}{
%    minimum width=6mm,
%    path picture={
%        \draw[circle] 
%            (path picture bounding box.south east) -- (path picture bounding box.north west)
%            (path picture bounding box.south west) -- (path picture bounding box.north east);
%        \node[draw=none] at ($(path picture bounding box.south)+(0,0.13)$)     {\small #1};
%        \node[draw=none] at ($(path picture bounding box.west)+(0.13,0)$)      {\small #2};
%        \node[draw=none] at ($(path picture bounding box.north)+(0,-0.13)$)    {\small #3};
%        \node[draw=none] at ($(path picture bounding box.east)+(-0.13,0)$)     {\small #4};
%        }
%    }
% }
%
%\resizebox{.5\textwidth}{!}{
% \begin{tikzpicture}[scale=0.3,-,>=stealth',shorten >=.2pt,auto,
%     semithick, initial text=, ampersand replacement=\&,]
%
%  \matrix[nodes={draw, fill=none, shape=rectangle, minimum height=.2cm, minimum width=.2cm, align=center}, row sep=.6cm, column sep=.3cm] {
%    \node[draw=none] (r) {$R(z)$};
%%   \& \coordinate (aux0);
%   \& \node[circle,add={-}{+}{}{}] (circle) {};
%   \node[draw=none] (ez) at ([xshift=1cm,yshift=.15cm]circle)  {$e(z)$};
%   \& \coordinate (aux4);
%   \& \node[rectangle,draw,
%	minimum width=1cm,
%	minimum height=1cm,
%        label=\textbf{Controller}] (cz) {{\sc C(z)}};
%   \node[draw=none] (ud) at ([xshift=1cm,yshift=.15cm]cz)  {$U(z)$};
%   \& \node[rectangle,draw,
%	minimum width=1cm,
%	minimum height=1cm,
%        label=\textbf{Plant}] (gs) {{\sc G(z)}};
%   \node[draw=none] (y) at ([xshift=1cm,yshift=.15cm]gs)  {$Y(z)$};
%   \& \node[circle,add={+}{+}{}{}] (circle2) {};
%%   \node[draw=none] (tyz) at ([xshift=-1cm,yshift=.15cm]circle2)  {$Y(z)$};
%%   \node[draw=none] (nu) at ([yshift=1cm]circle2)  {$\nu(z)$};
%   \& \coordinate (aux1);
%   \& \node[draw=none] (yz) {$\tilde Y(z)$};\\
%   \& 
%   \& \coordinate (aux5); 
%   \& \node[rectangle,draw,
%	minimum width=1cm,
%	minimum height=1cm] (deltag) {{\sc $\Delta G(z)$}};
%   \&
%   \& \coordinate (aux6);\\
%   \& \coordinate (aux3); 
%   \&
%   \& 
%   \& 
%   \&
%   \& \coordinate (aux2);\\
%  };
%
%
%%  \path[->] (v1) edge (q1.north);
%%  \path[->] (v2) edge (q2.north);
%  \path[->] (r) edge (circle.west);
%%  \path[->] (nu) edge (circle2.north);
%  \path[->] (aux1) edge (yz);
%  \path  
%   (aux4.north) edge (aux5.south)
%   (aux5.east) edge (deltag.west)
%   (deltag.east) edge (aux6.west)
%   (aux6.south) edge (circle2.south)
%   (circle.east) edge (cz)
%   (cz.east) edge (gs.west)
%   (gs.east) edge (circle2.west)
%   (circle2.east) edge (aux1.west)
%   (aux1.south) edge (aux2.north)
%   (aux2.west) edge (aux3.east); 
%  \path[->]  (aux3.north) edge (circle.south);
% \end{tikzpicture}
%}
% \caption{Sampled System. \label{fig:sampledsystem}}
%\end{figure}

%\begin{myprop}
%Considering the quantisation noise $\nu(k)$ as an output error in the
%discrete plant model, 
%%{\it i.e.}, it does not affect the plant dynamics, but its output samples, 
%$\Delta{G_{Q}(z)}$ can be obtained from
%%
%\begin{equation}
%\label{eq:quantisation_tf}
%\Delta{G_{Q}(z)}=\frac{\nu(z)}{U(z)},
%\end{equation}
%where $\nu(z)$ is the z-transform of the random process $\nu(k)$,
%which can be calculated through the power spectral density of
%$\nu(k)$~\cite{poularikas2000transforms}.  
%Thus, the $\Delta{G_{Q}(z)}$ can expressed as
%%
%\begin{equation}
%\label{eq:deltag_var}
%\Delta{G_{Q}(z)}=\frac{\sigma_{Q}^{2}}{U(z)},
%\end{equation}
%where $\sigma_{Q}^{2}$ is the variance of the quantisation noise and
%$U(z)$ is the z-transform of control signal ({\it i.e.}, the z-transform of
%the signal from the digital controller), which is provided to the plant by
%means of ZOH.
%\end{myprop}
%
%%
%\begin{proof}
%%
%Suppose that the plant discrete model in Figure~\ref{fig:sampledsystem} can
%be represented by the following pulse transfer function (without
%quantisation noise effect):
%%
%\begin{equation}
%\label{eq:tfwithout}
%G(z)=\frac{Y(z)}{U(z)}=\frac{B(z)}{A(z)}=\frac{b_{0}+b_{1}z^{-1}+b_{2}z^{-2}+\cdots+b_{M}z^{-M}}{1+a_{1}z^{-1}+a_{2}z^{-2}+\cdots+a_{N}z^{-N}},
%\end{equation}
%%
%\noindent where $Y(z)$ is the z-transform of the plant output signal without
%noise $y(k)$, $U(z)$ is the z-transform of the control signal, which is
%provided to the plant, $A(z)$ and $B(z)$ are the plant pulse transfer
%function numerator and denominator, respectively.
%
%Considering the quantisation noise $\nu(k)$ as an output error in the
%discrete plant model, an additive noise in the plant output $y(k)$ resulting
%in the noisy output $\hat{y}(k)$ can be represented as
%%
%\begin{equation}
%\hat{y}(k)=y(k)+\nu(k).
%\end{equation}
%
%Note that the model in Eq.~\eqref{eq:tfwithout} is equivalent to the
%following difference equation:
%%
%\begin{equation}
%\begin{split}
%y(k)=-a_{1}y(k-1)-a_{2}y(k-2)-\cdots - a_{N}y(k-N)\\
%+b_{0}u(k)+b_{1}u(k-1)+b_{2}u(k-2)+\cdots+b_{M}u(k-M),
%\end{split}
%\end{equation}
%
%\noindent and consequently $\hat{y}(k)$ can be represented by the following
%difference equation:
%%
%\begin{equation}
%\label{eq:differencewithnoise}
%\begin{split}
%\hat{y}(k)=-a_{1}y(k-1)-a_{2}y(k-2)-\cdots - a_{N}y(k-N)\\
%+b_{0}u(k)+b_{1}u(k-1)+b_{2}u(k-2)+\cdots+b_{M}u(k-M)+\nu(k).
%\end{split}
%\end{equation}
%
%The z-transform of Eq.~\eqref{eq:differencewithnoise} is
%%
%\begin{equation}
%\begin{split}
%\hat{Y}(z)=(-a_{1}z^{-1}-a_{2}z^{-2}-\cdots-a_{N}z^{-N})Y(z)\\
%+(b_{0}+b_{1}z^{-1}+b_{2}z^{-2}+\cdots+b_{M}z^{-M})U(z)+\nu(z),
%\end{split}
%\end{equation}
%dividing both sides by $U(z)$, it is equivalent to:
%\begin{equation}
%\frac{\hat{Y}(z)}{U(z)}=\hat{G}(z)=[1-A(z)]G(z)+U(z)B(z)+\frac{\nu(z)}{U(z)}.
%\end{equation}
%
%Considering $\hat{G}(z)=G(z)+\Delta{G_{Q}(z)}$ and Eq.~\eqref{eq:tfwithout},
%the intended expression can be obtained as
%%
%\begin{equation}
%\Delta{G_{Q}(z)}=\frac{\nu(z)}{U(z)} 
%\end{equation}
%\end{proof}

%Eq.~\eqref{eq:deltag_var} indicates that the modelling of the quantisation noise
%effect on the plant transfer function demands an estimation of the variance
%of quantisation noise $\sigma_{Q}^{2}$.  To obtain such estimation, the
%methodology presented by~\cite{widrow2008quantisation}
%is used.  The authors provide an algorithm for computing the mean square of
%the output noise due to $Q1$ and $Q2$.  Considering that each quantiser
%contribution can be represented by a white noise with zero mean
%(Assumption~\ref{whitenoise}), the mean square can be considered as the
%variance of noise ($\sigma_{Q}^{2}$), which can be calculated as
%%
%\begin{equation}
%\sigma_{Q}^{2}=\frac{\lambda_{1}^{2}\cdot \sum {h_{Q1}^{2}} \cdot q_{1}+\lambda_{2}^{2}\cdot \sum {h_{Q2}^{2}} \cdot q_{2}}{12},
%\end{equation}
%%
%\noindent where $\lambda_{1}$ and $\lambda_{2}$ are the scale factors, {\it i.e.} constant gains usually employed before the ADC and after the DAC in order to avoid the overload of converters  %\red{[define scale factors]} 
%for the quantisation $Q1$ and $Q2$, respectively; $q_{1}$ and $q_{2}$ are the quantisation step for $Q1$ and $Q2$, respectively, and $\sum {h_{Q1}^{2}}$ and $\sum
%{h_{Q2}^{2}}$ are the squares sums of the impulse response obtained by means
%of a transfer function from $Q1$ and $Q2$ to the plant output, respectively.
%
%The transfer functions $H_{Q1}(z)$ and $H_{Q2}(z)$, from which the squares
%sums of the impulse response are needed, can be obtained through the
%analysis of the system in Figure~\ref{fig:sampledsystem}.

%\red{[The following text needs to be completed. (Notice typos, reference to wrong signal $R$, and to wrong TF $C,G$.]}
%
%Considering that
%the quantiser $Q1$ is the source of a white noise with zero mean $\nu_{1}$
%and $Q2$ is the source of a white noise with zero mean $\nu_{2}$, the
%following equation can be written for the system in
%Figure~\ref{fig:sampledsystem}
%%
%\begin{equation}
%\resizebox{0.48\textwidth}{!}{
%$\tilde{Y}(z)=\tilde R(z)C(z)G(z)-\tilde{Y}(z)C(z)G(z)+\nu_{1}(z)+\nu_{2}(z)G(z),$
%}
%\end{equation}
%
%\noindent which can be rewritten as
%%
%\begin{equation}
%\label{eq:outputfunctions}
%\tilde{Y}(z)=C(z)G(z)H(z)R(z)+H(z)\nu_{1}(z)+G(z)H(z)\nu_{2}(z)
%\end{equation}
%$$H(z)=\frac{1}{1+C(z)G(z)}$$
%
%Note that $H(z)$ cancels the poles of both $G(z)$ and $C(z)$, thus we need only verify the stability of $H(z)$
%to validate the stability of the whole system.
%These equations are sufficient for computing $\sigma_{Q}^{2}$.  However,
%there is still a gap for a complete model for $\Delta{G_{Q}(z)}$ in
%Eq.~\eqref{eq:deltag_var}, {\it i.e.}, the z-transform of control action
%$U(z)$.  This indicates that the quantisation noise effect on a discrete
%plant model mainly depends on three main factors:
%%
%\begin{itemize}
%%
%\item \textbf{Sample time:} the sample time effect is reflected over pulse transfer function that covers the effects of DAC ant its interpolator (ZOH) \red{[clarify/revise based on the comments made above.]};
%%
%\item \textbf{ADC and DAC step:} \red{[again, do we need both?]} these are the main component of quantisation noise power, but an $8$-bits converter already reduces the effect of quantisation noises in the output for negligible values.
%%
%\item \textbf{Signal-to-noise ratio (SNR): } the dependence on the value of $U(z)$ \red{[notice that now we have reduced the problem to depend on the effect of the reference signal transform R(z) - pls revise.]} shows the importance of the SNR for handling quantisation noise; for a high SNR of the digital controller, the quantisation effect is negligible. This happens during the transient of the system, when high control actions are employed. However, it can be relevant during the steady state. A digital controller with integral action might be sufficient for eliminating such influence at steady state.
%%
%\end{itemize}
%
%An alternative expression for $\Delta{G_{Q}(z)}$ can be obtained from
%Eq.~\eqref{eq:deltag_var}, transforming it into a function of the reference
%$R(z)$.  Once $U(z)$ is the control signal obtained via digital controller
%$C(z)$ and error signal, {\it i.e.},
%%
%\begin{equation}
%U(z)=C(z)\cdot[R(z)-Y(z)],
%\end{equation} 
%
%\noindent where $Y(z)$ is the first term of Eq.~\eqref{eq:outputfunctions}, thus:
%\begin{equation}
%\label{eq:controlsignal}
%\resizebox{0.48\textwidth}{!}{
%$U(z)=C(z)\cdot \left[R(z)-\frac{C(z)G(z)}{1+C(z)G(z)}R(z)\right]=\frac{C(z)}{1+C(z)G(z)}R(z).$
%}
%\end{equation} 
%
%Substituting Eq.~\eqref{eq:controlsignal} in Eq.~\eqref{eq:deltag_var}, a
%model for $\Delta{G_{Q}(z)}$ depending on $R(z)$ can be obtained:
%%
%\begin{equation}
%\Delta{G_{Q}(z)}=\sigma^{2}_{Q}\cdot \frac{1+C(z)G(z)}{C(z)} \cdot \frac{1}{R(z)},
%\end{equation}
%
%\noindent where $R(z)$ can be considered as a step signal (for stabilization
%purpose) with height equal to $r$, which can be rewritten as
%%
%\begin{equation}
%\Delta{G_{Q}(z)}=\frac{\sigma^{2}_{Q}}{r}\cdot \frac{1+C(z)G(z)}{C(z)} \cdot \frac{z-1}{z}.
%\end{equation}
%
%A more precise description of $\Delta{G_{Q}(z)}$ can be obtained
%considering FWL \red{[clarify what you mean with FWL, introduce FWL function explicitly and precisely]} effects on digital controller coefficients, as described in
%Bessa {\it et al.}~\cite{Bessa16}, {\it i.e.}, using $FWL[C(z)]$ instead of
%the nominal $C(z)$:
%%
%\begin{equation}
%\label{eq:deltaG_final}
%\Delta{G_{Q}(z)}=\frac{\sigma^{2}_{Q}}{r}\cdot \frac{1+FWL[C(z)]G(z)}{FWL[C(z)]} \cdot \frac{z-1}{z}.
%\end{equation}
%
%Finally, Eq.~\eqref{eq:deltaG_final} is a suitable frequency domain model
%for quantisation noise effects on closed-loop systems.  This model depends
%on the plant pulse transfer function ($G(z)$), which is described by
%Eq.~\eqref{eq:pulsetf}, the height of step reference signal $r$, and the
%digital controller model.

%\subsection{Modelling Error on Parameters Representation} 
%\label{sec:}

%\red{[TO BE COMPLETED.]}
% We use CounterExample-Guided Inductive Synthesis (CEGIS), as
% illustrated in Fig~\ref{fig:CEGIS}.

%specialised it for stream refactoring.  
% adding JST for stream refactoring

\subsection{Architecture of the program synthesiser}
\label{synthesiser-general}
%
%% \red{[I would recommend to clarify the domain of definition of $\sigma$, particularly over the set where $x$ ranges over. 
%% (Be cognisant that control audience might be unfamiliar with the cegis architecture.) ]}
% 
The input specification provided to the program synthesiser is of the form
$\exists \vec{P} .  \forall \vec{x}.  \sigma(\vec{x}, \vec{P})$ where
$\vec{P}$ ranges over functions, $\vec{x}$ ranges over ground terms and
$\sigma$ is a quantifier-free formula.  We interpret the ground terms over
some finite domain $\mathcal{D}$.

The design of our synthesiser is given in Figure~\ref{fig:CEGIS} and consists
of two phases, {\sc Synthesise} and {\sc Verify}, which interact via a
finite set of test vectors {\sc inputs} that is updated incrementally. 
Given the aforementioned specification $\sigma$, the {\sc synth} procedure
tries to find an existential witness $\vec{P}$ satisfying the specification
$\sigma(\vec{x}, \vec{P})$ for all $\vec{x}$ in {\sc inputs} (as opposed to
all $\vec{x} \in \mathcal{D}$).
%
If {\sc synthesise} succeeds in finding an witness $\vec{P}$, this witness
is a candidate solution to the full synthesis formula.  We pass this
candidate solution to {\sc verify}, which checks whether it is a full
solution ({\it i.e.}, $\vec{P}$ satisfies the specification $\sigma(\vec{x},
\vec{P})$ for all $\vec{x}\in\mathcal{D}$).
%
%determines whether
%it does satisfy the specification on all inputs by checking
%satisfiability of the following verification formula:
%
%$\exists \vec{x} . \lnot \sigma(\vec{x}, \vec{P})$
%If this formula is unsatisfiable, 
%% the candidate solution is in fact a 
%% full solution.
%
If this is the case, then the algorithm terminates.  Otherwise, additional
information is provided to the {\sc synthesise} phase in the form of a new
counterexample that is added to the {\sc inputs} set and the loop iterates
again (note that, for now, we ignore the second feedback signal ``Increase
Precision'' provided by the {\sc Verify} phase in Figure~\ref{fig:CEGIS} as it
is specific to control synthesis and will be described in the next section).

%% the witness $\vec{x}$ denote an input on which the
%% candidate solution fails to meet the specification.  Thus, $\vec{x}$ is
%% added to the {\sc inputs} set, and the loop iterates again.

It is
worth noting that each iteration of the loop adds a new input to the
finite set $\text{\sc inputs}$ being used for synthesis.  Given that
the full set of inputs $\mathcal{D}$
%\blue{[which by now should have a name, as suggested to be introduced above]} 
is finite, this means that the refinement loop
can only iterate a finite number of times.

\subsection{Synthesis for control}
\label{synthesis-elements}

\paragraph{Specification of the stability property}

Next, we describe the specific property that we pass to the program
synthesiser as the specification $\sigma$.  There are two different
algorithms in DSVerifier that can be used for stability
verification~\cite{daes20161,Bessa16}, one based on
Schur's decomposition and another one based on Jury's
criteria~\cite{astrom1997computer}.  Here, we choose Jury's
method~\cite{astrom1997computer}, due to its efficiency.

We use this method to check the stability in the $z$-domain for the
characteristic polynomial $S(z)$ defined
in~\eqref{eq:internal_stab_lemma}.

% of the matrix $\left( \begin{array}{c} Y \\ U \end{array}\right)$, where $Y(z)$ is the closed-loop system output and U(z) is the controller output. 
%% \red{[unclear: what matrix? We need to clarify the relationship
%% between the char. poly. $S(z)$ and the later quantities $\Delta
%% N_{G}(z), \Delta D_{G}(z)$ and corresponding controller's
%% structure.]}.
%
We consider the following form for $S(z)$:
%
$$
S(z) = a_0z^N+a_1z^{N-1}+\cdots+a_{N-1}z+a_N=0, a_0\neq0
$$
%
%\cdsay{is this for the matrix $\left( \begin{array}{c} Y \\ U \end{array}\right)$?}

% Assuming that $\hat{G}(z) = \frac{\hat{N_G}(z)}{\hat{D_G}(z)}$
% and $\hat{C}(z) = \frac{\hat{N_C}(z)}{\hat{D_C}(z)}$, then:
% $$S(z) = FWL[N_C(z)] \times \hat{N_G}(z) + FWL[D_C(z)] \times \hat{D_G}(z)$$

Next, the following Jury matrix
$M = [m_{ij}]_{(2N−2)\times N}$ is built from $S(z)$ coefficients:
%
$$
M=\left( 
\begin{array}{c}
V^{(0)}\\
V^{(1)}\\
\vdots\\
V^{(N-2)}
\end{array}
\right), 
$$
%
where $V^{(k)} = [v^{(k)}_{ij} ]_{2\times N}$ such that:
%
$$
v_{ij}^{(0)}=\left\{
\begin{array}{ll}
a_{j-1}, & \texttt{if}~i=1\\
v_{(1)(N-j+1)}^{0},&\texttt{if}~i=2
\end{array}
\right.
$$
%
$$
v_{ij}^{(k)}=\left\{
\begin{array}{ll}
0,&\texttt{if}~j>n-k\\
v_{1j}^{(k-1)}-v_{2j}^{(k-1)} . \frac{v_{11}^{(k-1)}}{v_{21}^{(k-1)}}, & \texttt{if}~j\leq n-k ~\texttt{and}~i=1\\
v_{(1)(N-j+1)}^{k},& \texttt{if}~j\leq n-k ~\texttt{and}~i=2\\
\end{array}
\right.
$$
%
where $k \in \mathbb{Z}$, such that $0 < k < N - 2$. 
$S(z)$ is the
characteristic polynomial of a stable system if and only if the following four conditions hold:
\begin{itemize}
\item $R_1: S(1) > 0$;
\item $R_2: (−1)^N S(−1) > 0$;
\item $R_3: |a_0| < a_N$;
\item $R_4: m_{11} > 0 \wedge m_{31}>0 \wedge m_{51}>0 \wedge \cdots \wedge m_{(2N{-}3)(1)}>0$.
\end{itemize}

The stability property is then encoded by creating a
constraint of the form:
$$
\phi_\mathit{stability} \equiv (R_1 \wedge R_2 \wedge R_3 \wedge R_4).
$$
%\cdsay{figure out what is existentially quantified for the final $\sigma$.} 
%% \red{[Yes, intuitively all clear, however for the sake of clarity I would like to see the explicit mapping between the above formula and the previously used specification 
%% $\sigma(x, P)$, with clear elucidation of the relationship between the domain of $z$ and that of $x$.]}

% \red{[clear elucidation of the mapping between this and the domains used 
% in the general description of the synthesiser]}

\paragraph{Fixed-point computation in program synthesis}

The program synthesis engine uses fixed-point arithmetic.  Namely, we use
the domain $\mathbb{R}\param{I}{F}$ for the controller's coefficients and
the domain $\mathbb{R}\param{I_b}{F_b}$ for the plant's coefficients, 
where $I$ and $F$, as well as $I_b$ and $F_b$, denote the integer and
fractional size, respectively, and $\mathbb{R}\langle I_b,F_b \rangle
\supseteq \mathbb{R}\langle I,F \rangle$.
%
%the coefficients for $C(z)$ 
%$N_c$ and $D_c$ 
%in the domain $\mathbb{R}\param{I}{F}$, 
%and $\mathbb{R}^{m_C}\param{I}{F}$, respectively, 
%where $n_C$ and $m_C$ denote the order of the corresponding polynomials
%where $I$ and $F$ denote the integer and fractional size, respectively.
%Regarding the plant, its coefficients are this is synthesised over the 
%domain $\mathbb{R}^{n_G+m_G}\param{I_b}{F_b}$, where $n_G$ and $m_G$ 
%denote the order of the polynomials representing the 
%numerator and denominator of the plant's transfer function, respectively. 
%Again, $I_b$ and $F_b$ denote the integer and fractional size, respectively, with
%$\mathbb{R}\langle I_b,F_b \rangle \supseteq \mathbb{R}\langle I,F \rangle$.

Given the use of fixed-point arithmetic, 
we need to examine the effect of discretisation during these operations.
Let $\tilde C(z)$ and $\tilde G(z)$ be the transfer functions represented
using fixed-point bit vectors.%  with integer size $I_b$ and fractional
% size $F_b$, whose domain is $\mathbb{R}\langle I_b,F_b \rangle \supseteq \mathbb{R}\langle I,F \rangle$.
%
\begin{align}
\small
\label{digital_controller_plant_tf}
\tilde C(z)&=\frac{\tilde \beta_{0}+\tilde \beta_{1}z^{-1}+...+\tilde \beta_{M_C}z^{-M_C}}{\tilde \alpha_{0}+\tilde \alpha_{1}z^{-1}+...+\tilde \alpha_{N_C}z^{-N_C}}, \\
%\label{plant_tf}
\tilde G(z)&=\frac{\tilde b_{0}+\tilde b_{1}z^{-1}+...+\tilde b_{M_G}z^{-M_G}}{\tilde a_{0}+\tilde a_{1}z^{-1}+...+\tilde a_{N_G}z^{-N_G}}.
\end{align}

Since the controller is synthesised in the $\mathbb{R}\langle I,F \rangle$
domain,
%
 $$\forall i \leq N_C, j \leq M_C\ \  \tilde \beta_{i} \equiv \beta_{i} \wedge \tilde \alpha_{j} \equiv \alpha_{j} \Leftrightarrow \tilde C(z) \equiv C(z)$$
%
However, given a real plant $\hat{G}(z)$, and a discretising function
%
\begin{align}
\small
\label{FWL_tf}
&f_b : \mathbb{R} \rightarrow  \mathbb{R}\langle I_b,F_b \rangle \triangleq \tilde x = f_b(x) : \tilde x=x-(x\mod\tilde x)\\
&FWL_b[P \in \mathcal{P}] = \tilde P \in \mathcal{P}^{n}\langle I_b,F_b \rangle : c_i \in \vec{P} \wedge \tilde c_i \in \tilde \vec{P}=f_b(c_i)  \nonumber
\end{align}
%
we calculate $\tilde G(z)=FWL_b(\hat{G}(z))$
%
\begin{align}
% \label{digital_controller_tf}
% \tilde C(z)&=\frac{(\hat{\beta}_{0}+\Delta_q \hat{\beta}_{0}) +...+(\hat{\beta}_{M_C}+\Delta_b \hat{\beta}_{M_C})z^{-M_C}}{(\hat{\alpha}_{0}+\Delta_q \hat{\alpha}_{0})+...+(\hat{\alpha}_{N_C}+\Delta_q \hat{\alpha}_{N_C})z^{-N_G}} \nonumber \\
% &= C(z) =\hat{C}(z) + \Delta{C}_b(z)\\
\label{digital_plant_tf}
\tilde G(z)&=\frac{(\hat{b}_{0}+\Delta_b \hat{b}_{0}) +...+(\hat{b}_{M_G}+\Delta_b \hat{b}_{M_G})z^{-M_G}}{(\hat{a}_{0}+\Delta_b \hat{a}_{0})+...+(\hat{a}_{N_G}+\Delta_b \hat{a}_{N_G})z^{-N_G}} \nonumber \\
&\Rightarrow \vec{\tilde G} =\vec{\hat{G}}+\Delta_b{\vec{G}}=\vec{G}+\Delta_p{\vec{G}}+\Delta_b{\vec{G}}.
\end{align}
where $\Delta_b\hat{c}_i=\hat{c}_i\mod \tilde{c}_i$
and $\Delta_b{G}$ represents the plant uncertainty caused by
the rounding off effect.  We may encompass all the uncertainty as
$\Delta{\vec{G}}=\Delta_p{\vec{G}}+\Delta_b{\vec{G}}$. 
Since the complexity of SAT solvers is worst-case exponential in the size of
the problem instance, our algorithm tries to solve the problem first for a
low $I_b$, $F_b$, iteratively increasing the precision if it is insufficient.

\paragraph{Our synthesis problem}
The synthesis problem we are trying to solve is the following:
find a digital controller $C(z)$ %denoted by  $N_c$ and $D_c$ 
that makes the closed-loop system stable 
for all possible uncertainties 
%$\Delta{\vec{p}}$ (\ref{eq:uncertainty}).
$\tilde G(z)$ (\ref{digital_plant_tf}).
%, where we precisely describe 
%$(\Delta N_G, \Delta D_G)$. 
% Note that, similar to the rest of the paper, we use the notation 
% $\tilda{N_G}(z) = N_G(z) + \Delta N_G(z)$ and 
% $\hat{N_C}(z) = N_C(z) + \Delta N_C(z)$.
When mapping back to the notation used for describing the general architecture 
of the program synthesiser, the controller $C(z)$ denotes $P$ and 
$\tilde G(z)$ represents $x$.

As mentioned above, we compute the coefficients for $C(z)$ 
%$N_c$ and $D_c$ 
in the domain $\mathbb{R}\param{I}{F}$, 
%and $\mathbb{R}^{m_C}\param{I}{F}$, respectively, 
%where $n_C$ and $m_C$ denote the order of the corresponding polynomials
and those for $\tilde G(z)$ in the domain
$\mathbb{R}\langle I_b,F_b \rangle$.
While the controller's precision $\param{I}{F}$ is given, 
we can vary $\param{I_b}{F_b}$ such that 
$\mathbb{R}\langle I_b,F_b \rangle \supseteq \mathbb{R}\langle I,F \rangle$.

\paragraph{The {\sc Synthesise} and {\sc Verify} phases}
The {\sc synthesise} phase uses BMC to 
compute a solution $C(z)$. % An important remark is that this 
% controller is synthesised in the $\mathbb{R}\langle I,F \rangle$
% domain.

There are two approaches for the {\sc verify} phase.
% based on the described model.  
The first approach uses interval arithmetic \cite{moore1966interval}
to represent the coefficients
$[{c}_i-\Delta_p{c}_i-\Delta_b{c}_i\ ,\
{c}_i+\Delta_p{c}_i-\Delta_b{c}_i+(2^{-F_b})]$ 
and rounds operations outwards during calculations. This effectively allows
us to simultaneously evaluate the whole collection of plants $\mathfrak{G}$
plus the effects of numeric calculations in the engine.
Synthesised controllers using this verifier will always
be stable for all plants in the family.
% and only generate counterexamples when the precision of the
%engine is insufficient (\emph{ie} no stable plant could be found). 
Our preliminary experiments showed that a synthesis approach
using this verification engine has poor performance and we
%Given the high computational cost of this
%approach, 
therefore designed a second approach, which we describe below.
Our experimental results described in Section~\ref{sec:experiments}, 
%try both approaches and provide the time for the fastest. 
show that the speedup yielded by the second approach 
is in most cases of at least two orders of magnitude.

The second approach is illustrated in Figure~\ref{fig:CEGIS} and uses a
two-stage verification: the first stage performs potentially unsound
fixed-point operations assuming a precision \param{I_b}{F_b}, and the
second stage restores soundness by validating these operations
using interval arithmetic on the synthesised controller.

In more detail, in the first stage, denoted by {\sc Uncertainty} in Figure~\ref{fig:CEGIS}, 
assuming a precision \param{I_b}{F_b} 
we check whether the system is unstable for the current candidate solution,
{\it i.e.}, if $\neg \phi_\mathit{stability}$ is satisfiable for $S(z)$.
If this is the case, then we obtain a counterexample 
$\tilde G(z)$, 
%denoted by values for each coefficients of $\Delta N_{G}(z)$ and $\Delta D_{G}(z)$, 
%\red{[see my comment above - these quantities pop up too abruptly.]},
which makes the closed-loop system unstable. 
This uncertainty is added to 
the set {\sc inputs} such that, in the subsequent {\sc synthesise} phase, 
we obtain a candidate solution consisting 
of a controller $C(z)$, which makes the closed-loop 
system stable for all the uncertainties accumulated in {\sc inputs}.

If the {\sc Uncertainty} verification stage concludes that
the system is stable for the current candidate solution,
then we pass this solution to the second verification stage, {\sc Precision}, 
which checks the propagation of the error in the fixed-point 
calculations using a Fixed-point Arithmetic Verifier based on 
interval arithmetic. 
%
% To achieve this, we use a nondeterministic plant selected from $\mathfrak{G}(z)$ where the expected error of the bit vector calculations is assumed to be within the expected bounds and verified after synthesis using the interval version of the program. 
%
% The first verification phase produces a counterexample if some instance of $\mathfrak{G}(z)$ does not verify the properties and the CEGIS loop iteratively looks for a new controller that verifies the supplied counterexamples. The second phase verifies the propagation of the error in the fixed point calculations.
% If the second phase fails verification, we increase the precision of the analyser and feed the CEGIS loop again to find a sound solution. 
% vouches for the use of CEGIS in most of our benchmarks.
% \cdsay{end of integration}

If the {\sc precision} verification returns $\mathit{false}$, then we
increase the precision of \param{I_b}{F_b} and re-start the {\sc
  synthesise} phase with an empty {\sc inputs} set.  Otherwise, we
found a full sound solution for our synthesis problem and we are done.

In the rest of the paper, we will refer to the two approaches 
for the {\sc verify} phase as one-stage and two-stage, respectively.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Illustrative Example} \label{sec:running-ex}

We will illustrate our approach with a classical cruise control example,
extracted from the literature~\cite{Astrom08}.  The example will allow us to
highlight the challenges that arise when using finite precision arithmetic
in digital control.  We are given a discrete plant model (with a time step
of $0.2$s), which is represented by the following $z$-expression:
%
\begin{equation}
\label{Eq:running-example-plant}
G\left(z\right) := \frac{0.0264}{z-0.9998}.
\end{equation}

Using an optimisation tool, the authors
of~\cite{DBLP:conf/hybrid/WangGRJF16} have designed a high-performance
controller for this plant, which is characterised by the following
$z$-domain transfer function:
%
\begin{equation}
\label{Eq:running-example-controller}
C\left(z\right) := \frac{2.72z^2 - 4.153z + 1.896}{z^2 - 1.844z + 0.8496}.
\end{equation}
%
The authors of~\cite{DBLP:conf/hybrid/WangGRJF16} claim that this controller
$C(z)$ (Eq.~\eqref{Eq:running-example-controller}) stabilizes the
closed-loop system for the discrete plant model $G(z)$
(Eq.~\ref{Eq:running-example-plant}).  However, if the effects of finite
precision arithmetic are considered during verification (e.g., fixed-point
arithmetic and word-length), then this closed-loop system becomes unstable.
%
For instance, an implementation of $C(z)$ using \param{4}{16} fixed-point
numbers (i.e., $4$ bits for the integer part and $16$ bits for the
fractional part) can be modeled as follows:
%
\begin{equation}
\label{Eq:running-example-controller-quantised}
C_{q}\left(z\right) {:=} \frac{2.719986z^2{-}4.153000z
{+}1.895996}{z^2{-}1.843994z+0.849594} \;.
\end{equation} 
%
The resulting system using the typical series loop configuration, where
$C_{q}\left(z\right)$ and $G\left(z\right)$ are in the forward path, is
unstable.

\subsection{Program synthesis for the example}

We now demonstrate how our approach solves the synthesis problem for the
example given in the previous section.  We start with the candidate solution
with all coefficients zero, as given below, and a precision of $I_b=16$,
$F_b=24$.  The digital controller is designed with standard techniques for
control systems~\cite{Kuo:2002:ACS:579453, Ogata:1987:DCS:26170}.
%
%\red{[Be more explicit, write out spurious candidate solution with zero coefficients. KISS - you don't want to lose the control audience here.]}.
%
%$$
%\begin{array}{ll}
%N_C(z) {=} 0z^2{+}0z{+}0\\
%D_C(z) {=} 0z^2{+}0z{+}0\\
%\end{array}
%$$
$$ C(z)=\frac{0z^2{+}0z{+}0}{0z^2{+}0z{+}0} $$

In the first {\sc verify} stage, the {\sc uncertainty} 
check finds the following counterexample:
%
$$ \tilde G(z) = \frac{0.026506}{1.000610z+1.002838} $$

%$$
%\begin{array}{ll}
%\Delta N_G(z) = 0.026506\\
%\Delta D_G(z) = 1.000610z+1.002838
%\end{array}
%$$

We add this counterexample to {\sc inputs} and initiate the {\sc synthesise}
phase, where we obtain the following candidate solution:
%
%\red{[ How is the order of the candidate solution chosen here?]}
$$
C(z)=\frac{12.402664z^2{-}11.439667z{+}0.596756}{4.003906z^2{-}0.287949z{+}0.015625}
$$

This time, the {\sc uncertainty} check does not find any
counterexample and we pass the current candidate solution to the {\sc
precision} verification stage.
%
% In the second {\sc verify} phase, we check whether the current candidate
% solution above is the final solution. It is not and the verifier signals
% that it is due to lack of precision.
%
We obtain the result $\mathit{false}$, meaning that the current precision is
insufficient.  Consequently, we increase our precision to $I_b=20$, $F_b=28$.
%
Since the previous counterexamples were obtained at lower precision, we
remove them from the set of counterexamples.  Back in the {\sc synthesise}
phase, we re-start the process with a candidate solution with all
coefficients $0$.  Next, the {\sc uncertainty} verification stage provides
the first counterexample at higher precision:
%
%$$
%\begin{array}{ll}
%\Delta N_G(z) {=} 0.026314\\
%\Delta D_G(z) {=} 0.999024z{-}1.004785
%\end{array}
%$$
$$ \tilde G(z) = \frac{0.026314}{0.999024z{-}1.004785} $$

In the next {\sc synthesise} phase, we get a new candidate solution that
eliminates the new, higher precision counterexample:
%
$$ C(z)=\frac{11.035202z^2{+}5.846100z{+}4.901855}{1.097901z^2{+}0.063110z{+}0.128357} $$
%
%$$
%\begin{array}{ll}
%N_C(z) {=} \\
%D_C(z) {=} \\
%\end{array}
%$$

This candidate solution is validated as the final solution by both stages
{\sc uncertainty} and {\sc precision} in the {\sc verify} phase.
In particular, Figure~\ref{fig:bode} shows the Bode diagram of the illustrative example from~\cite{DBLP:conf/hybrid/WangGRJF16} 
(Fig.~\ref{fig:original}) and from our synthesiser (Fig.~\ref{fig:cegiscontroller}). 
The phase margin is negative as shown in Figure~\ref{fig:original}, which means that the controller 
from~\cite{DBLP:conf/hybrid/WangGRJF16} is unstable, considering FWL effects. 
However, the final solution from \tool is stable since it presents an infinite phase margin 
and a gain margin of $17.8 dB$.

%, which fails to find a new counterexample (i.e. returns UNSAT).

% \blue{
% For the same cruise control system, the synthesis may be done considering that the plant model have parametric uncertainties ($\Delta G_{P}(z)$). Thus, the synthesised controller must to stabilize for any parametric variation $\Delta p$ inside of a given region. Specially, in this example, was assumed that each coefficient of $G(z)$ may vary $\pm 0.5\%$.
% }

% \blue{Initially, the controller denominator and numerator  ($D_{C}(z)$ and $N_{C}(z)$) are started with zero coefficients, as in previous example. Thus,  in the first {\sc verify} phase a counterexample is presented, that is added to {\sc inputs}, whis is usede in the first  {\sc synthesise} phase. The resulting controller is:
% }
% $$
% {
% \blue{
% C(z)=\frac{0.77516z^2{+}-0.71497z{+}0.03729}{0.25024z^2{-}0.01799z{+}0.00097}
% }}
% $$

% \blue{In the second {\sc verify} phase, the controller failed to stabilize the system for all the range of uncertainties, another counterexample was provided, and the following candidate controller is presented in another {\sc synthesise} phase:}
% $$
% {
% \blue{
% C(z)=\frac{0.68970z^2{+}0.36538z{+}0.30637}{0.06862z^2{+}0.00394z{+}0.00802}
% }}
% $$

% \blue{
% This candidate solution is validated as the final solution by the next 
% {\sc verify} phase.
% }

\begin{figure*}
\centering
\resizebox{.8\textwidth}{!}{
 \begin{tikzpicture}[scale=0.3,->,>=stealth',shorten >=.2pt,auto, semithick, initial text=, ampersand replacement=\&,]
  \matrix[nodes={draw, fill=none, shape=rectangle, minimum height=.2cm, minimum width=.2cm, align=center
},
          row sep=1cm, column sep=1.5cm] {
   \coordinate (aux1);
   \& \coordinate (aux2);
   \&;\\
   \node[fill=gray!20] (synth) {{\sc Synthesise}};
   \&
   complexnode/.pic={ 
     \node[rectangle,draw,dashed,
	minimum width=7cm,
	minimum height=.8cm,
	label={\sc Verify},] (verif) {};
   \node[fill=gray!20] (verif1) at ([xshift=-2cm]verif.center) {{\sc ~Uncertainty~~~~}};
   \node[fill=gray!20] (verif2) at ([xshift=2cm]verif.center) {{\sc ~~~~~Precision~~~}};
   } 
   \& \node[ellipse, fill=gray!20] (done) {{\sc Done}};\\
   %% \node[fill=yellow!20] (verif) {{\sc ~~~~~~Uncertainty~~~~~~}};
   %% \&
   %% \node[fill=yellow!20] (verif2) {{\sc ~~~~~~Precision~~~~~~~}};\\
   \& \\
   \& \\
   \node (gp) {Closed-loop System Search};
   \&
   complexnode/.pic={ 
     \coordinate (aux);
   \node (bmc) at ([xshift=-2cm]aux.center) {BMC-based Verifier};
   \node (fp)  at ([xshift=2cm]aux.center) {Fixed-point \\Arithmetic Verifier};
   }   
    \\
  };

   \path
    ([yshift=2em]synth.east) edge node[xshift=-0.5em] {Candidate solution} ([yshift=2em]verif1.west)
    ([yshift=-2em]verif1.west) edge node {Counterexample} ([yshift=-2em]synth.east)
    ([xshift=5em]verif1.south) edge node[align=center] {Candidate\\ $P$} ([xshift=5em]bmc.north)
    ([xshift=5em]verif2.south) edge node[align=center] {Candidate\\ $P$} ([xshift=5em]fp.north)
    ([xshift=-5em]bmc.north) edge node[align=center]  {UNSAT/\\model} ([xshift=-5em]verif1.south)
    ([xshift=-5em]fp.north) edge node[align=center]  {True/\\False} ([xshift=-5em]verif2.south)
    (verif2) edge node {} (done)
    ([xshift=5em]synth.south) edge node[align=center] {Inputs} ([xshift=5em]gp.north)
    ([xshift=-5em]gp.north) edge node[align=center] {UNSAT/\\candidate} ([xshift=-5em]synth.south)
    (aux1) edge (synth.north);
   \path[-]
   (verif2.north) edge node[align=center] {} ([xshift=6.7cm]aux2)
   ([xshift=6.7cm]aux2) edge node[align=center] {Increase Precision} (aux1);

 \end{tikzpicture}
}
 \caption{Counterexample-Guided Inductive Synthesis of Closed-loop System (Step~D)
\label{fig:CEGIS}}
\end{figure*}

\begin{figure}
    \centering
    \begin{subfigure}[b]{0.45\textwidth}
        \includegraphics[width=\textwidth]{figures/runningexample_bd0.pdf}
        \caption{Original controller~\cite{DBLP:conf/hybrid/WangGRJF16}.}
        \label{fig:original}
    \end{subfigure}
    ~
    \begin{subfigure}[b]{0.45\textwidth}
        \includegraphics[width=\textwidth]{figures/runningexample_bd2.pdf}
        \caption{Controller synthesised by \tool.}
        \label{fig:cegiscontroller}
    \end{subfigure}
    \caption{Bode diagram for original~\cite{DBLP:conf/hybrid/WangGRJF16} and synthesised closed-loop systems.}\label{fig:bode}
\end{figure}

Figure~\ref{fig:step} shows the step responses of the closed-loop system 
with the original controller~\cite{DBLP:conf/hybrid/WangGRJF16} (Figure~\ref{fig:step0}), 
the first candidate solution (Figure~\ref{fig:step1}), and the final solution from \tool (Fig.~\ref{fig:step2}). 
The step response in Figure~\ref{fig:step0} confirms the stability loss if we consider FWL effects. 
However, Figure~\ref{fig:step2} shows that the first candidate controller is able to stabilize 
the closed-loop system without uncertainties, but it is rejected by \tool since this solution 
is not considered to be sound. Finally, Figure~\ref{fig:step2} shows a stable behavior 
for the final (sound) solution, which presents a lower settling time.

\begin{figure*}
    \centering
    \begin{subfigure}[b]{0.3\textwidth}
        \includegraphics[width=\textwidth]{figures/runningexample_step0.pdf}
        \caption{Orignal controller.}
        \label{fig:step0}
    \end{subfigure}
    ~
    \begin{subfigure}[b]{0.3\textwidth}
        \includegraphics[width=\textwidth]{figures/runningexample_step1.pdf}
        \caption{First candidate solution.}
        \label{fig:step1}
    \end{subfigure}
    ~
    \begin{subfigure}[b]{0.3\textwidth}
        \includegraphics[width=\textwidth]{figures/runningexample_step2.pdf}
        \caption{Final solution by \tool.}
        \label{fig:step2}
    \end{subfigure}
    \caption{Step responses for original~\cite{DBLP:conf/hybrid/WangGRJF16} closed-loop system with FWL effects and for each {\sc synthesise} step of \tool.}\label{fig:step}
\end{figure*}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Experimental Evaluation}\label{sec:experiments}

This section is split into three parts. Section~\ref{experimental-setup}
discusses the experimental setup and the employed benchmarks, while
Section~\ref{experimental-objectives} describes the experimental objectives. 
Section~\ref{experimental-results} describes the experimental results
related to the application of CEGIS to the standard closed-loop control
system benchmarks described in Section~\ref{experimental-setup}.

%---------------------------------
\subsection{Description of the Benchmarks}
\label{experimental-setup}
%---------------------------------

The first set of benchmarks uses the discrete plant of a cruise control
model for the car accounting for rolling friction, aerodynamic drag, and
gravitational disturbance force~\cite{Astrom08} as follows
%
\begin{equation}
\label{cruise-control-c1}
G_1(z)=\frac{0.0264}{z-0.998}. \nonumber
\end{equation}

%Two proportional-integral (PI) controllers are designed by Wang {\it et al.} 
%for the cruise control plant $G(z)$~\cite{DBLP:conf/hybrid/WangGRJF16}, 
%which has the following mathematical model
%
%\begin{equation}
%\label{cruise-control-g1}
%C_{1}(z)=\frac{1.51z-1.3586}{z-1}, \nonumber
%\end{equation} 
%
%\begin{equation}
%\label{cruise-control-g2}
%C_{2}(z)=\frac{2.72z^2 - 4.153z + 1.896}{1.0z^2 - 1.844z + 0.8496}. \nonumber
%\end{equation}

The second set of benchmarks considers a simple spring-mass 
damper~\cite{DBLP:conf/hybrid/WangGRJF16}, where the discrete 
plant dynamics is represented by the following $z$-expression
%
%where both discrete 
%plant dynamics and controller are represented by the following 
%z-expression, resp.
%
\begin{equation}
\label{spring-mass-damper-g}
G_2(z)=\frac{5\times{10^{-5}}z + 5\times{10^{-5}}}{z^2 - 2z + 1.0001}. \nonumber
\end{equation} 
%
%\begin{equation}
%\label{spring-mass-damper-c}
%C_3(z)=\frac{1.28\times{10^{3}}z^2 - 1.354240\times{10^{3}}z + 0.074879\times{10^{3}}}{z^2 - 1.4990z + 0.4995}. \nonumber
%\end{equation}

A third set of benchmarks that consists of a physical plant for satellite
applications is considered~\cite{Franklin15}, in order to evaluate the
full capabilities of the present approach.  In particular, satellites often require
attitude (pose) control for proper orientation of antennas and sensors
w.r.t.~earth.  The satellite attitude control is typically used for
three-axis attitude tracking, but here only one axis at a time is
considered.  The motion equation for the one-axis system, despising
disturbance torques, is given by
%
\begin{equation}
\label{eq:satelliteode}
I\cdot \ddot{\theta} = \tau_{C}, 
\end{equation}

\noindent where $I$ is the inertia moment of satellite about its mass
center, $\tau_{C}$ is the control torque applied by thusters, and $\theta$
is the controlled attitude angle.  Normalizing Eq.~\eqref{eq:satelliteode}
by defining the control signal $u=\frac{\tau_{C}}{I}$ and taking Laplace
transform, the following $s$-domain transfer function $G(s)$ is obtained
%
\begin{equation}
\label{eq:satellitetf}
G_{3}(s)=\frac{\theta(s)}{u(s)}=\frac{1}{s^2},
\end{equation}

\noindent using the ZOH discretisation defined in Eq.~\eqref{eq:pulsetf}, 
the following $z$-domain model is obtained
%
\begin{equation}
G_{3}(z)= \frac{T^{2}}{2} \frac{z+1}{(z+1)^{2}}. \nonumber
\end{equation}

%The discrete controllers $C_{4}$ and $C_{5}$ for the satellite plant $G_{3}$, discretized with sample time $1s$ and $2s$ respectively, 
%are represented by the following z-expression:
%
%\begin{equation}
%\label{satellite-b2}
%C_{4}(z)=\frac{2.88z^2 - 4.896z + 2.074}{1.0z^2 - 0.42z - 0.3465 - 0.03915}. \nonumber
%\end{equation} 
%
%\begin{equation}
%\label{satellite-c2}
%C_{5}(z)=\frac{0.43z^2 - 0.2408z - 0.1316}{1.0z^2 + 1.5z + 0.97 + 0.183}. \nonumber
%\end{equation}

The fourth (and last) set of benchmarks considers the following description
for the plant $G_{4}(s)$, which is typically used for evaluating stability
margins~\cite{bhattacharyya97, keel_Bhattacharyya_examples}
%
\begin{equation}
\label{exampleA}
G_{4}(s)=\frac{s-1}{s^{2}-s-2},
\end{equation}
%
\noindent where $G_{4}(z)$ is obtained using the ZOH discretisation defined
in Eq.~\eqref{eq:pulsetf} with three different sample times ($0.1$, $0.05$,
and $0.03$ seconds), respectively, as follows:
%
\begin{equation}
\label{exampleA-sampletime1}
G_{4a}(z)=\frac{0.100342181002722z - 0.110876810062963}{1.0z^{2} - 2.12624017619613z + 1.10517091807565}, \nonumber
\end{equation}
%
%
\begin{equation}
\label{exampleA-sampletime2}
G_{4b}(z)=\frac{0.0500422033454653z - 0.0526068264456340}{1.0^{2} - 2.05640034257636z + 1.05127109637602}, \nonumber
\end{equation}
%
%
\begin{equation}
\label{exampleA-sampletime3}
G_{4c}(z)=\frac{0.0300090687252212z - 0.0309228417953966}{1.0^{2} - 2.03228208009387z + 1.03045453395352}. \nonumber
\end{equation}

All experiments were conducted on a 12-core 2.40\,GHz Intel Xeon E5-2440
with 96\,GB of RAM and Linux OS.  All times given are wall clock time in
seconds as measured by the UNIX date command.  For the approach using a
two-stage verification engine we apply a timeout of 8 hours per benchmark,
and 24 hours for the approach using a one-stage engine.

%---------------------------------
\subsection{Objectives}
\label{experimental-objectives}
%---------------------------------

Using the closed-loop control system benchmarks described in
Section~\ref{experimental-setup}, our experimental evaluation aims to answer
two research questions:
%
\begin{enumerate}

\item RQ1 \textbf{(performance)} does the CEGIS approach provide a feasible
solution in reasonable time considering robust stability of FWL digital
controllers?

\item RQ2 \textbf{(sanity check)} are the synthesised controllers sound
and can their stability be confirmed outside of our model?

\end{enumerate}

%---------------------------------
\subsection{Results}
\label{experimental-results}
%---------------------------------

We give the runtimes required to synthesise a stable controller for each
benchmark in Table~\ref{tab:results}.  Here, \textit{Plant} is the discrete
or continuous plant model, \textit{Benchmark} is the name of the employed
benchmark, \textit{I} and \textit{F} represent the number of integer and
fractional bits of the stable controller, respectively, while \textit{Time}
is the total time (in seconds) required to synthesise a stable controller
for the given plant.

For the majority of our benchmarks, the conjecture explained in
Section~\ref{synthesis-elements} holds and the two-stage verification 
engine %inductive back-end 
is able to find a stable solution in less than five minutes.  This is possible if the
inductive solutions need to be refined with few counterexamples and
fixed-point precision increases.  However, two benchmarks in our set
required too many counterexamples to refine their solution (marked with $^*$
in Table~\ref{tab:results}).  In these cases, the one-stage engine 
%interval arithmetic back-end 
is able to complement the two-stage approach and synthesise a solution
instead, albeit with significantly larger runtimes.  This performance
difference is due to the fact that the one-stage verification engine does
not take advantage of the inductive conjecture inherent to CEGIS, but
instead fully explores the counterexample space.  The result is a
performance difference of at least two orders of magnitude in our
experiments, leading to the one-stage engine timing out on the majority of
our benchmarks.  Table~\ref{tab:results} only lists the faster of the two
verification engines for each benchmark, which in $12$ out of $15$
benchmarks is the two-stage one.

\begin{table}
\centering
\begin{tabular}{c | c | l | r r | r l}
\# & Plant  & Benchmark                  & $I$ & $F$ & Time    & \\\cline{1-6}
1  & $G_1$  & CruiseControl02            &   4 &  16 & 17\,s   & \\
2  & $G_1$  & CruiseControl02$^\dagger$  &   4 &  16 & 46\,s   & \\
3  & $G_2$  & SpringMassDamper           &  15 &  16 & 28\,s   & \\
4  & $G_2$  & SpringMassDamper$^\dagger$ &  15 &  16 & \xmark  & \\
5  & $G_3$  & SatelliteB2                &   3 &   7 & 7\,s    & \\
6  & $G_3$  & SatelliteB2$^\dagger$      &   3 &   7 & 6601\,s & \tbmark\\
7  & $G_3$  & SatelliteC2                &   3 &   5 & 2\,s    & \\
8  & $G_3$  & SatelliteC2$^\dagger$      &   3 &   5 & 76\,s   & \tbmark\\
9  & $G_4a$ & a\_ST1\_IMPL1              &  16 &   4 & 2704\,s & \\
10 & $G_4a$ & a\_ST1\_IMPL2              &  16 &   8 & 538\,s  & \\
11 & $G_4a$ & a\_ST1\_IMPL3              &  16 &  12 & 12\,s   & \\
12 & $G_4b$ & a\_ST2\_IMPL1              &  16 &   4 & 318\,s  & \\
13 & $G_4b$ & a\_ST2\_IMPL2              &  16 &   8 & 967\,s  & \\
14 & $G_4b$ & a\_ST2\_IMPL3              &  16 &  12 & 9798\,s & \\
15 & $G_4c$ & a\_ST3\_IMPL1              &  16 &   4 & 6304\,s & \\
\end{tabular}

\vspace{0.5em}
\scriptsize{* = solved with the one-stage engine, $\dagger$ = with uncertainty}
\vspace{0.5em}

\caption{Experimental results obtained with \tool\label{tab:results}}
\end{table}
%---------------------------------

The median runtime for our benchmark set is $197$\,s, implying that \tool
can synthesise half of the controllers in less than $5$ minutes.  Overall,
the average synthesis time amounts to approximately $30$ minutes.  With
respect to RQ1 we consider these times short enough to be of practical use
to control engineers.  We further conclude that the approach
using a two-stage verification engine is able to synthesise stable
controllers in $12$ out of the $15$ benchmarks, and can be complemented
using the one-stage engine for $2$ additional benchmarks, where the
inductive conjectures fail.  Both verification engines together enabled
controller synthesis for $14$ out of $15$ benchmarks.  For the $1$ remaining
benchmark our approach failed to synthesise a stable controller in the
requested time limits.  This can be addressed by either increasing the fixed
point word widths considered or using floating point arithmetic instead. 
The synthesised controllers were confirmed to be stable outside of our
model representation using MATLAB, positively answering RQ2.
A~link to the full experimental environment, including scripts to reproduce
the results, all benchmarks and the \tool tool, is provided in the
footnote.\footnote{\url{https://drive.google.com/file/d/0ByIexo3Z5N91QkhrMi1rdkZFTWc/view}}

\subsection{Threats to validity}

We report a favourable assessment of \tool over a diverse set of real-world
benchmarks. Nevertheless, this set of benchmarks is limited within the
scope of this paper and \tool's performance may not generalise to other
benchmark scenarios. This can be addressed by expanding the benchmark set
in future experiments.

Furthermore, our algorithm to select suitable FWL word widths to model
plant behaviour employs a heuristic based on user-provided controller
word-width specifications.  Given the encouraging results of our benchmarks,
this heuristic appears to be strong enough for the current benchmark set,
but this may not generalise.  Further experiments towards determining
suitable plant FWL configurations may thus be necessary in future work.

Finally, the experimental results we obtained using \tool for stability
properties may not generalise to other properties.  The inductive nature of
the two-stage back-end of \tool increases performance significantly compared
to the one-stage back-end, but this performance benefit introduced by CEGIS'
inductive generalisations may not be observed for other controller
properties.  Additional experiments are necessary to confirm that the
performance of our inductive synthesis approach can be leveraged in those
scenarios as well.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Related Work}\label{sec:related}

\paragraph{Robust synthesis of Linear Systems}

The problem of parametric control synthesis based on stability measures for
continuous Linear Time Invariant (LTI) Single Input-Single Output (SISO)
systems has been researched for several decades.  On a theoretical level it
is a solved problem~\cite{wonham1967pole} for which researches continuously
seek better results based on a number of properties in addition to
stability.  A vast range of pole placement techniques such as Moore's
algorithm for eigenstructure assignment~\cite{klein1977eigenvalue} or the
more recent Linear Quadratic Regulator (LQR)~\cite{bemporad2002explicit}
have been used in several studies with increasing degrees of success to
ensure stability.  The latter approach highlights the importance of
conserving energy during the control process, which results in lower running
costs.  Since real systems are subject to tolerance and noise as well as the
need for economy, more recent studies focus on the problems of achieving
robust stability with minimum gain~\cite{schmid2014unified,
konigorski2012pole}.  However, when extending the problem to digital
controller synthesis, many of these techniques lack the ability to produce
sound or stable results because they disregard the effects of quantisation
and rounding.  Recent papers on implementations/synthesis of LTI digital
controllers~\cite{das2013lqr,ghosh2013fpga} focus on time discretisation,
failing to account for these error-inducing effects and can therefore result
in digitally unstable systems that have been proven to be robustly stable in
a continuous space.  This situation demands the inclusion of formal methods
to verify the effects of space discretisation on the implementation of the
controller.

\paragraph{Formal Verification of Linear Digital Controllers}

Researchers in formal verification of control systems have studied various
effects of discretising the dynamics, including delayed
response~\cite{Duggirala2015}, and Finite Word Length (FWL) constraints with
the goal to either verify (e.g.~\cite{daes20161}) or optimise
(e.g.~\cite{oudjida2014design}) implementations.  There are in fact two
different problems arising from FWL effects.  The first corresponds to the
error in the dynamics caused by the inability of the hardware to represent
the exact dynamics of the system whilst the second relates to rounding
errors during computation.  In~\cite{fialho1994stability} a stability
measure based on the error of the digital dynamics ensures that the FWL
errors do not make the digital system unstable.  A more recent
approach~\cite{DBLP:journals/automatica/WuLCC09} uses $\mu$ calculus to
directly model the digital controller so that the selected parameters are
stable by design.  Most studies following this verification research field
focus on finding a correct implementation of a known desired controller,
looking for optimal parameter representations using FWL, but ignoring the
effects of rounding errors due to issues of mathematical tractability.  The
analyses in~\cite{DBLP:conf/hybrid/WangGRJF16, DBLP:conf/hybrid/RouxJG15}
rely on an invariant computation on the discrete system dynamics using
Semi-Definite Programming (SDP).  Whilst the former uses BIBO properties to
determine stability, the latter uses Lyapunov-based quadratic invariants. 
In both cases, the SDP solver uses floating-point arithmetic and soundness
is checked by upper bounding the error.  An alternative approach is taken
by~\cite{park2016scalable}, where the verification of existing code is
performed against a known model by extracting an LTI model of the code
through symbolic execution.  In order to account for rounding errors, an
upper bound is introduced in the verification phase.  If the error of the
implementation is lower than this tolerance level, then the verification is
successful.

\paragraph{Robust Synthesis of FWL Digital Controllers}

Given the current knowledge in control system's synthesis, there is no other
existing study on the automatic synthesis of fixed-point digital controllers
considering FWL effects.  Parameter synthesis tools such
as~\cite{cimatti2013parameter} use bounded model checking and fixed-point
computations to find parameters based on user defined specifications, but
they often operate in the continuous domain and are ill-suited for robust
analysis since they have no direct means of evaluating robustness.  Other
tools such as~\cite{economakos2016automated} are directly aimed at robust
stability problems, but they fail to explore the effects of FWL in their
implementation.  In order to provide a correct by design digital controller,
\cite{alur2016compositional} requires a user-defined finite state
abstraction to synthesise a digital controller based on high level
specifications.  Whilst this approach directly overcomes the challenges
presented by the FWL problem, it still requires error-prone user
intervention.  A different solution that looks at the other end of the scale
is an approach that synthesises word lengths for known problems as developed
in~\cite{jha2013swati}; however, this provides neither an optimal word
synthesis nor a comprehensive solution for the problem.  It is from this
vacuum that we seek to explore a CEGIS approach to FWL controller synthesis.

\paragraph{CEGIS}

In recent years the development of program synthesis has found many ways to
obtain correct by design programs from high-level specifications.  One such
approach~\cite{itzhaky2010simple} looks to inductively synthesise invariants
to generate the desired programs.  Since these digital programs are already
constrained by designed to factors such as FWL, these approaches are ideal
for the synthesis of parametric controllers. 
In~\cite{DBLP:conf/cdc/RavanbakhshS15}, the authors use CEGIS for the
synthesis of switching controllers for stabilizing continuous-time plants
with polynomial dynamics.  The work extends to its application on affine
systems, finding its major challenge in the hardness of solving linear
arithmetic on present SMT solvers.  Since this approach uses switching
states instead of linear dynamics in the digital controller, it entirely
circumvents the FWL problem.  It is also not suitable for the kind of
control we seek to synthesise.  What is needed in this regard is the
combination of a synthesis engine with a control verification tool that
addresses the challenges presented here in the form of FWL effects and
stability measures for LTI SISO controllers.  We take the former
from~\cite{DBLP:conf/lpar/DavidKL15} and the latter from~\cite{daes20161}
whilst enhancing the procedure by evaluating the quantisation effects of
the Hardware interfaces (ADC/DAC) to obtain an accurate discrete-time FWL
representation of the continuous dynamics.

%In \cite{DBLP:journals/corr/PakmehrWJVF13}, Pakmehr et al.  design a control software verification framework for gas turbine engines.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Conclusions} \label{sec:conclusions}

We have presented a CEGIS-based approach implemented in a tool called \tool, 
which uses inductive synthesis in conjunction with an algorithm for verifying 
robust closed-loop stability. In particular, \tool is able to compute a parametric solution 
for a stabilizing digital controller given a continuous (or discrete) plant 
specification.  This enables us to automatically synthesise correct-by-construction 
digital controllers for a given model of their physical environment.  
Our approach addresses the effects of the quantisers (A/D and D/A converters) 
as time discretisation elements with quantisation noise.  Our experimental results 
show that \tool is able to synthesise stable controllers in most benchmarks within
reasonable time frames fully automatically.
%
% and can be 
%complemented using the interval arithmetic back-end, where the inductive conjectures fail. 
%Both back-ends enabled controller synthesis for all benchmarks within reasonable time frames.
%
Future work will be the extension of this CEGIS-based approach to further
classes of systems, including state-space representation.  
We will also consider performance requirements while
synthesising the digital controller.

\bibliographystyle{abbrv}
\bibliography{paper}

%APPENDICES are optional
%\balancecolumns
%\appendix
%Appendix A

\end{document}