### 2017-05-28 emsoft2017/emsoft-paper.tex

\documentclass[sigconf]{acmart}

\usepackage{xspace,framed}

\setcopyright{acmcopyright}
\acmDOI{10.475/123_4}
\acmISBN{123-4567-24-567/08/06}
\acmConference[EMSOFT '17]{The ACM SIGBED International Conference on Embedded Software }{October 15--20, 2017}{Seoul, South Korea}
\acmYear{2017}
\copyrightyear{2017}
\acmPrice{15.00}

\newcommand\tool{{DSSynth Toolbox}\xspace}

\begin{document}

\title{DSSynth Toolbox: Automated Synthesis of Digital Controllers \\ for Physical Plants via MATLAB Toolbox}

\begin{abstract}
We present an automated approach implemented as a MATLAB Toolbox to synthesize stable and safe digital-control systems for physical plants represented as linear and time invariant models. The tool developed in the current paper is named as DSSynth (Digital-Systems Synthesis), which is a MATLAB Toolbox that synthesizes digital controllers w.r.t. the stability specification, considering finite word-length (FWL) effects for physical plants represented by transfer-function or state-space equations. DSSynth considers fixed-point numerical representation for the digital controller and dynamic input ranges for the closed-loop control system. Currently, there are two versions of the DSSynth toolbox: command-line and MATLAB application.
\end{abstract}

%
% The code below should be generated by the tool at
% http://dl.acm.org/ccs.cfm
% Please copy and paste the code instead of the example below. 
%
\begin{CCSXML}
<ccs2012>
<concept>
<concept_id>10010520.10010570</concept_id>
<concept_desc>Computer systems organization~Real-time systems</concept_desc>
<concept_significance>500</concept_significance>
</concept>
<concept>
<concept_id>10010520.10010553.10010562</concept_id>
<concept_desc>Computer systems organization~Embedded systems</concept_desc>
<concept_significance>300</concept_significance>
</concept>
<concept>
<concept_id>10011007.10010940.10010992.10010998.10003791</concept_id>
<concept_desc>Software and its engineering~Model checking</concept_desc>
<concept_significance>500</concept_significance>
</concept>
<concept>
<concept_id>10011007.10010940.10010992.10010998</concept_id>
<concept_desc>Software and its engineering~Formal methods</concept_desc>
<concept_significance>100</concept_significance>
</concept>
<concept>
<concept_id>10003752.10003790.10011192</concept_id>
<concept_desc>Theory of computation~Verification by model checking</concept_desc>
<concept_significance>300</concept_significance>
</concept>
</ccs2012>
\end{CCSXML}

\ccsdesc[500]{Computer systems organization~Real-time systems}
\ccsdesc[300]{Computer systems organization~Embedded systems}
\ccsdesc[500]{Software and its engineering~Model checking}
\ccsdesc[100]{Software and its engineering~Formal methods}
\ccsdesc[300]{Theory of computation~Verification by model checking}

\keywords{Formal Synthesis, Digital Controllers, MATLAB Toolbox, Fixed-point Arithmetic, Transfer-Function, State-Space, Digital Control Synthesis}

\maketitle

\input{paper-body}

\bibliographystyle{ACM-Reference-Format}
\bibliography{references}

\end{document}