### 2017-05-31 emsoft2017/emsoft-paper.tex
reviewed section 2.1
Signed-off-by: Lucas Cordeiro <lucasccordeiro@gmail.com>

\documentclass[sigconf]{acmart}

\usepackage{xspace,framed}
\usepackage{listings}

\setcopyright{acmcopyright}
\acmDOI{10.475/123_4}
\acmISBN{123-4567-24-567/08/06}
\acmConference[EMSOFT '17]{The ACM SIGBED International Conference on Embedded Software }{October 15--20, 2017}{Seoul, South Korea}
\acmYear{2017}
\copyrightyear{2017}
\acmPrice{15.00}

\newcommand\tool{{DSSynth Toolbox}\xspace}

\begin{document}

\title{DSSynth: An Automated Synthesis Tool for Physical Plants \\ Digital Controllers via MATLAB Toolbox}

\begin{abstract}
We present an automated MATLAB Toolbox, named as DSSynth 
(Digital System Synthesis), to synthesize stable and safe digital-control 
systems for physical plants represented as linear time-invariant
systems with single-input and single-output. DSSynth synthesizes digital 
controllers w.r.t. the stability and safety specifications, considering finite word-length 
effects for physical plants represented by transfer-function or state-space 
equations. DSSynth considers fixed-point numerical representation for the 
digital controller and dynamic input ranges for the closed-loop control system. 
Currently, there are two DSSynth toolbox versions: command-line and 
MATLAB application. The resulting toolbox enables application of synthesis to 
real-world systems by control engineers.
\end{abstract}

%
% The code below should be generated by the tool at
% http://dl.acm.org/ccs.cfm
% Please copy and paste the code instead of the example below. 
%
\begin{CCSXML}
<ccs2012>
<concept>
<concept_id>10010520.10010570</concept_id>
<concept_desc>Computer systems organization~Real-time systems</concept_desc>
<concept_significance>500</concept_significance>
</concept>
<concept>
<concept_id>10010520.10010553.10010562</concept_id>
<concept_desc>Computer systems organization~Embedded systems</concept_desc>
<concept_significance>300</concept_significance>
</concept>
<concept>
<concept_id>10011007.10010940.10010992.10010998.10003791</concept_id>
<concept_desc>Software and its engineering~Model checking</concept_desc>
<concept_significance>500</concept_significance>
</concept>
<concept>
<concept_id>10011007.10010940.10010992.10010998</concept_id>
<concept_desc>Software and its engineering~Formal methods</concept_desc>
<concept_significance>100</concept_significance>
</concept>
<concept>
<concept_id>10003752.10003790.10011192</concept_id>
<concept_desc>Theory of computation~Verification by model checking</concept_desc>
<concept_significance>300</concept_significance>
</concept>
</ccs2012>
\end{CCSXML}

\ccsdesc[500]{Computer systems organization~Real-time systems}
\ccsdesc[300]{Computer systems organization~Embedded systems}
\ccsdesc[500]{Software and its engineering~Model checking}
\ccsdesc[100]{Software and its engineering~Formal methods}
\ccsdesc[300]{Theory of computation~Verification by model checking}

\keywords{Formal Synthesis; Digital Control Systems; MATLAB Toolbox; Finite-Word Length; Verification}

\maketitle

%---------------------------------------------------
\section{Introduction}
%---------------------------------------------------

Linear-Time Invariant (LTI) systems are used in several applications
to support non-trivial tasks when applied to embedded systems 
(e.g., robot manipulator, power plants and guidance system). 
In addition, controllers synthesis for LTI systems is richly studied 
by different researchers~\cite{mazo2010pessoa,DBLP:conf/emsoft/RavanbakhshS16,economakos2016automated}; 
however, there are many challenges when 
digital-control systems theory is used due to the finite-word 
length (FWL) effects~\cite{Guang2013, Istepanian2001}, time discretization 
and quantization noise, which are typically introduced by Analogue-to-Digital 
(ADC) and Digital-to-Analogue (DAC) conversion, leading to truncation 
or round-off errors~\cite{astrom1997computer}.

Embedded digital control systems perform computations over digital signals 
to influence the system behavior~\cite{Ogata2001}; it can be mathematically 
expressed as difference equations, transfer functions or state-space equations. 
Here, the focus is on transfer-function and state-space representations. 
The following expression presents the general form of a digital control system 
using the transfer function representation:
%
\begin{equation}
\small
\label{eq:transferfunction}
H(z)=\frac{B(z)}{A(z)}=\frac{b_{0}+b_{1}z^{-1}+...+b_{M}z^{-M}}{1+a_{1}z^{-1}+...+a_{N}z^{-N}},
\end{equation}
%
\noindent where $z^{-1}$ is called backward-shift operator, $A(z)$ and $B(z)$ are 
the denominator and numerator polynomials, respectively.
%
In state-space format, the digital control system represents the behavior 
of a system via a state evolution equation $\dot{x}(n+1)$ and an instantaneous 
output equation $y(n)$, as follows:
%
\begin{equation}
\begin{split}
\dot{x}(n+1) &= A x(n) + B u(n)
\\
y(n) &= C x(n) + D u(n), 
\end{split}\label{eq:ss-example}
\end{equation}

\noindent where $A$, $B$, $C$ and $D$ are matrices that fully specify a digital system.

Recently, a new methodology to synthesize digital-control systems was proposed, 
named as DSSynth (Digital System Synthesis)~\cite{abate2017, abatecav2017}, 
which is based on the Counter-Example Guided Inductive Synthesis 
(CEGIS) approach~\cite{DBLP:conf/asplos/Solar-LezamaTBSS06}. Here, CEGIS is used
as a program synthesis engine able to produce controllers for highly non-trivial 
specifications with a very high degree of automation. Program synthesis engines 
as CEGIS use a specification as the starting point, and subsequently 
produce a sequence of candidate programs from a given template. The candidate programs 
are iteratively refined to eventually satisfy the specification. Modern synthesis engines combine 
automated testing, genetic algorithms and SMT-based automated reasoning~\cite{DBLP:journals/corr/AlurFSS16a, DBLP:conf/lpar/DavidKL15}. 
Currently, Abate {\it et al.}~\cite{abate2017} synthesize stable, software-implemented 
embedded controllers, along with a physical plant model represented by transfer functions 
or state-space equations~\cite{abate2017,abatecav2017} using a CEGIS-based approach.

Notably, there are toolboxes in MATLAB with functions and scripts to facilitate the 
digital system design and implementation~\cite{matlab-toolbox}. In fact, there is a MATLAB 
Toolbox named as DSVerifier Toolbox~\cite{issta2017}, which automatically detects 
specific errors related to digital system design ({\it e.g.}, stability, limit cycle and overflow) 
using symbolic model checking techniques based on Boolean Satisfiability (SAT) and 
Satisfiability Modulo Theories (SMT) solvers, given the physical plant and the digital controller. 
However, there is no toolbox in MATLAB to perform digital control synthesis; in particular, based on 
CEGIS considering FWL effects.

The present paper addresses this problem and describes a MATLAB toolbox for DSSynth~\cite{abate2017, abatecav2017} 
in MATLAB's environment named as \tool. The main advantage regarding the use of a MATLAB toolbox 
lies on designing physical plants in MATLAB and then promptly synthesizing a stable digital controller. 
Additionally, when using the \tool, an engineer is able to design a digital system with MATLAB, 
through transfer-function or state-space representations, consider low-level systems parameters 
(implementation features and numerical format), and then gets a stable synthesized digital-controller.

%---------------------------------------------------
\section{Synthesizing Digital Controllers with \tool}
%---------------------------------------------------

%---------------------------------------------------
\subsection{\tool Architecture}
%---------------------------------------------------

The proposed synthesis methodology for closed-loop digital control
systems is based on the DSSynth tool~\cite{abate2017, abatecav2017}; 
it can be split into two main stages as follows: manual (user) and 
automated (DSSynth) procedures, as illustrated in Fig.~\ref{fig:synthesis-flow}. 
%
\begin{figure}[ht!]
\centering
\includegraphics[width=0.45\textwidth]{synthesis-flow.pdf}
\caption{Digital Controller Synthesis Methodology.}
\label{fig:synthesis-flow}
\end{figure}

In step $1$, the digital-control representation is define, which can be in 
transfer-function or state-space equations. In step $2$, the physical plant 
with the respective coefficients must be designed, according to equations 
(\ref{eq:transferfunction}) and (\ref{eq:ss-example}). And finally, in step $3$, 
the numerical representation for the digital controller implementation must be 
defined, {\it i.e.}, the FWL format that includes the number of bits for the integer 
and fractional parts and the dynamic range inputs. The results from the manual 
procedure is a digital-system implementation for the given physical plant in MATLAB.

The synthesis process starts in step $A$, when the \tool gets the digital-system in MATLAB, 
and then translates it into a ANSI-C program; then in step $B$, the DSSynth Tool is invoked. 
Finally, in step $C$, the digital controller is synthesized based on CEGIS. The output generated 
in MATLAB is the synthesized digital controller, which can be represented in transfer-function 
or state-space format. The synthesis is \emph{successful} if a digital controller is correctly synthesized;
otherwise the synthesis has status \emph{failed} if any parameter is defined incorrectly 
in previous steps, or there is no solution for the designed digital controller.

%---------------------------------------------------
\subsection{\tool Procedures}
%---------------------------------------------------

The \tool performs the following automated procedures in order to synthesize a digital controller:

\begin{enumerate}
\item \textbf{Setup}: obtain the plant, fixed-point format and dynamic input ranges, and translate them to a struct in MATLAB;
\item \textbf{Parse}: obtain the digital plant implementation, and then translate it to an ANSI-C file;
\item \textbf{Execute}: obtain the ANSI-C file from the previous step, call CBMC-CEGIS as back-end program synthesis tool, and then perform the automated synthesis;
\item \textbf{Extract}: obtain the .log that is generated after the synthesis phase and then check the synthesized digital-controller.
\item \textbf{Report}: obtain the digital-controller, translate it to a MATLAB system, and then show the result to the user.
\end{enumerate}

%---------------------------------------------------
\subsection{\tool Features}
%-----------------------------------------------

The \tool's features can be described as follows:

\begin{itemize}
\item \textbf{Representation}: the synthesis could be performed for digital-systems in transfer-function or state-space equations;
\item \textbf{Fixed-Point Arithmetic}: the digital controller is synthesized considering FWL effects, in fixed-point numeric representation;
\item \textbf{Interface}: the synthesis could be executed using the command-line version, and a GUI Application version. In addition, in the GUI Application, the user is able to plot the step response in order to reproduce the stability of the digital controller synthesized;
\end{itemize}
%---------------------------------------------------
\subsection{\tool Usage}
%---------------------------------------------------

%---------------------------------------------------
\subsubsection{Command Line Version}
%---------------------------------------------------

Users must provide a digital system described as a MATLAB system  using a \texttt{``tf''} (for transfer-function) or an \texttt{``ss''} (for state-space) command (cf. step $1$ of Fig.~\ref{fig:synthesis-flow}).

\tool is called via command line in MATLAB as \texttt{synthesize(plant, intBits, fracBits, maxRange, minRange)}, where \texttt{plant} is the physical plant in state-space or transfer-function representation, \texttt{intBits} is the integer part, \texttt{fracBits} is the fractional part, \texttt{maxRange} and \texttt{minRange} are the maximum and minimum dynamic range, respectively.

After executing \texttt{synthesize} command in MATLAB, the output prints the synthesized digital controller according to the digital system representation, {\it i.e.}, transfer-function or state-space equations.

%---------------------------------------------------
\subsubsection{MATLAB Application Version} 
%---------------------------------------------------

A graphical user interface application was developed (Fig.~\ref{fig:gui-for-tf}), in order to favor digital-control systems synthesis in MATLAB, besides improving usability and, consequently, attracting more digital-system engineers. Users can provide all required parameters for digital-system synthesization: physical plant specification, fixed-point implementation and dynamic inputs. 
%
\begin{figure}
  \includegraphics[width=0.5\textwidth]{screens_dssynth.png}
  \caption{GUI Application for Transfer-Function Synthesis.}
  \label{fig:gui-for-tf}
\end{figure}

%---------------------------------------------------
\subsection{Illustrative Example}
%---------------------------------------------------

To illustrate the \tool's usage, Fig.~\ref{toolbox-usage} shows the synthesis for a physical plant defined in Eq.(~\ref{equation_plant}) using the transfer-function representation, where ``num'' and ``den'' represent the numerator $A(z)$ and denominator $B(z)$, respectively.
%
\begin{equation}
\label{equation_plant}
H(z)=\frac{B(z)}{A(z)}=\frac{-0.06875z^{2}}{z^2-1.696z+0.7089},
\end{equation}
%
\begin{figure}[ht]
\scriptsize
\begin{lstlisting}[xleftmargin=.025\textwidth,xrightmargin=.025\textwidth, frame=single,]
>> num = [-0.06875 0 0];
>> den = [1.0000 -1.696 0.7089];
>> system = tf(num,den,0.002);
>> y = synthesize(system,8,8,1,-1);
>> SYNTHESIS SUCCESSFUL
>> y = (-0.9983z^2 + 0.09587z + 0.1926)/(z^2+0.5665+0.75);
\end{lstlisting}
\vspace{-0.2cm}
\caption{Synthesis of a Digital Controller for the physical plant defined in Eq.(~\ref{equation_plant}) in MATLAB, with a fixed-point format  $\left\langle 8,8\right\rangle$.}
\label{toolbox-usage}
\end{figure}
%
In the Fig.~\ref{toolbox-usage}, the value returned in $y$ represents the digital controller synthesized and stable. In order to validate and reproduce the stability in the digital system composed by the physical plant defined in Eq.(~\ref{equation_plant}) and the digital controller represented by $y$, the user can invoke the step response using the command \texttt{dstep}, and then observe that the graph plotted shows a stable digital system, as can be seen in Fig.~\ref{step-response}.
%
\begin{figure}[ht]
  \includegraphics[width=0.5\textwidth]{step-response.eps}
  \caption{Step response for the Eq.~\eqref{equation_plant} describing a stable system.}
  \label{step-response}
\end{figure}

\section{Conclusions}

\tool is able to synthesize dynamic physical plants designed in MATLAB, through transfer-function and state-space representations based on Counter-Example Guided Inductive Synthesis (CEGIS)~\cite{DBLP:conf/asplos/Solar-LezamaTBSS06}.

Given the current literature in formal synthesis, there is no other MATLAB toolbox for synthesize stable digital systems, while taking into account implementation aspects (fixed-point arithmetic).

As future work, \tool will perform synthesis to generate stable digital controllers and be combined with DSVerifier Toolbox~\cite{issta2017} to verify the stability of the digital controller synthesized, and also, with DSValidator Toolbox~\cite{dsvalidator}, in order to reproduce and validate the stability property.

\bibliographystyle{ACM-Reference-Format}
\bibliography{references}

\end{document}