### 2017-07-08 ase2017/paper.tex

\documentclass[10pt,conference]{IEEEtran}

\usepackage{microtype}
\usepackage{xspace,framed}
\usepackage{listings}
\usepackage{graphicx}
\usepackage{epstopdf} %converting to PDF
\usepackage{subfig} 
\usepackage{xcolor,colortbl}
\usepackage{amssymb}
\usepackage{amsmath}
\usepackage{hyperref}
\usepackage{cite}	
\usepackage{mathtools}
\usepackage{tikz}
\usetikzlibrary{positioning, automata, shapes.arrows, calc, shapes, arrows}
\usetikzlibrary{patterns}
\usepackage{pgfplots}

%\setcopyright{acmcopyright}
%\acmDOI{10.475/123_4}
%\acmISBN{123-4567-24-567/08/06}
%\acmConference[ASE '17]{The ACM SIGBED International Conference on Embedded Software }{October 15--20, 2017}{Seoul, South Korea}
%\acmYear{2017}
%\copyrightyear{2017}
%\acmPrice{15.00}

\renewcommand{\baselinestretch}{0.99}
\newcommand\tool{{DSSynth Toolbox}\xspace}
\newcommand{\fwl}[1]{\mathcal{FWL}[#1]}

\begin{document}

\title{DSSynth: An Automated Digital Controller Synthesis Tool for Physical Plants} 
% via MATLAB Toolbox}

\author{\IEEEauthorblockN{Alessandro Abate$^{1}$, Iury Bessa$^{2}$, Dario Cattaruzza$^{1}$, Lennon Chaves$^{2}$, Lucas Cordeiro$^{1,2}$, \\ 
Cristina David$^{1}$, Pascal Kesseli$^{1}$, Daniel Kroening$^{1}$ and Elizabeth Polgreen$^{1}$}
\IEEEauthorblockA{$^{1}$University of Oxford, Oxford, United Kingdom \\
$^{2}$Federal University of Amazonas, Manaus, Brazil}}

%\author{\IEEEauthorblockN{Erickson H. da S. Alves, Lucas C. Cordeiro, Eddie B. de Lima Filho}
%		\IEEEauthorblockA{Federal University of Amazonas, Brazil\\
%			E-mails: erickson.alves@indt.org.br, lucascordeiro@ufam.edu.br, eddie@ctpim.org.br}}

\maketitle

\begin{abstract}
%
We present an automated MATLAB Toolbox, named \emph{DSSynth} (Digital-System
Synthesizer), to synthesize sound digital controllers for physical plants
that are represented as linear time-invariant systems with single input and
output.  In~particular, DSSynth synthesizes digital controllers that are
sound w.r.t.~stability and safety specifications.  DSSynth considers the
complete range of approximations, including time discretization,
quantization effects and finite-precision arithmetic (and its rounding
errors).  We demonstrate the practical value of
this toolbox by automatically synthesizing stable and safe controllers for intricate
physical plant models from the digital control literature.
The resulting toolbox enables the application of program synthesis
to real-world control engineering problems.
%
\end{abstract}

%
% The code below should be generated by the tool at
% http://dl.acm.org/ccs.cfm
% Please copy and paste the code instead of the example below. 
%
\begin{IEEEkeywords}
Formal Synthesis; Digital Control Systems; MATLAB Toolbox; Finite-Word Length; Verification
\end{IEEEkeywords}

%---------------------------------------------------
\section{Introduction}
%---------------------------------------------------

Control theory targets the construction of reliable systems by providing
mathematical guarantees about the stability and the desired performance of
closed-loop dynamics, where outputs of discrete plant $G(z)$ are fed back
and compared to a reference signal towards which a controller $C(z)$ should
steer~\cite{astrom1997computer}.  Fig.~\ref{fig:typical-control-system}
depicts a typical closed-loop digital control system.
%
\begin{figure}[ht!]
\centering
\includegraphics[width=0.4\textwidth]{closedloopseries.pdf}
\caption{Typical Closed-loop Control System}
\label{fig:typical-control-system}
\end{figure}

%Embedded digital control systems perform computations over digital signals 
%to influence the system behavior~\cite{Ogata2001}; it can be mathematically 
%expressed as difference equations, transfer functions or state-space equations. 
%Here, the focus is on transfer function and state-space representations.

There are various ways to model control systems.  Two common ways to
represent the dynamics of such systems are transfer functions and
state-space equations (difference or differential equations).  In this
paper, we consider systems with a single input and a single output (SISO)
given either as transfer functions or state-space linear time-invariant
(LTI) models.  LTI models are a common sub-class, and are frequently used in
cyber-physical systems (e.g., in robotics, smart grids, power plants, and
avionics).  A discretized plant or a digital controller is, in generally,
given as follows when using a transfer function representation:
%
\begin{equation}
\small
\label{eq:transferfunction}
H(z)=\frac{B(z)}{A(z)}=\frac{b_{0}+b_{1}z^{-1}+...+b_{M}z^{-M}}{a_{0}+a_{1}z^{-1}+...+a_{N}z^{-N}},
\end{equation}
%
In this representation, $z^{-1}$ is called the backward-shift operator;
$A(z)$ and $B(z)$ are the denominator and numerator polynomials; and $N$ and
$M$ are the order of the denominator and numerator polynomials, respectively.

In a state-space form, the behavior of a system is represented via a state
evolution equation $x(n+1)$ and an instantaneous output equation $y(n)$, as
follows:
%
\begin{equation}
\begin{split}
x(n+1) &= A x(n) + B u(n)
\\
y(n) &= C x(n) + D u(n), 
\end{split}\label{eq:ss-example}
\end{equation}
%
where $A$, $B$, $C$ and $D$ are matrices that fully specify the model.

There is a rich body of literature on automating the synthesis of
controllers for LTI systems~\cite{mazo2010pessoa,
DBLP:conf/emsoft/RavanbakhshS16, economakos2016automated}.  However, there
are many challenges when digital control is used to implement the
controller, mainly due to effects of finite-word length (FWL), which leads
to truncation or round-off errors~\cite{Guang2013, Istepanian2001}, and time
discretization and quantization noise, which are typically introduced by
Analogue-to-Digital (ADC) and Digital-to-Analogue (DAC)
conversion~\cite{astrom1997computer}.

We have recently proposed a new method to synthesize digital control systems
for LTI systems, which we have named \emph{DSSynth} (Digital-System
Synthesizer)~\cite{abate2017, abatecav2017}.  The algorithm implements
Counter-Example Guided Inductive Synthesis
(CEGIS)~\cite{DBLP:conf/asplos/Solar-LezamaTBSS06}.  CEGIS is used to
realize a program synthesis engine that is able to generate sound digital
controllers for highly non-trivial system specifications with a very high
degree of automation.  The program synthesis engine is given a specification
and subsequently produces a sequence of candidate programs from a template. 
The candidate programs are iteratively refined to eventually satisfy the
specification.  Modern synthesis engines such as the one used here combine
automated testing, genetic algorithms and SMT-based
reasoning~\cite{DBLP:journals/corr/AlurFSS16a, DBLP:conf/lpar/DavidKL15}. 
Using the CEGIS-based approach, DSSynth synthesizes stable digital
controllers for a given physical plant represented by either transfer
functions or state-space equations.

There are existing toolboxes for MATLAB that provide functions and scripts
to support design and implementation of digital
controllers~\cite{matlab-toolbox}.  In~particular, there is a MATLAB Toolbox
named \emph{DSVerifier}~\cite{issta2017, DBLP:journals/tc/BessaIPCF17,
DBLP:journals/dafes/BessaICF16}, which automatically detects specific errors
in a given digital system design (e.g., errors related to stability, limit
cycle and overflow) using symbolic model checking based on Boolean
Satisfiability (SAT) and Satisfiability Modulo Theories (SMT) solvers.  The
Pessoa~\cite{mazo2010pessoa} toolbox performs synthesis of correct-by-design
digital controllers.  Pessoa is based on approximate bisimulation, which
allows control engineers to replace differential equations, describing a
physical plant, by a respective finite-state machine.  However, Pessoa does
not take advantage of recent advances in bit-accurate verification of
programs to synthesize controllers w.r.t.~FWL effects.  MATLAB offers a tool
named \emph{Control System Tuner}~\cite{autotuner} for automatically
designing controllers based on optimization or graphical analysis.  However,
the Control System Tuner is unable to consider FWL effects during controller
synthesis.  To the best of our knowledge, there is no MATLAB toolbox able to
perform digital controller synthesis that considers the complete range of
approximations, including time discretization, quantization effects, and
finite-precision arithmetic and its rounding errors.

The present tool paper addresses this limitation and describes a MATLAB
toolbox for DSSynth.  MATLAB is very commonly used for modeling physical
plants.  Thus, the integration into MATLAB enables rapid synthesis of a
stable and safe digital controller for such plants without switching the
environment. Experimental results show that \tool is able to efficiently 
synthesize stable and safe controllers for a set of intricate benchmarks taken 
from the literature:  the median runtime for our benchmarks set is $124$\,s, 
and the average synthesis time amounts to $35.5$\,s for closed-loop control 
systems represented by state-space equations, while for transfer functions, the average synthesis
time amounts to $123.6$\,s.

%---------------------------------------------------
\section{Counterexample-Guided Inductive Synthesis for Control Systems}
\label{sec:CEGIS}
%---------------------------------------------------

%\textcolor{red}{Cristina: can you please generalize our CEGIS implementation for control here?}

%% The input specification provided to the program synthesizer is of the form
%% $\exists P .\, \forall a.\, \sigma(a, P)$, where $P$ ranges over functions
%% (where a function is represented by the program computing it),
%% $a$ ranges over ground terms, and $\sigma$ is a quantifier-free formula. 
%% We~interpret the ground terms over some finite domain~$\mathcal{D}$.

Our synthesizer iterates between two main phases, an inductive synthesis
phase, {\sc synthesise}, and a validation phase, {\sc verify}.  The two
phases interact via a finite set of test vectors, which is updated
incrementally.  Given a stability and safety specification, the inductive
synthesis procedure tries to find a candidate solution satisfying the
specification for the given set of test inputs.
%
If the synthesis phase succeeds in finding a witness, this witness is a
candidate solution to the full synthesis formula.  We pass this candidate
solution to the validation phase, which checks whether it is a full solution
(i.e., satisfies the specification for all possible inputs).  If this is the
case, then the algorithm terminates.  Otherwise, additional information is
provided to the inductive synthesis phase in the form of a new
counterexample, C-ex, that is added to the set of test inputs, and the loop
iterates again.

We use two different instantiations of the synthesiser
in order to find stable and safe digital
controllers. The two approaches differ in
the manner in which they check the safety specification
for all reachable states.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{figure}[htb]
{\scriptsize
\centering
\resizebox{.5\textwidth}{!}
{
  \begin{tikzpicture}[scale=0.3,->,>=stealth',shorten >=.2pt,auto, semithick, initial text=, ampersand replacement=\&,]
  \matrix[nodes={draw, fill=none, shape=rectangle, minimum height=.2cm, minimum width=.2cm, align=center
},
          row sep=.6cm, column sep=.9cm] {
   \coordinate (aux1);
   \& \coordinate (aux2);
   \&;\\
   \coordinate (aux3);
   \& \coordinate (aux4);
   \&;\\
   \coordinate (aux5);
   \& \coordinate (aux6);
   \&;\\
   \node[minimum width=1.5cm, minimum height=0.6cm, fill=gray!20] (synth) {{\sc Synthesize}};
   \&
   complexnode/.pic={ 
     \node[rectangle,draw,dotted,
	minimum width=6cm,
	minimum height=1cm,
        pattern=north west lines, pattern color=gray!20,
	label={\sc ~~~~~~~~~~~~Verify},] (verif) {};
     \node[minimum width=1cm, minimum height=0.6cm, fill=gray!20] (verif1) at ([xshift=-2cm]verif.center) {{\sc 1.Safety}};
     \node[minimum width=1cm, minimum height=0.6cm, fill=gray!20] (verif2) at ([xshift=0cm]verif.center) {{\sc 2.Precision}};
     \node[minimum width=1cm, minimum height=0.6cm, fill=gray!20] (verif3) at ([xshift=2cm]verif.center) {{\sc 3.Complete}};
     %\node[minimum width=1cm, minimum height=0.6cm, fill=gray!20] (verif4) at ([xshift=3.1cm]verif.center) {{\sc 5.Sampling}};
   } 
   \& \node[ellipse, fill=gray!20] (done) {{\sc Done}};\\
   \& \\
   \node[minimum height=0cm] (gp) {\sf Program Search};
   \&
   complexnode/.pic={ 
     \coordinate (aux);
   \node (bmc) at ([xshift=-2cm]aux.center) {\sf BMC-based \\ \sf Verifier};
   \node (fp)  at ([xshift=0cm]aux.center) {\sf Fixed-point \\ \sf Arithmetic\\ \sf Verifier};
   \node (sv)  at ([xshift=2cm]aux.center) {\sf Completeness\\ \sf Verifier};
   %\node (cv)  at ([xshift=3.1cm]aux.center) {\sf Sampling\\Verifier};
   }   
    \\
  };

   \path
    ([yshift=2em]synth.east) edge node[xshift=-0.5em,align=center] {K} ([yshift=2em]verif1.west)
    ([yshift=-2em]verif1.west) edge node {C-ex} ([yshift=-2em]synth.east)
    ([xshift=-5em]fp.north) edge node[align=center]  {} ([xshift=-5em]verif2.south)
    ([xshift=-5em]sv.north) edge node[align=center]  {} ([xshift=-5em]verif3.south)
    %([xshift=-5em]cv.north) edge node[align=center]  {T/F} ([xshift=-5em]verif4.south)
    ([xshift=5em]verif1.south) edge node[align=center] {} ([xshift=5em]bmc.north)
    ([xshift=5em]verif2.south) edge node[align=center] {} ([xshift=5em]fp.north)
    ([xshift=5em]verif3.south) edge node[align=center] {} ([xshift=5em]sv.north)
    %([xshift=5em]verif4.south) edge node[align=center] {$K$} ([xshift=5em]cv.north)
    ([xshift=-5em]bmc.north) edge node[align=center]  {} ([xshift=-5em]verif1.south)
    (verif) edge node {PASS} (done)
    ([xshift=5em]synth.south) edge node[align=center] {} ([xshift=5em]gp.north)
    ([xshift=-5em]gp.north) edge node[align=center] {} ([xshift=-5em]synth.south)
    (aux3) edge (synth.north);
   \path[-]
   (verif2.north) edge node[align=center] {} ([xshift=0cm]aux6)
   ([xshift=0cm]aux6) edge node[align=center] {Increase Precision} (aux5)
   (verif3.north) edge node[align=center] {} ([xshift=6.7cm]aux4)
   ([xshift=6.7cm]aux4) edge node[align=center] {Increase Unfolding Bound} (aux3);
   %(verif4.north) edge node[align=center] {} ([xshift=10.5cm]aux2)
   %([xshift=10.5cm]aux2) edge node[align=center] {Increase Sampling Rate} (aux1);

 \end{tikzpicture}
}}
\caption{CEGIS with multi-staged verification}
\label{fig:CEGIS-precision-increment}
\end{figure}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

The first approach, presented in Fig.~\ref{fig:CEGIS-precision-increment}, 
uses a multi-staged verification process. 
It starts by devising a digital controller that stabilizes the model while remaining safe for a
pre-selected time horizon ($k$) and a single initial state; then, it
employs a multi-staged verification process as follows:
(i) The first verification stage ({\sc safety}) checks that the candidate
solution, which we synthesized to be safe for at least one initial
state, is safe for \emph{all} possible initial states, i.e., does not reach
an unsafe state within $k$ steps.
(ii) The second verification stage ({\sc precision})
 restores soundness with respect to the plant precision
by using interval arithmetic \cite{moore1966interval} to validate the 
operations performed by the previous stage. 
(iii) The third verification stage ({\sc complete}) checks that the current
$k$ is large enough to ensure safety for any $k'{>}k$.  Here, we compute the
completeness threshold $\overline{k}$ for the current candidate controller 
$K$~\cite{abatecav2017}, i.e., the number of
iterations required to sufficiently unwind the closed-loop state-space
model such that the boundaries are not violated for any larger number of
iterations, and
check that $k{\geq}\overline{k}$.
%
\begin{figure}
\centering
{\scriptsize
  \resizebox{.5\textwidth}{!}
  {
\begin{tikzpicture}[scale=0.3,->,>=stealth',shorten >=.2pt,auto, semithick, ampersand replacement=\&,]
  \matrix[nodes={draw, fill=none, shape=rectangle, minimum height=.2cm, minimum width=.2cm, align=center},row sep=.8cm, column sep=.2cm] {
   \coordinate (aux1);
   \& \coordinate (aux2);
   \& ;\\
   \&\node[fill=gray!20,align=center,xshift=-1.5cm] (pre) {{\sc 1. pre-}\\{\sc processing}};
   \& \node[fill=gray!20,align=center] (abstract) {\sc 4. abstract};
   \& \coordinate (aux); \\ 
   \&
   \& \node[fill=gray!20,align=center] (synth) {\sc 2. synthesize};
   \& \node[fill=gray!20,align=center, minimum width=4cm] (verify) {\sc 3. verify ($\phi$)};
      \node[draw=none] (SAT) at ([xshift=2.7cm,yshift=-.3cm]verify)  {\sc PASS};
   \& \node[ellipse, fill=gray!20] (done) {{\sc Done}};\\
   \&
   \& \node[draw,rectangle,align=center] (KSAT) {Program \\ Search};
   \& complexnode/.pic={
     \coordinate (AA);
     \node[draw,rectangle,align=center] (AAV) at ([xshift=-1.5cm]AA.center) {Abstract\\Acceleration};
     \node[draw,rectangle,align=center] (AAC) at ([xshift=1.5cm]AA.center) {Abstraction \\ Verifier};
    }\\
  };
  \path
    (pre.east) edge node[align=center] {} (abstract.west)
    (abstract.south) edge node{} (synth.north)
    ([yshift=1em]synth.east) edge node {$K$} ([yshift=1em]verify.west)
    (aux) edge (abstract.east) 
    (verify.east) edge (done.west);
  \path
    ([xshift=+.6cm]KSAT.north) edge[left] node[xshift=-.3cm] {} ([xshift=0.6cm]synth.south) 
    ([xshift=-.6cm]synth.south) edge node[align=center,xshift=.3cm] {} ([xshift=-.6cm]KSAT.north)
    ([yshift=.5cm]AAV.east) edge node{} ([yshift=.5cm]AAC.west)
    ([yshift=-.5cm]AAC.west) edge node{} ([yshift=-.5cm]AAV.east)
    (AAC.north) edge node{} ([xshift=4.9cm]verify.south)
    ([xshift=-5cm]verify.south) edge node{} (AAV.north);
  \path[-] 
     (verify.north) edge node {C-ex} (aux);
\end{tikzpicture}
}
}
\caption{Abstraction-based CEGIS}
\label{fig:CEGARIS} 
\end{figure}

Instead of unfolding the dynamics, the second approach (illustrated in
Fig.~\ref{fig:CEGARIS}) employs {\em abstract
acceleration}~\cite{cattaruzza2015unbounded} to simultaneously evaluate all
possible progressions of the model.  This eliminates the need for a
completeness threshold, since the time horizon considered in abstract
acceleration is infinite.  An additional preprocessing phase is used to
compute a set of initial bounds on $K$ based on input constraints.  Note
that these bounds will be used by the {\sc synthesize} phase to reduce the
size of the solution space.
%
Since this model is based on abstraction, we initially perform the synthesis
over a basic model that abstracts away the complexity of the specification
and is only refined to a more complex model when a counterexample is found.
%
% Thus, we may only look at the gain of the
% model to synthesise a controller and to add further restrictions on specific sections of the reach
% tube when these prove to invalidate the solution.
%
The use of this Counterexample-Guided Abstraction-Refinement results in much
faster synthesis and verification since it is applied to both phases,
allowing for faster results whilst maintaining soundness for an infinite
time horizon.

%\textcolor{red}{We have to talk about the naive and AA back-ends, which are based on our CEGIS framework}

%---------------------------------------------------
\section{Synthesizing Digital Controllers with \tool}
%---------------------------------------------------

%---------------------------------------------------
\subsection{\tool Architecture}
%---------------------------------------------------

The proposed synthesis methodology for closed-loop digital control
systems is based on the DSSynth tool~\cite{abate2017, abatecav2017}, 
which can be split into two main stages as follows: manual and 
automated steps, as illustrated in Fig.~\ref{fig:synthesis-flow}. 
%
\begin{figure}[ht!]
\centering
\includegraphics[width=0.45\textwidth]{synthesis-flow.pdf}
\caption{Digital Controller Synthesis Methodology.}
\label{fig:synthesis-flow}
\end{figure}

In Step~1, the user selects the digital controller representation, which
can be a transfer function or a state-space model.  In Step~2, the
physical plant (in the form of Equation~(\ref{eq:transferfunction}) or
(\ref{eq:ss-example}), depending on representation) must be
given~\cite{astrom1997computer}.  Finally, in Step~3, the numerical
representation for the digital controller implementation must be set by the
user, i.e., the FWL format that defines the number of bits of the integer
and fractional parts as well as the dynamic range inputs.  The result of the
input given by the user is a specification of the physical plant.

After that, the automated synthesis process starts with Step~A, where the
\tool translates the digital system specification from the MATLAB format
into an ANSI-C program.  In Step~B, our CEGIS engine is invoked to
synthesize the digital controller w.r.t.~the digital system specification. 
Finally, in Step~C, the synthesized digital controller is produced.  The
output generated by the \tool is the synthesized digital controller
represented either in transfer function or state-space equation.  The
synthesis is considered to be \emph{successful} if a digital controller is
correctly synthesized w.r.t.~FWL effects; otherwise, if any parameter is
incorrectly defined in previous steps, or if there is no solution (no
controller) for the respective physical plant (given the implementation
requirements), then the synthesis is considered to have \emph{failed}.

Our CEGIS engine is implemented as an integrated module within the C bounded
model checker (CBMC)~\cite{cbmc}.  CBMC transforms the ANSI-C representation
of our closed-loop control system model into its internal format.  We
instrument this IR for each synthesis or verification scenario accordingly,
and use CBMC as an oracle to answer our queries.  CBMC itself relies on an
underlying SAT or SMT solver to solve verification conditions.  We~model FWL
effects, time discretization and the presence of quantization noise
explicitly using CBMC's nondeterminism API (e.g., \texttt{nondet},
\texttt{CPROVER\_assume} intrinsic functions).  Our module is included in
CBMC~5.7 and is available for
download.\footnote{\url{https://github.com/diffblue/cbmc/archive/cbmc-5.7.zip}}
We further created a VirtualBox OVA image with our tool pre-installed,
including our MATLAB-generated benchmarks and benchmark shell script with
instructions to run all benchmarks and reproduce our
experiments.\footnote{\url{www.cprover.org/DSSynth/controller-synthesis-cav-2017.tar.gz}}
The current development branch of our toolbox is available in a public
GitHub
repository.\footnote{\url{https://github.com/ssvlab/dsverifier/tree/master/toolbox-dssynth}}

The entire implementation consists of $400$ lines of MATLAB scripts for the
UI module, $585$ and $800$ lines of C code in the naive and abstract
accelerator back-ends, respectively (cf.~Section~\ref{sec:CEGIS}), $4000$
lines of C code for utility functions such as interval and fixed-point
arithmetic, as well as $1000$ lines of C++ code in the control synthesis
module in CEGIS CBMC.  We evaluate \tool using $18$ SISO control system
benchmarks.  Additionally, documentation and videos about the \tool are
available at \url{http://dsverifier.org/dsverifier/dssynth-toolbox/}.

%---------------------------------------------------
\subsection{\tool Procedures}
%---------------------------------------------------

The \tool performs the following automated steps
to synthesize a given digital controller:
%
\begin{enumerate}
%
\item \textbf{Setup}: obtains the physical plant, fixed-point format and
dynamic input ranges, and translates them to a specific structure in MATLAB.
%
\item \textbf{Parse}: obtains the digital system specification and translates 
it into an ANSI-C program.
%
\item \textbf{Execution}: obtains the ANSI-C program from the previous
(parse) step and calls our CEGIS engine as a back-end program synthesis tool
to perform the automated synthesis;
%
\item \textbf{Extraction}: obtains the ``.log'' file that is generated 
after the synthesis phase and then checks the synthesized digital-controller.
%
\item \textbf{Report}: obtains the digital controller from the previous
(extraction) step and translates it into MATLAB, for presentation to the
user.
%
\end{enumerate}

%---------------------------------------------------
\subsection{\tool Features}
%-----------------------------------------------

The features of the \tool can be summarized as follows:
%
\begin{itemize}
%
\item \textbf{Representation}: the synthesis can be 
performed for digital systems specified via 
transfer function or as state-space models.
%
\item \textbf{Fixed-Point Arithmetic}: the digital 
controller is synthesized considering FWL effects, 
using the given fixed-point numerical representation.
%
\item \textbf{Interface}: the synthesis can be executed via command-line or
graphical-user interface (GUI).  For the GUI application, the user can plot
the step response to reproduce the stability in closed-loop of the
synthesized digital controller.
%
\end{itemize}

%---------------------------------------------------
\subsection{\tool Usage}
%---------------------------------------------------

%---------------------------------------------------
\subsubsection{Command Line Version}
%---------------------------------------------------

Users must provide a digital system described as a MATLAB model using a
\texttt{tf} (for transfer function) or an \texttt{ss} (for state-space
model) command (step $1$ of Fig.~\ref{fig:synthesis-flow}).  The \tool is
called via the command line in MATLAB as

\begin{center} 
\texttt{synthesize(plant, intBits, fracBits, maxR, minR)}
\end{center}

\noindent where \texttt{plant} is the physical plant in state-space or
transfer function representation, \texttt{intBits} is the integer part,
\texttt{fracBits} is the fractional part, \texttt{maxR} and \texttt{minR}
are the maximum and minimum dynamic range, respectively (steps $2$ and
$3$ of Fig.~\ref{fig:synthesis-flow}).
%
After executing the \texttt{synthesize} command, the \tool provides the
synthesized digital controller.

%---------------------------------------------------
\subsubsection{MATLAB Application Version} 
%---------------------------------------------------

A graphical user interface application was developed (illustrated in
Fig.~\ref{fig:gui-for-tf}) to support digital controller synthesis in
MATLAB.  Its aim is to improve usability and, consequently, to attract more
engineers.  Users provide the following inputs for digital controller
synthesis within the UI: physical plant specification, fixed-point
representation and the range of the inputs.

\begin{figure}[ht]
    \centering
    \subfloat[Definition of the system representation and the physical plant]{
		\includegraphics[width=0.2\textwidth]{step1.png}
		\label{step1}}
		\hfil
    \subfloat[Definition of implementation aspects and input ranges]{
	\includegraphics[width=0.2\textwidth]{step2.png}
		\label{step2}}
		\hfil
    \subfloat[Digital controller synthesized by DSSynth]{
	\includegraphics[width=0.2\textwidth]{step3.png}
		\label{step2}}
		\hfil
    \subfloat[Step response for the synthesized digital controller]{
	\includegraphics[width=0.22\textwidth]{step4.png}
		\label{step3}}
		\hfil
    \caption{DSSynth GUI for controller synthesis using transfer function representation}
    \label{fig:gui-for-tf}
\end{figure}

%---------------------------------------------------
\subsection{Illustrative Example}
%---------------------------------------------------

To illustrate the usage of the \tool, Fig.~\ref{toolbox-usage} shows each
step of the synthesis process for a physical plant described in
Eq.~\eqref{equation_plant}, which represents a unmanned aerial vehicle
(UAV) quadcopter system~\cite{bouabdallah}.  This physical plant is
expressed by a transfer function $G(z)$, where $B(z)$ and $A(z)$ represent
the numerator and denominator, respectively, as
%
\begin{equation}
\label{equation_plant}
G(z)=\frac{B(z)}{A(z)}=\frac{-0.06875z^{2}}{z^2-1.696z+0.7089}.
\end{equation}

\begin{figure}[ht]
\scriptsize
\begin{lstlisting}[xleftmargin=.025\textwidth,xrightmargin=.025\textwidth, frame=single,]
>> num = [-0.06875 0 0];
>> den = [1.0000 -1.696 0.7089];
>> system = tf(num,den,0.002);
>> y = synthesize(system,8,8,1,-1);
>> SYNTHESIS SUCCESSFUL
>> y = 
>>  -0.9983z^2 + 0.09587z + 0.1926
>>  --------------------------------
>>       z^2 + 0.5665z + 0.75
\end{lstlisting}
\vspace{-0.2cm}
\caption{Synthesis of a digital controller for the physical plant defined in Eq.(\ref{equation_plant}) in MATLAB, with a fixed-point format  $\left\langle 8,8\right\rangle$}
\label{toolbox-usage}
\end{figure}

In Fig.~\ref{toolbox-usage}, ``num'' represents $B(z)$ and ``den'' represent
$A(z)$, while $y$ represents the synthesized stable digital controller,
which is defined by Eq.(\ref{equation_controller}).
%
\begin{equation}
\label{equation_controller}
C(z)=\frac{-0.9983^{2}+0.09587z+0.1926}{z^2+0.5665z+0.75}.
\end{equation}

A digital system is stable iff all of its poles are inside the
$z$-plane unity circle, i.e., the poles must have a modulus less than
one~\cite{astrom1997computer}.  In order to compute the poles for the
digital system represented by Equations~\eqref{equation_plant} and
\eqref{equation_controller}, the user can invoke the \texttt{feedback} and
\texttt{series} commands in MATLAB~\cite{matlab-toolbox} to obtain a
general equation regarding the transfer function representation, as
illustrated in Fig.~\ref{combine-controller-plant}.

\begin{figure}[ht]
\scriptsize
\begin{lstlisting}[xleftmargin=.025\textwidth,xrightmargin=.025\textwidth, frame=single,]
>> num = [-0.99832 0.09587 0.1926];
>> den = [1 0.5665 0.75];
>> controller = tf(num,den,0.002);
>> num = [-0.06875 0 0];
>> den = [1.0000 -1.696 0.7089];
>> plant = tf(num,den,0.002);
>> sys = feedback(series(controller, plant),1)
>> sys =
>>        0.06863z^4 - 0.006591z^3 - 0.01324z^2
>> ---------------------------------------------------
>> 1.069z^4 - 1.136z^3 + 0.4849z^2 - 0.8704z + 0.5317
\end{lstlisting}
\vspace{-0.2cm}
\caption{General equation for the closed-loop control system using the series configuration as
illustrated in Fig.~\ref{fig:typical-control-system}}
\label{combine-controller-plant}
\end{figure}

The general equation that represents the closed-loop control system, using
the series configuration as illustrated in
Fig.~\ref{fig:typical-control-system}, is described by
Eq.~\ref{equation_general} as follows
%
\begin{equation}
\small
\label{equation_general}
\frac{N(z)}{D(z)}=\frac{0.06863z^{4} - 0.006591z^{3} - 0.01324z^{2}}{1.069z^{4} - 1.136z^{3} + 0.4849z^{2} -0.8704z + 0.5317}.
\end{equation}

For the stability check, if any root of the denominator polynomial $D(z)$ in
Eq.~\ref{equation_general} has modulus equal or greater than one, then the
system is unstable; otherwise, it is stable.  Indeed, the roots computed by
the function \texttt{roots} in MATLAB~\cite{matlab-toolbox} result in the
following poles as illustrated in Fig.~\ref{roots-of-dz}, which means that
DSSynth has synthesized a stable digital controller.

\begin{figure}[ht]
\scriptsize
\begin{lstlisting}[xleftmargin=.025\textwidth,xrightmargin=.025\textwidth, frame=single,]
>> den = [1.069 -1.136 0.4849 -0.8704 0.5317]
>> r = roots(den)
>> r =
>>  -0.2912 + 0.8061i
>>  -0.2912 - 0.8061i
>>   0.8225 + 0.0219i
>>   0.8225 - 0.0219i
\end{lstlisting}
\vspace{-0.2cm}
\caption{Roots of the polynomial $D(z)$ given by Eq.~\ref{equation_general}.}
\label{roots-of-dz}
\end{figure}

Simulating the digital system represented by Equations
\eqref{equation_plant} and \eqref{equation_controller} to validate and
reproduce the system stability, the user can obtain the step response using
the command \texttt{dstep} in MATLAB, and then observe that the plotted
graph indicates a stable digital system (Fig.~\ref{step-response}).

\begin{figure}[ht]
  \includegraphics[width=0.5\textwidth]{step-response.eps}
  \caption{Step response for Eq.~\eqref{equation_plant} describing a stable system for the UAV quadcopter system}
  \label{step-response}
\end{figure}

%---------------------------------------------------
\section{\tool Experimental Evaluation}
%---------------------------------------------------

%---------------------------------------------------
\subsection{Benchmarks Description}
\label{benchmarks-description}
%---------------------------------------------------

Our evaluation consists of $18$ SISO control system benchmarks extracted
from the literature~\cite{abate2017, abatecav2017, bouabdallah, acrobot,
cstr, KOKOTOVIC198023, gajic2008optimal, Franklin15, maglev, converters,
CTMS, DBLP:journals/tc/BessaIPCF17, DBLP:journals/dafes/BessaICF16}, which
include physical plants for a unmanned aerial vehicle (UAV), a cruise
control system, a spring-mass damper, a satellite, a direct current (DC)
servo motor, a helicopter, the inverted pendulum, a pendulum, magnetic
suspension, a magnitezed pointer, 1/4 car suspension, a computer tape
driver, a flexible beam, a guidance system, a USCG cutter tampa, a
continuous stirred tank reactor, a voltage regulator, and an acrobot control
system.

%The first set of benchmarks uses the discrete model of a cruise control system for a car, and accounts for rolling friction, aerodynamic drag, and the gravitational disturbance force. %cruise-control
%The second set of benchmarks considers the discrete model of a simple spring-mass damper plant. %spring-mass damper
%A third set of benchmarks uses the discrete model for satellite attitude dynamics, which require attitude control for orientation of antennas and sensors w.r.t. Earth. %satellite
%The fourth and fifth set of benchmarks describe the discrete model of a DC servo motor velocity dynamics. %servor-motor
%The fifth set of benchcmarks corresponds an USCG cutter Tampa Heading Angle that describes the heading angle  dynamics of a Coast Guard cutter.
%The sixth set of benchmarks contain a flexible beam system that is a plant usually employed in vibration studies. It is  composed by a flexible metallic structure with piezoceramic actuators and sensors.%flexible beam
%The seventh set of benchmarks describes the discrete model for the Helicopter Longitudinal Motion, which provides the longitudinal motion dynamics of a helicopter. %helicopter
%The eighth set of benchmarks contains the discrete model for the known Inverted Pendulum, which describes a pendulum dynamics with its center of mass above its pivot point. %inverted pendulum
%The ninth set of benchmarks contains the Magnetic Suspension discrete model, which describes the dynamics of a mass that levitates with support only of a magnetic field. %magnetic suspension
%The tenth set of benchmarks contains the Computer Tape Driver discrete model, which describes a system to read and write data on a storage device. %tape driver
%The eleventh set of benchmarks contains the unmanned aerial vehicle (UAV) discret model, in particular, a quadrotor attitude system. %uav
%The twelfth set of benchmarks contains the 1/4 Car Suspension plant presents a physical model that connects a car to its wheels and allows relative motion between both parts. %1/4 car suspension
%The thirteenth set of benchmarks contains the Magnetized Pointer plant describes a physical model employed in analogue gauges and indicators that is rotated through interaction with magnetic fields. %magnitezed pointer
%The fourteenth set of benchmarks contains the pendulum which is system composed by a swinging point mass suspended from  a frictionless pivot through a negligible mass rod. %pendulum
%The fifteenth set of benchmarks  contains the Acrobot benchmark corresponds to the model of an two-link  underactuated robot manipulator, named Acrobot, linearized by partial  feedback linearization~\cite{acrobot}. %acrobot
%The sixteenth set of benchmarks contains a voltage regulator benchmark~\cite{KOKOTOVIC198023} that consists in a linear model of a synchronous machine connected to an  infinite bus.
%The seventeenth set of benchmarks contais a continuous stirred tank reactor which describes a linear model for PH dynamics of a PH  neutralization process of an aqueous solution of sodium acetate with  hydrochloric acid in a continuous flow stirred-tank reactor.%cstr
%And finally, the last set of benchmarks contaians the Guidance Control System benchmark which is another high-order system  model used by Chen et al.~\cite{CHEN1979389}  in order reduction  techniques. Guidance systems are used to determine the path of an autonomous vehicle movement.

%-----------------------------------
\subsection{Experimental Setup}
\label{experimental-setup}
%-----------------------------------

For all evaluated benchmarks, the signal input range related to the control
system is between $-1$ and $1$.  The implementation features used for
digital controller are: $8$~bits for the integer part (including the sign
bit) and $8$~bits for the fractional part.  We have run the \tool using
CEGIS with multi-staged verification (cf.~Section~\ref{sec:CEGIS}).  All
experiments with \tool v$1$.$0$.$0$ were conducted on an otherwise idle
Intel Core i$7-2600$ $3.40$ GHz processor, with $24$ GB of RAM, running
Ubuntu $64$-bits.

%---------------------------------------------------
\subsection{Objectives}
%---------------------------------------------------

The main objectives of our experimental evaluation is to (1)~evaluate the
\tool performance to produce stable and safe controllers, and also
(2)~confirm the stability and safety of the synthesized digital controllers
outside of our model using MATLAB.

%, using both the transfer function and state-space representations.

%---------------------------------------------------
\subsection{Results}
%---------------------------------------------------

%\textcolor{red}{Why are the experiments only mentioning stable controllers? Up until this point, we discuss stable and safe controllers.}

For our experiments, we observed that the \tool correctly synthesizes stable
and safe digital controllers for our set of benchmarks represented by
transfer functions or state-space equations.  We employed a diverse set of
real-world benchmarks extracted from the control literature.  However, note
that this set of benchmarks is still limited within the scope of this paper,
and the performance may not be generalized to other benchmarks.  The median
runtime for our benchmarks set is $124$\,s, and the average synthesis time
amounts to $35.5$\,s for closed-loop control systems represented by
state-space equations, while for transfer functions, the average synthesis
time amounts to $123.6$\,s.

Additionally, we are able to reproduce the stability and safety for all
evaluated control system benchmarks.  As an example, we have reproduced the
stability property in the synthesized controller for a DC motor plant
described in Eq.~\eqref{tf-synth-controller}, simulating the step response
in MATLAB.  Fig.~\ref{tf-step-response} shows that the digital controller
synthesized by the \tool stabilizes the closed-loop control system using the
series connection (illustrated in Fig.~\ref{fig:typical-control-system}) for
the DC motor plant.  We have applied this procedure to all synthesized
controllers to confirm the reliability of our experimental results.
%
\begin{equation}
\label{tf-synth-controller}
C(z)=\frac{-0.3466796875z+0.015625}{-0.5z^{2}+0.19921875z}.
\end{equation}

\begin{figure}[ht]
  \includegraphics[width=0.5\textwidth]{tf-step-response.eps}
  \caption{Step response for the Eq.~\eqref{tf-synth-controller} describing a stable system for the DC motor plant.}
  \label{tf-step-response}
\end{figure}

Regarding the state-space representation, we have reproduced the stability property for 
the closed-loop control system of a pendulum represented by Eq.~\eqref{ss-synth-controller}. 
%
\begin{equation}
\label{ss-synth-controller}
\begin{split}
x(n+1) &= A x(n) + B u(n)
\\
y(n) &= C x(n) + D u(n), 
\end{split}\
\end{equation}
%
\noindent where,
%
$$A = \left[\begin{array}{cc}-1.999909737361046&-1.000000000000000\\1&0\end{array}\right],$$
$$B = \left [\begin{array}{c}4\\0\end{array}\right],$$
$$C = \left [\begin{array}{cc}-1.756887232846049&-1.7494556607764090\end{array}\right],$$
$$\quad D = \left [\begin{array}{c}9.8\end{array}\right].$$

For this particular closed-loop control system, the feedback matrix (see
Eq.~\eqref{ss-synth-feedback}) generated by the \tool also leads to a stable
system, as sillustrated in Fig.~\ref{ss-step-response}.
%
\begin{equation}
\label{ss-synth-feedback}
\begin{split}
K &= \left [\begin{array}{cc}-0.5898&-0.1914\end{array}\right]
\end{split}\
\end{equation}

\begin{figure}[ht]
  \includegraphics[width=0.5\textwidth]{ss-step-response.eps}
  \caption{Step response for the Eq.~\eqref{ss-synth-controller} describing a stable system for the pendulum plant.}
  \label{ss-step-response}
\end{figure}

%The digital controllers synthesized by \tool are all stable using the MATLAB step response
%as a validation procedure, which means that our results are reliable for control engineers. 
The synthesized controllers are also confirmed to be safe outside of our model
representation using MATLAB, which means that our results are reliable for control engineers. 
In particular, we unfold the system dynamics up to a given depth, 
compute the system input and output via the feedback matrix, and then check the states bounds.
 A link to the full experimental environment, including the scripts, benchmarks, 
 videos, documentation and the tool, is provided online.\footnote{\url{www.dsverifier.org}}

%--------------------------------------
\section{Conclusions}
%--------------------------------------

The \tool can automatically synthesize stable and safe digital controllers
for intricate dynamic physical plants, represented in MATLAB either as a
transfer function or a state-space equation.  The \tool is the first fully
automated synthesis tool that is algorithmically and numerically sound,
considering various error sources in the implementation of the digital
control algorithm and in the computational modeling of plant
dynamics~\cite{abate2017, abatecav2017}.
%
Additionally, given the current literature in controller synthesis, there is
no other MATLAB toolbox for synthesizing stable and safe digital controllers
for physical plants that considers all aspects of their digital
implementation.
%
As future work, the \tool will perform synthesis considering additional
requirements (e.g., performance) and will be combined with the
DSVerifier/DSValidator Toolboxes~\cite{issta2017,dsvalidator} to
verify/validate further controller properties.

\bibliographystyle{IEEEtran}
\bibliography{references}

\end{document}