### 2018-04-29 SYNT_abstract/abstract.tex
abstract

\documentclass[submission]{eptcs}
\providecommand{\event}{SYNT 2018} % Name of the event you are submitting to
\usepackage{breakurl}             % Not needed if you use pdflatex only.
\usepackage{underscore}           % Only needed if you use pdflatex.

\title{Safe, Automated and Formal Synthesis of Digital Controllers 
for Continuous Plants}
%\author{Elizabeth Polgreen
%\institute{NICTA\\ Sydney, Australia}
%\institute{School of Computer Science and Engineering\\
%University of New South Wales\thanks{A fine university.}\\
%Sydney, Australia}
%\email{rvg@cs.stanford.edu}
%\and
%Co Author \qquad\qquad Yet S. Else
%\institute{Stanford Univeristy\\
%California, USA}
%\email{\quad is@gmail.com \quad\qquad somebody@else.org}
%}
%\def\titlerunning{Synthesis of Digital Controllers}
%\def\authorrunning{R.J. van Glabbeek, C. Author \& Y.S. Else}
\begin{document}
\maketitle

\begin{abstract}
We present a sound and automated approach to synthesizing safe,
digital controllers for physical plants represented as linear,
time-invariant models. The synthesis
accounts for errors caused by the digitization effects introduced by
digital controllers operating in either fixed-point or floating-point arithmetic. 
Our approach uses counterexample-guided inductive
synthesis (CEGIS): an inductive generalisation phase produces a
possible solution that is safe for some possible initial conditions but may
not be safe for all initial conditions. Safety for all initial conditions
is then verified either
via bounded model checking or abstract acceleration; if the verification step fails, a
counterexample is provided to the inductive generalisation and the
process iterates until a safe controller is obtained.  We demonstrate
the practical value of this approach by automatically synthesizing
safe controllers for physical plant models from the digital control
literature.
\end{abstract}

\section{Introduction}

Embedded control systems using fixed and floating-point arithmetic 
have become widespread
as the availability of low-cost devices that can perform highly
non-trivial control tasks has increased. Correct synthesis of
control software for such platforms is non-trivial due to the digital representation of continuous quantities 
that is used by the controller. This digital representation introduces 
errors due to finite-precision arithmetic, time discretization and 
A/D - D/A conversions.

%
Given an LTI model of a physical (continuous) plant, we present two automated approaches for generating
correct-by-construction digital controllers that address all these
challenges and satisfy a safety property for the physical plant. 
Both approaches make use of CounterExample-Guided
Inductive Synthesis (CEGIS)~\cite{jha-icse10,
  DBLP:conf/asplos/Solar-LezamaTBSS06}.  CEGIS is an
iterative process, where each iteration performs inductive
generalisation based on counterexamples provided by a verification module. 
The inductive generalisation step uses information 
about a limited number of inputs to compute a candidate solution
for all the possible inputs. Our two instantiations of CEGIS are
described next.

\emph{The first approach} starts by devising a
digital controller that stabilizes the system's model, while remaining safe for a
pre-selected time horizon and a single initial state; then, it verifies
unbounded-time safety by unfolding the dynamics of the LTI model, considering the
full set of initial states, and checking a \emph{completeness
threshold}~\cite{DBLP:conf/vmcai/KroeningS03}, i.e., the number of
iterations required to sufficiently unwind the closed-loop state-space
model such that the boundaries are not violated for any larger number of
iterations.  As~it requires unfolding up to the completeness threshold, this
approach can be computationally expensive if the completeness threshold is large.

\emph{The second approach}
employs \emph{abstract acceleration}~\cite{cattaruzza2015unbounded} to
evaluate all possible progressions of the LTI model simultaneously. 
This approach uses \emph{abstraction refinement},
enabling us to always start with a very simple description regardless of the
dynamics complexity, and only expand to more complex models
when a solution cannot be found.

We provide experimental results showing that both our approaches are able to
efficiently synthesize safe controllers for a set of intricate physical
plant models taken from the digital control literature.

\bibliographystyle{eptcs}       
\bibliography{../automatica/paper}  
\end{document}