### 2019-07-09 automatica/response.tex
response and manuscript

\documentclass{article}

\setlength\parindent{0pt}

\usepackage{amssymb,latexsym,amsfonts,amsmath, amsthm,color,mathrsfs}
\usepackage[dvipsnames]{xcolor}
\usepackage{amsthm}
\usepackage{mathtools}
\usepackage{dsfont}
\usepackage{mathtools}
\usepackage{todonotes}
%\usepackage{utf8}
\usepackage{fullpage}
\def\SS#1{{\textcolor{red}{ {\bf SS:} #1}}}
\renewcommand{\b}{\textcolor{blue}}
%\newcommand{\red}{\textcolor{red}}
%
\newcommand{\vold}[1]{\colorbox{red!20}{\bfseries #1}}
\newcommand{\vnew}[1]{\colorbox{blue!30}{\bfseries #1}}

\renewcommand{\t}{\new{(t)}}
\newcommand{\tp}{\new{(t+1)}}
\newcommand{\fei}[1]{{\color{purple}#1 \color{black}}}
\newcommand{\rev}[1]{\textcolor{blue}{#1}}

\begin{document}
\begin{center}
{\Large Responses to the Reviews}\\\vspace{0.3cm}
\today
\end{center}
\vspace{1cm}
Dear Associate Editor and Reviewers,\vspace{0.3cm}
\\
We would like to deeply thank you for reviewing our paper and providing us with the constructive comments and suggestions. We have carefully addressed
all the comments and incorporated them in the revised manuscript.

We present below a detailed discussion of all comments, questions, and concerns raised by the reviewers. The major modifications are highlighted in
blue in the revised manuscript. For clarity,  we also use \vold{v1} and \vnew{v2} in the response letter for referring respectively to the previous
and current versions of the manuscript.
%I do not like color coding that much. And definitely not with red and blue.
\\
\\
With kind regards,\vspace{0.1cm}
\\
The Authors
%
\newpage

\section{General response}

\section{Response to Reviewer (1)}
\bigskip

{\bf Comment: }{\itshape This paper presents an approach for synthesizing 
digital/software controllers for continuous plants (linear time invariant / LTI) 
such that they satisfy safety/stability requirements, as would show up in closed-loop 
control of cyber-physical systems (CPS). The approach extends earlier work of the authors, 
the primary of which in my view is toward floating-point implementation as opposed to only fixed-point. 
I did not entirely buy that using CEGIS is a contribution, as from looking over the earlier related publications 
(e.g., ASE'17, CAV'17, HSCC'17, SYNT'18 abstract), all of them use CEGIS.}

\vspace{1em}

{\bf Response: }

This journal submission encompasses and summarises the mentioned conference articles (CAV, HSCC, ASE) or workshop abstract (SYNTH). 
It is a more mature, complete, and comprehensive presentation of the conference material. 
Of these conference articles, the CAV/HSCC contributions presented the theory underlying the approach, 
whereas the ASE is a tool paper. 
The application of the CEGIS architecture to the synthesis of digital controllers for continuous plants is indeed a novel contribution of this line of work.

\todo[inline]{Do we actually say that CEGIS is a contribution? We should change that. Check that we don't say CEGIS is a contribution. AA: I think the reviewers complain about our earlier conference publications. We should rectify this view. I've added the text above. }

\vspace{2em}
{\bf Comment: } {\itshape he results are evaluated on about a dozen and a half examples from the controls literature all with relatively small
dimensionality (2 to 5 real variables in the continuous state space), and while the authors claim the software artifacts and benchmarks are available
for evaluation, when this reviewer browsed to the link (http://www.cprover.org/DSSynth/), a page with "TBD" was shown. I did a little digging and
eventually found what I presume is the same / similar artifacts here: https://github.com/ssvlab/dsverifier and/or http://dsverifier.org/}

\vspace{1em}
{\bf Response: }

Fixed this now

\todo[inline]{Elizabeth: say what the problem was, fix this if possible, then thank the reviewer for pointing this out. }

\vspace{2em}
{\bf Comment: } {\itshape On the control theory side, the approach uses standard results to synthesize stabilizing controllers for LTI systems,
specifically linear state feedback controllers (u=Kx with a possible reference signal to track), under the assumption the plant is controllable. It
seems the approach requires full-state observability as well, which is a fairly standard assumption as one can typically implement various filters /
estimators (Kalman, Leuenberger, etc.) to ensure state observability assuming detectability of the plant, but the authors should clarify this point to
make the assumptions more precise.}

\vspace{1em}
{\bf Response: }
We do need full observability of the (states of the) model, and have clarified this in Section 3.1, as follows: {\it We assume full observability of the states of the model, namely the output states correspond with the model variables.  } We recall this assumption later in the text.

\vspace{2em}
{\bf Comment: } {\itshape Theorem 2 states completeness of the approach, specifically that there exists a finite k for which BMC can be done up-to to
provide a guarantee of future safety for all larger k' greater than k. Assuming the benchmarks all work up to some completeness threshold, I would like to see
the completeness threshold for each benchmark in table 1. The reason being, I find this result a bit surprising in the given context (albeit not
unbelievable given the focus on LTI) and would like to see further evidence illustrating the point beyond the fairly informal proof, and it would
greatly enhance the existing discussion of timeouts, etc. related to completeness thresholds being too high for some cases.}

\vspace{1em}

{\bf Response: }

We have added the completeness threshold for each benchmark in the table. Note that the completeness threshold is dependent on the controller, 
and a controller with a smaller completeness threshold may exist.

The proof of Theorem 2 has been elaborated and made more comprehensive and detailed.

\todo[inline]{Elizabeth: add completeness thresholds to table}

\vspace{2em}

{\bf Comment: }
{\itshape I would like to see some additional discussion and/or comparison to hybrid systems reachability. While the focus of this work is on lower-level
software synthesis, if the approach is to be believed in the context of closed-loop systems satisfying specifications, then I would like some
discussion/comparison to that vast existing literature. Specifically, I presume the reason the benchmarks are not scalable beyond ~n=4 dimensionality
(as the n=5 benchmark appears to timeout), which is trivially small at this stage for linear systems (e.g., given a recent HSCC'19 paper on verifying
$R^{billion}$ linear systems, or at least n=100 to n=1000 with standard approaches like SpaceEx), is due to overapproximation error growth with the
interval representation of plant behaviors, whereas most scalable approaches utilize better data structures for representing geometric sets
(zonotopes, support functions, star sets, Taylor models, etc.). This likewise limits the possible applicability of the
approach for nonlinear plants, although there are a host of technical issues in that potential extension. }

\vspace{1em}
{\bf Response: }

We have reviewed relevant literature on controller synthesis based on finite abstractions, cf. Belta and Tabuada.

Literature on reachability analysis tackles a different problem, namely one of verification, rather than one of control synthesis (which this work focuses on).

Admittedly reachability analysis tool scale quite well (as the reviewer points out), however this is under clear assumptions on model dynamics.

One might argue that reachability analysis tools could be embedded within the CEGIS framework, however they currently lack full soundness: namely they are algorithmically sound, but not numerically sound. On the contrary our work is sound on both sides. (Shall we mention that indeed we do reachability analysis via Axelerator in CAV17?)

\vspace{2em}
{\bf Comment: }
{\itshape Finally, while checking the $2^n$ vertices
are sufficient (theorem 1), I doubt this is necessary, as e.g. the star set approaches avoid this blow-up that would also limit the dimensionality
the approach can work toward.}
\vspace{1em}

{\bf Response: }

\rev{It is probably true that the check on all vertices might not be necessary in general, that that this check might be simplified in special cases or under special conditions. The check however, as it currently is implemented, does not represent the bottleneck of our overall method, so we can safely rely on its current form, as it stands. }

\vspace{2em}

{\bf Comment: }
{\itshape While I may have missed it, I didn't see how the controller gain K is initially selected. I presume under the given assumptions a standard approach
such as pole placement, but this should be clarified}
\vspace{1em}
{\bf Response: }

The controller gain is initially nondeterministically selected by the bounded model checking synthesis step, i.e., the SAT solver returns the first assignment to the controller
values it finds
that satisfies the boolean formula which encodes the specification, by using Conflict Driven Clause Learning.
We have added the following sentence to clarify this:
\todo[inline]{add this}

\vspace{2em}
\todo[inline]{fix reviewer 1 typos}

\section{Response to Reviewer (3)}

{\bf Comment: }{\itshape
The verification condition, given in Theorem 1, is wrong. Here is an example system, with dynamics, initial set, and unsafe set.

\begin{verbatim}
x+ = -0.5*y
y+ = 0.5*x
\end{verbatim}

Rotation by 90 degrees in counterclock wise and shrinking it to the origin by 0.5.
\begin{verbatim}
initial set is 0 <= x <= 1 and 0 <= y <= 1.
unsafe set -3/20 <= x <= -2/20 and -3/20 <= y <= -2/20.
\end{verbatim}

It is easy to see that after two transformations, the initial set would be
\begin{verbatim}
-0.25 <= x <= 0 and -0.25 <= y <= 0.
\end{verbatim}
The unsafe set lies completely within the reachable set, but none of the edges in the hyperrectangle is unsafe. The next time the initial set reaches
this quadrant, the reachable set is
\begin{verbatim}
-1/(2^6) <= x <= 0
-1/(2^6) <= y <= 0.
\end{verbatim}

I believe that for this case, the K returned from completeness is 4 or 5.

The same theorem is also present in [1]. The proof, upon careful inspection, is not a proof of the statement given in the theorem. It only proves
that the reachable set is the convex set obtained by the trajectories of the corner states. However, it does not prove that if the trajectories are
safe, their convex combination is also safe. In fact, as given in the above counterexample, the vertices of polytope are safe does not mean that the
convex combination is safe.

I think this bug needs to be fixed in the BMC part of the verifier. Instead of checking the individual trajectories, you should in fact compute the
convex combination of trajectories (reachable set) and then perform the verification.}

\vspace{1em}
{\bf Response: }
\todo[inline]{Alessandro: Response to this (Theorem 1)}
\vspace{2em}

\rev{
Many thanks for this note. 
Notice that in the submitted manuscript we had focused on special shapes of initial and safe sets, namely (convex) hyper-boxes, which included the origin. 
In the amended draft, we have better clarified these assumptions, and 
we have emphasised that the initial set is assumed to lie within the safe set: indeed, the contrary would trivially result in an unsafe condition. 
In conclusion, recall that the sets are hyper-boxes, thus convex, and that in view of the linear dynamics of the considered models, in practice both are assumed to contain the origin.  }

\section{Response to Reviewer (4)}

{\bf Comment:  } {\itshape The safety property in Eqn 3 focuses on interval constraints. One can generate auxiliary variables such that range constraints over them can encode
more complex safety properties. It would be nice to bring this up in the discussion after equation 3 that $x_i$ need not be just $i$-th component of the
state.  Can these also incorporate temporal properties (say STL properties with bounded time-horizon) without having to unroll the transition and
create auxiliary variables that track history explicitly?}

\vspace{1em}
{\bf Response: }

\todo[inline]{Eli: can we generalise according to the suggestion of the reviewer?}

\vspace{2em}
{\bf Comment: } {\itshape The paper would benefit with a restructuring of the discussion. Figure 2 and 3 will be more useful
earlier in the paper (Section 1 or 3).}

\vspace{1em}
{\bf Response: }
We have moved these figures to [] 
\todo[inline]{move figure}
\vspace{2em}

{\bf Comment: } {\itshape
Some other relevant applications of CEGIS to CPS/dynamical systems include: Synthesizing switching logic for
safety and dwell-time requirements in ICCPS, 2010; Synthesis of optimal switching logic for hybrid systems in EMSOFT, 2011, Synthesis of Optimal
Fixed-Point Implementation of Numerical Software Routines in NSV'13, Bridging boolean and quantitative synthesis using smoothed proof search,
POPL'14, Synthesis of fixed-point programs in EMSOFT'13}

\vspace{1em}
{\bf Response: }
We have added a discussion about these applications and results to the related work as follows:
\todo[inline]{add discussion}

\vspace{2em}
{\bf Comment: } {\itshape Could you add one example to the end of Section 4 to show $R_1$, $R_2$, $R_3$, $R_4$ for a simple LTI system?

\vspace{1em}
{\bf Response: }

\todo[inline]{Eli: can you clarify?}

\vspace{2em}
{\bf Comment: } {\itshape 
Section 4 is one the most interesting parts
of the paper and its current presentation is a bit sketchy. What are $m_{ij}$?} }

\vspace{1em}
{\bf Response: }

\todo[inline]{Eli: can we elaborate?}

\bibliographystyle{alpha}
\bibliography{}

\end{document}